### 2016-03-07 paper/paper.tex
Added latex version of the paper.

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,english,american,draftfoot]{ulb_thesis}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{float}
\usepackage{url}
\usepackage{slashed}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{makeidx}
\makeindex

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\floatstyle{ruled}
\newfloat{algorithm}{tbp}{loa}[chapter]
\providecommand{\algorithmname}{Algorithm}
\floatname{algorithm}{\protect\algorithmname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
 \input{ulb_ntheorem.std}
\newenvironment{lyxlist}[1]
{\begin{list}{}
{\settowidth{\labelwidth}{#1}
 \setlength{\leftmargin}{\labelwidth}
 \addtolength{\leftmargin}{\labelsep}
 \renewcommand{\makelabel}[1]{##1\hfil}}}
{\end{list}}

\makeatother

\usepackage{babel}
\addto\captionsenglish{\renewcommand{\algorithmname}{Algorithm}}

\begin{document}
\selectlanguage{english}%

\title{Engineering a Runtime System for AQL}

\selectlanguage{american}%

\author{Andrea Morciano}

\degreeaward{Docteur en XXX\\
(Département XXX, Faculté XXX)}
\maketitle
\begin{dedication}
To my parents.\end{dedication}
\begin{acknowledgments}
Thank you, my advisor!
\end{acknowledgments}
\vspace*{4cm}

\emph{``A great citation here}''

\begin{flushright}
X
\par\end{flushright}
\begin{abstract}
abstract of the thesis
\end{abstract}
\tableofcontents{}

\listoffigures

\listoftables

\mainmatter

\include{intro}

\chapter{Information Extraction}

\section{Definition}

Information Extraction (IE) is the discipline that addresses the task
of extracting structured information from unstructured sources automatically.
Sources usually take the form of text documents. The most explored
aspect of IE is the extraction of \emph{named entities.} Another main
topic, that has become object of research only in recent years, is
the extraction of \emph{relationships} between entities. IE is a field
with contributions from many different communities: Machine Learning,
Information Retrieval, Database, Web, Document Analysis. The interest
in towards it is motivated by the evergrowing amount of data that
is generated by our society, that is at a point in which manual analysis
is infeasible. Thus IE promises to bring value to many application
domains, most notably the enterprise world and the Web. There are
two main approaches to IE: \emph{rule-based }and \emph{statistical}.

\section{Example Applications}

\paragraph{News Tracking}

The activity of tracking events in news articles. It can result in
a lot of useful services, like automatic creation of multimedia news,
linking articles to information pages on the entities found, etc.

\paragraph{Data Cleaning}

Extracting structured forms from flat data strings (containing, e.g.,
addresses). It allows more effective decuplication of information,
among the other things.

\paragraph{Citation Databases}

Articles, conferences sites, individual research sites and similar
are explored to obain formatted citations of publications, later stored
in publicly accessible databases. They are capable to forward references
and may provide aggregate statistics and scoring information. See
for instance Google Scholar (\url{https://scholar.google.it/}).

\paragraph{Relationship Web Search}

Relationship extraction would be a very useful feature to integrate
into web search engines, as the keyword search that they offer at
present time is only good for entity identification.

\section{Issues}

\subsection{Accuracy}

We measure the accuracy of an IE task with two quantities:
\begin{itemize}
\item \emph{precision}: the percentage of correctly extracted entities among
the extracted entities;
\item \emph{recall}: the percentage of entities extracted among all the
existing entities in the input source.
\end{itemize}
The main difficulties to achieving a good level of accuracy are:
\begin{itemize}
\item the availability of a great set of clues that might be very different
in nature (e. g. orghographic properties, part of speech, typical
words, etc.) and that might be difficult to combine;
\item the difficulty of finding out missed extractions;
\item the fact that with the advancement of research the extracted data
structures keep increasing in complexity (for instance it is becoming
more difficult to identify the boundaries of an entity in a text).
\end{itemize}
While we are able to reach a very good level of accuracy for entity
extraction ($90\%$), relationship extraction is still quite unreliable
($50-70\%$ accuracy), mainly due to its intrinsic complexity, superior
to that of the task of entity extraction.

\subsection{Running Time}

In its initial phase IE was interesing only to research communities.
With the enormous amount of data they need to process, companies are
now concerned with IE too, and \emph{scalability} has become a central
issue, while many IE systems don't address the problem of efficiently
carrying out extraction tasks with sufficient attention. Efficiency
is influenced most notably by the efficiency of the following tasks:
\begin{itemize}
\item filtering documents in order to actually examine only the ones that
have good chances to contain the desired information;
\item focusing on the parts of a document that contain relevant information;
\item processing steps, like database querying, that are typically very
expensive and that might have to be performed on the selected pieces
of input.
\end{itemize}
Recently many solutions have been proposed to target the scalability
issue; one of these is indeed SystemT.

\section{Entity Extraction\label{sec:Entity-Extraction}}

Named entities are elements of interest in a text. Example of entities
are person names, street addresses, institutions, countries, and so
on. In this section and in the next one, I assume that the output
of an extraction task is a series of labels inserted into an input
text document (that becomes \emph{annotated}), although there are
other possibilities.

\subsection{Rule-based Methods}

Rule-based methods employ sets of predicates that are ``fired''
independently. When a portion of input text satisfies a predicate,
the predicate's associated action is executed. A predicate can be
represented in the following generic form:

\begin{equation}
\textrm{"Contextual Pattern}\longrightarrow\textrm{Action"}
\end{equation}
\\
Contextual patterns are a way to identify entitities by exploiting
their properties or the context in which they are usually met. The
most common formalism used to express them is that of \emph{regular
expressions} over tokens of the input. Actions mark the entities that
have been identified, and might consist in storing them in a database
or adding delimiters directly in the text.\\
Most systems in this category present a cascaded structure: an input
document goes through a series of processing phases, and the output
of a phase is the input of the next one. A famous example is that
of cascading grammars: formal grammars are evaluated in sequence on
the input.

As I said, a contextual pattern seeks for (groups of) tokens that
have certain features. In the case of entity extraction, features
can be classified in the following categories:
\begin{itemize}
\item string representation;
\item orthography (e.g. small case, mixed case, number, etc.);
\item part of speech;
\item list of dictionaries they're contained into;
\item annotation obtained in previous phases.
\end{itemize}
Rules can be hand-coded by experts or learned through learning algorithms.
In the second case the goal is to cover each of the entities of interest
in the training set with at least one rule. The obtained rules should
have good recall and precision on new input. Moreover, when learning
a set of rules, we would like to achieve a good level of \emph{generalizability},
that is: we would like to find the minimum set of rules accounting
for the maximum portion of training data, with high precision. There
are two main strategies for rule learning: \emph{bottom-up} (start
from very specific rules and make them more and more general) and
\emph{top-down} (start with rules covering all existing instances,
then specialize them).

\subsubsection{Example Rules}

For the examples I use an idealized syntax, as in REF.
\begin{example}
Consider the task of identifying all mentions ISO standards in a text.
A rule for this purpose could be:

\begin{multline}
\left(\left\{ \textrm{String}=\textrm{"ISO"}\right\} \left\{ \textrm{String}=\textrm{"/IEC"}\right\} \left\{ \textrm{?}\right\} \left\{ \textrm{String}=\textrm{"/ASTM"}\right\} \left\{ \textrm{?}\right\} \right.\\
\left.\left\{ \textrm{Orthography type}=\textrm{number}\right\} \left\{ \left\{ \textrm{String}=\textrm{":"}\right\} \left\{ \textrm{Orthography type}=\textrm{number}\right\} \right\} \left\{ \textrm{?}\right\} \right)\\
\longrightarrow\textrm{ISO Standards}
\end{multline}

\end{example}

a
\begin{example}
Multiple entities can be matched at once. Imagine we need to find
mentions of (simple) street addresses in a text consisting of a street
name and a street number. A rule that matches the name and number
separatedly could be:

\begin{multline}
\left(\left\{ \textrm{Orthography type}=\textrm{mixed case word}\right\} \left\{ \textrm{*}\right\} \right)\textrm{: Name}\left(\left\{ \textrm{String}=\textrm{","}\right\} \right)\\
\left(\left\{ \textrm{Orthography type}=\textrm{number}\right\} \right)\textrm{: Number}\\
\longrightarrow\textrm{Street Name}=:\textrm{Name, }\textrm{Street Number}=:\textrm{Number}
\end{multline}

\end{example}

\subsection{Statistical Methods}

Statistical methods aim to decompose the source, assigning a label
to each element in the decomposition. We distinguish between three
types of statistical models:
\begin{itemize}
\item \emph{Token-level models}: they assign a label to each token of the
source. Since entities are usually comprised of multiple adjacent
tokens, the tags used are of the forms ``entity\_begin'', ``entity\_continue'',
``entity\_end'';
\item \emph{Segment-level models}: they try to find the best segmentation
of the source text;
\item \emph{Grammar-based models: }they use formal grammars, outputting
parse trees. All the valid parses are considered for an input document,
assigning a score to each. The parse with the highest score is retained.
\end{itemize}
I now give a brief description of Token-level Models and Segment-level
Models.

\subsubsection{Token-Level Models\label{sub:Token-Level-Models}}

\paragraph{Features}

Clues are of the form 
\begin{equation}
f:\left(x,y,i\right)\longmapsto\mathbb{R}
\end{equation}
where $x$ is a sequence of tokens, $i$ is a position in $x$ and
$y$ is a candidate label for the token at $i$. We distinguish between
these types of features:
\begin{itemize}
\item \emph{word features};
\item \emph{orthographic features;}
\item \emph{dictionary lookup features.}
\end{itemize}

\paragraph{Models for Labeling Token}

The best models are the ones that take into account dependencies between
tokens, among which we may find:
\begin{itemize}
\item \emph{ordered classifiers};
\item \emph{Hidden Markov models;}
\item \emph{Maximum Entropy Taggers;}
\item \emph{Conditional Markov Models;}
\item \emph{Conditional Random Fields }(the state of the art).
\end{itemize}

\subsubsection{Segment-Level Models}

\paragraph{Features}

In these models the label for a segment depends on the properties
of its token and on the previous segment. We can describe a segment
$s_{j}$ as:
\begin{equation}
s_{j}=\left(y_{i},l_{j},u_{j}\right)\label{eq:segment}
\end{equation}
where $y_{i}$ is the proposed label for $s_{j}$ and $l_{j},u_{j}$
are the start and end positions of $s_{j}$. Therefore, a feature
is of the form:
\begin{equation}
f\left(y_{i},y_{i-1},x,l_{j},u_{j}\right)
\end{equation}
where $x$ is the input sequence of tokens, $y_{i-1}$ is the proposed
label for the previous segment and the other symbols are defined as
for Equation \ref{eq:segment}. Besides token-level features (see
subsection \ref{sub:Token-Level-Models}), we can exploit the following
feature types:
\begin{itemize}
\item \emph{Similarity to an entity in a database;}
\item \emph{Length of the segment.}
\end{itemize}
There are also global segmentation models, that try to find the best
segmentation of a token sequence by maximizing a target function.

\section{Relationship Extraction}

When exctracting relationship between entities, we might face three
types of specific tasks:
\begin{itemize}
\item given a pair of entities, find the relationship between them;
\item given an entity $e$ and a relationship $r$, find all the other entities
$e'$ such that $\left(e,e'\right)\in r$;
\item given a big and open-ended input and a relationship $r$ find all
pairs of entities $e',e''$ such that $\left(e',e''\right)\in r$.
\end{itemize}

\subsection{Predicting the Relationship between a Entity Pair }

For the first task, we can exploit the following resources:
\begin{itemize}
\item \emph{surface tokens}: tokens that are usually placed between entities.
They are strong clues;
\item \emph{part of speech tags }(the most important being verbs);
\item \emph{syntactic parse tree}: allows grouping words in phrase types,
e.g. noun phrases, propositional phrases, and so on;
\item \emph{dependency graph}: it is a less expensive structure to compute
than the parse tree and it links a word to those that depend on it.
\end{itemize}
The main methods available to carry out the task are:
\begin{itemize}
\item \emph{Feature-based methods}, that simply transform the clues mentioned
above for usage by conventional classifier models;
\item \emph{Kernel-based methods}, that use kernel functions to encode the
similarity between two graphs;
\item \emph{Rule-based methods}, creating rules over structures around pairs
of entities.
\end{itemize}
The second task is a special case of the third one so I don't treat
it directly.

\subsection{Finding All Possible Entity Pairs Belonging to a Relationship}

The third task can be met especially when dealing with the Web. Usually
we can exploit the following resources to fulfill it:
\begin{itemize}
\item the \emph{types of arguments} of $r$ (that might need specific recognition
patterns);
\item a \emph{seed database} of pairs of entities belonging to $r$;
\item \emph{manually coded patterns}.
\end{itemize}
The generic procedure that is used in this case can be described with
these steps:
\begin{enumerate}
\item Use the seed database to learn the relevant extraction patterns;
\item Use the obtained patterns to define candidate triples of the form
$\left(e',e'',r\right)$;
\item Retain a subset of the candidate triples, using a statistical test.
\end{enumerate}
There exist also rule-based methods for the task.

\chapter{SystemT and AQL}

In this chapter, I give a description of \emph{SystemT} and its extraction
rule language \emph{AQL} (Annotation Query Language), and of a new
way of evaluating AQL queries, that uses a formalism derived from
\emph{Finite State Automata}. Producing and analysing a concrete runtime
system for AQL queries that employs this new method is the object
of my thesis. Since SystemT is a rule-based system, I now focus only
on this category.

\section{SystemT}

As I mentioned in the previous chapter, scalability has now become
a central issue to IE. Companies now rely on IE for many tasks; a
prominent example is \emph{Business Intelligence}. Unfortunately,
many systems developed in the past don't address this concern correctly.
Traditionally, most rule-based system rely on cascading grammars:
formal grammars are executed in sequence on the input, each grammar
representing a stage that takes as input the ouput of the previous
one. Rules in such grammars are matched using regular expressions:
if a part of the input text satisfies the regular expression associated
with a rule, that rule is activated (i.e. ``fires'') and the corresponding
action is executed. Evaluating a grammar on a document tends to be
costly, because simply evaluating a single rule might require scanning
the whole document. On large datasets, the running time becomes enormous.
What's more, these kind of systems are not able to fulfill the expressivity
requirements of the complex extraction tasks.\\
SystemT was developed at IBM to address these issues. This system
is based on a new approach to extraction rules: the \emph{algebraic
approach}. Here, data manipulations are viewed as operations in a
(relational) algebra. Extraction tasks are defined using \emph{annotators}
(see Section \ref{sec:Entity-Extraction}) whose rules are viewed
as \emph{queries} on input documents, that act as virtual databases.
Complex annotators are obtained by combining simple ones by means
of \emph{relational operators}. By doing so, all the optimization
techniques used in regular databases become available, but new techniques
are possible too, due to the characteristics of text documents. Another
advantage of SystemT is that it is capable to handle \emph{overlapping
rule matchings}, due to rulels being fired independently, which are
difficult to resolve in cascading grammars, where they tipically require
disambiguation policies that only partially solve the problem.\\
I now give an overview of SystemT's high-level structure. Then, I
talk about its data model, execution model and algebra.

\subsection{Architecture}

These are the main components of SystemT:
\begin{description}
\item [{Development~Environment:}] allows constructing annotators for
extraction tasks. Rules are expressed in AQL, and compiled into an
algebra. It supports an iterative development process.
\item [{Optimizer:}] seeks for the best query plan for an extraction task,
evaluating the most convenient optimization techniques in a cost-based
manner.
\item [{Runtime~Environment:}] given a query plan, it instantiates the
physical operators corresponding to the logical ones in the plan,
then it proceeds evaluating it on input documents.
\end{description}
Once the development of an annotator is complete, it is \emph{published}
to the optimizer and runtime. After optimization, the runtime evaluates
it on a continuous stream of documents.

\subsection{Data Model\label{sub:Data-Model}}

SystemT uses an \emph{object-relational} data model for annotations
on a text document, that allows applying logical operators over it,
that can be composed. There is an important assumption to mention
before continuing: an extraction task over a set of documents is performed
one document at a time. This means that any relationships between
entities in different documents are disregarded. This assumption is
crucial to some optimization techniques that SystemT uses. While there
exist tasks where considering these relationships would be useful
(think of the Web), still a large number of relevant tasks can be
carried out this way. In the following, I define the basic data types
of SystemT's data model. \\
I consider a finite alphabet $\varSigma$ of symbols (characters).
$\sum^{*}$ denotes the set of all strings of finite length over $\sum^{*}$.
A document is modeled as one such string. 
\begin{definition}
A document is a string $s\in\sum^{*}$.
\end{definition}

\noindent The most basic data type is the \emph{span}.
\begin{definition}
Given a string $s=\sigma_{1}...\sigma_{n}$ where $\forall i$, $\sigma_{i}\in\Sigma$,
with length$\left|s\right|=n$ and whose characters are indexed in
the natural way, a span is a substring $\left[i,j\right\rangle $
, where $i,j$ are indices of $s$ satisfying $1\leq i\leq j\leq n+1$.
A span of $s$ is also denoted as $s_{\left[i,j\right\rangle }$.
\end{definition}

\noindent Spans can be aggregated in \emph{tuples}, which are finite
sequences of spans\emph{. }Let us denote the set of all possible spans
of a string $s$ by $\textrm{Spans}\left(s\right)$ and by $\textrm{SVars}$
an infinite set of span variables, which can be assigned spans. A
$\left(V,s\right)-\textrm{tuple}$ is defined as follows:
\begin{definition}
Given a set $V\subseteq\textrm{SVars}$, $V$ being finite, and a
string $s\in\sum^{*}$, a $\left(V,s\right)-\textrm{tuple}$ is a
mapping $\mu:V\longmapsto\textrm{Spans}\left(s\right)$. When $V$
is clear from the context, we might call $\mu$ simply a $s-\textrm{tuple}$.
\end{definition}

\noindent A set of tuples is called a \emph{relation. }Here, I focus
on \emph{span relations}, which I formally define as $\left(V,s\right)-\textrm{relations}$.
\begin{definition}
\label{span-relation}A $\left(V,s\right)-\textrm{relation}$ is a
set of $\left(V,s\right)-\textrm{tuples}$. As before, we can speak
of $s-\textrm{relation}$ when $V$ is clear.
\end{definition}

\noindent Each operator in the algebra takes one or more span relations
as input and outputs a single span relation.

\subsection{Execution Model\label{sub:Execution-Model}}

A single document is conceived as a \emph{local annotation database},
to which annotators are applied in order to build \emph{views}. In
general, a local database fits into main memory. Local databases are
contained in a \emph{global annotation database}. The runtime of SystemT
takes a global database and it annotates its local databases. The
procedure is described by Algorithm \ref{alg:Annotating-all-local}.

\begin{algorithm}
$E\longleftarrow\left\{ \textrm{algebra expression}\right\} $

for localDB in globalDB do

begin
\begin{enumerate}
\item $\left\{ \textrm{Read localDB into main memory}\right\} $
\item $R\longleftarrow E\left(\textrm{localDB}\right)$
\item $\left\{ \textrm{Add \ensuremath{R} to localDB}\right\} $
\item $\left\{ \textrm{Write modified localDB to disk}\right\} $
\end{enumerate}
\caption{\label{alg:Annotating-all-local}Annotating all local databases in
a global database (taken from REF).}

\end{algorithm}

\noindent In this procedure, step 2 in the for loop is the most expensive.

\subsection{Algebra of Operators \label{sub:Algebra-of-Operators}}

The operators of SystemT's algebra can be classified in three groups:
\begin{itemize}
\item \emph{relational operators;}
\item \emph{span extraction operators;}
\item \emph{span aggregation operators.}
\end{itemize}
In addition, there exist some \emph{span selection predicates} that
are used for span selection.

\subsubsection{Relational Operators}

Relational operators are the usual operators of relational algebra
that appear in classical database query plans. Important examples
are:
\begin{itemize}
\item \emph{selection} ($\sigma$);
\item \emph{projection }($\pi$);
\item \emph{Cartesian product }($\times$);
\item \emph{Union} ($\cup$);
\item \emph{Intersection} ($\cap$).
\end{itemize}

\subsubsection*{Span Extraction Operators}

Loosely speaking, span extraction operators take a pattern and a document
as input and output the maximal set of spans that match that pattern.
There are two main span extraction operators:
\begin{lyxlist}{00.00.0000}
\item [{\emph{standard~regular~expression~matcher~($\varepsilon_{re}$)}:}] this
operator takes a regular expression $r$ as input and it identifies
all non-overlapping matchings of $r$ in the current document, from
left to right.
\item [{\emph{dictionary~matcher~($\varepsilon_{d}$)}:}] this operator
outputs all the spans that match some entry in a given dictionary.
\end{lyxlist}
Although the dictionary matcher might seem useless since there is
a regular expression matcher, it has some advantages over it, as the
fact that it can find overlapping matchings or that it enforces the
semantics of word boundaries.

\subsubsection*{Span Aggregation Operators}

Span aggregation operators are used to aggregate spans in a meaningful
way. They are of two types:
\begin{itemize}
\item \emph{Consolidation}: they are used to coalesce overlapping spans
that were matched using patterns for the same concept. They are:

\begin{lyxlist}{00.00.0000}
\item [{\emph{containment~consolidation~($\varOmega_{c}$)}:}] it discards
a span if it is fully contained into another one.
\item [{\emph{overlap~consolidation~($\varOmega_{o}$)}:}] it merges
spans that overlap repeatedly.
\end{lyxlist}
\item \emph{Block ($\beta$)}: it matches a series of spans, each at a distance
from its neighbor span(s) that is not superior to a threshold. It
is thought to identify regularity. It is tuned by two parameters:
a \emph{distance constraint} to control regularity and a \emph{count
constraint }that establishes the minimum number of spans in a block.
\end{itemize}

\subsubsection{Span Selection Predicates}

Consider two spans $s_{1},s_{2}$. The main span predicates that may
be used for selection are the following:
\begin{lyxlist}{00.00.0000}
\item [{$s_{1}\preceq_{d}s_{2}$}] \noindent when $s_{1},s_{2}$ do not
overlap, $s_{1}$ precedes $s_{2}$ and there are less than $d$ characters
between them;
\item [{$s_{1}\simeq s_{2}$}] when the two spans overlap;
\item [{$s_{1}\subset s_{2}$}] when $s_{1}$ is strictly contained in
$s_{2}$;
\item [{$s_{1}=s_{2}$}] when $s_{1}$ equals $s_{2}$.
\end{lyxlist}

\subsection{Optimization Techniques}

As I mentioned in the introduction of this chapter, SystemT can make
use of the optimization techniques from traditional database systems,
but there exist some peculiar aspects of SystemT and span extraction
tasks that enable new optimization methods.
\begin{remark}
The effect of document-at-a-time processing is that the span relations
produced and consumed for a single document by operators are very
small in size and often empty.\label{small-relations}
\end{remark}

\begin{remark}
Evaluating an annotator on a large set of document is a \emph{CPU-intensive
process}. This is because the running time is by far dominated by
the execution of the operators $\varepsilon_{re}$ and $\varepsilon_{d}$,
that are applied to each document in an input set.
\end{remark}

\begin{remark}
Spans are nothing but \emph{intervals}, so we can exploit \emph{interval
algebra}.\label{interval-algebra}
\end{remark}

\noindent With Remark \ref{small-relations} and Remark \ref{interval-algebra}
in mind, running time can be reduced in a number of ways. I present
them briefly.
\begin{description}
\item [{Regular~Expression~Strength~Reduction}] Some classes of regular
expressions, as defined in the POSIX standard, are executed more efficiently
by using \emph{specialized engines }that are able to improve performance.
For example, an expression that looks for a finite number of strings
in a text is evaluated more efficiently by a string-matching engine.
\item [{Shared~Dictionary~Matching~(SDM)}] Dictionary lookups are usually
very expensive, as we need to consult the dictionary thousand of times
in a typical setting. Instead of evaluating each $\varepsilon_{d}$
operator in an extraction task independently for each input document,
it is evaluated once and for all at the beginning of the process,
and the obtained matches are shared among the single documents.
\item [{Conditional~Evaluation~(CE)}] As we know, an annotator is evaluated
indipendently on each input document. Thus, by employing some heuristic,
it can be guessed if a (sub)query will give any matches in a document
without loss of generality, and in case it does not, it is not evaluated
on it.
\item [{Restricted~Span~Extraction~(RSE)}] It consists in executing
expensive operations on selected regions of a document. It is used
for \emph{join} operators involving dictionary matching operators
and/or a regular expression matching operators in their argments.
One input operator to a join is evaluated on the whole document, while
the other operators are modified to consider only some neighborhood
of the results from the first one. These neighborhoods are established
by some ad hoc heuristic.
\end{description}
The techniques described allow the use of a \emph{cost-based plan
optimizer}: in a query plan, the subgraphs suitable for optimization
are identified, and all the possible optimized plans for them are
formulated. In the end, the best ones are retained. Results of sample
experiments on a large dataset are decribed in REF and REF.

\section{AQL}

AQL is the concrete language used by SystemT to express annotators,
designed to support the execution model I described in Subsection
\ref{sub:Execution-Model}. It is a declarative language with a syntax
very similar to that of SQL. It supports all the operators that I
presented in Subsection \ref{sub:Algebra-of-Operators}. When coding
an extraction task in AQL, we are able to build a series of views
of a document, that are of progressively higher level, and can be
based on lower-level views. The content of a view corresponds with
annotations in the text. An input document is modeled as a view too,
which is provided by default. The main advantages of AQL are:
\begin{itemize}
\item it allows formulating \emph{complex low-level }patterns in a declarative
fashion;
\item it enables \emph{modularization }and \emph{reuse} of the queries,
making development and maintance of \emph{complex high-level structures}
easier.
\end{itemize}

\section{A Formalism for AQL's Core: VStack and VSet Automata}

AUTHORS describe in REF a formal model that captures the core functionality
of AQL. They show a way to represent annotators expressed in AQL by
means of modified finite state automata, namely \emph{VStack} and
\emph{VSet automata}. In my work, I define and implement a runtime
system based on this new formalism. I now describe the model and discuss
the relative expressive power of its main elements. First of all,
we need to be more precise on some basic concepts.

\subsection{Basic Definitions}

I consider $\varSigma$ and $\varSigma^{\ast}$, defined as in Subsection
\ref{sub:Data-Model}. Let us start with the definition of a language.
\begin{definition}
A language L is a subset of $\varSigma^{\ast}$.
\end{definition}

\noindent Now we can state formally define regular expressions, by
giving their language.
\begin{definition}
Regular expressions over $\varSigma$ are those that belong to the
language $\gamma$ defined as:
\begin{equation}
\gamma:=\textrm{Ø}\mid\epsilon\mid\sigma\mid\gamma\vee\gamma\mid\gamma\cdot\gamma\mid\gamma^{\ast}
\end{equation}

\noindent where:
\begin{itemize}
\item $\slashed{O}$ is the empty language;
\item $\epsilon$ is the empty string;
\item $\sigma\in\varSigma$;
\item $\vee$ is the ordinary disjunction operator;
\item $\cdot$ is the concatenation operator;
\item $\ast$ is the Kleene Closure.
\end{itemize}
\end{definition}

\noindent Additionally, we might use $\gamma^{+}$ as a shortcut for
$\gamma\cdot\gamma^{\ast}$, and $\varSigma$ as an abbreviation of
$\sigma_{1}\vee...\vee\sigma_{n}$ (with a slight abuse of notation).
We also need the definition of a \emph{string relation}.
\begin{definition}
A n-ary string relation is a subset of $\left(\varSigma^{\ast}\right)^{n}$.
\end{definition}

\noindent I have already defined span relations (Definition \ref{span-relation}).
They allow us to give a precise definition of annotators, by introducing
the concept of \emph{document spanner}.
\begin{definition}
Given a string $s$, a document spanner P is a function that maps
$s$ to a $\left(V,s\right)-\textrm{relation}$ $r$ , where $V:=\textrm{Svars}\left(P\right)$.
We have $r=P\left(s\right).$ We say $P$ is n-ary if $\left|V\right|=n$.
\end{definition}

\noindent I now present two types of spanners that will be useful
later on. The first one is that of \emph{Hierarchical Spanners}. First,
let us see when a $s-\textrm{tuple}$ is hierarchical.
\begin{definition}
\label{hierarchical-tuple}Given a string $s$, a document spanner
P and an $s-\textrm{tuple}$ $\mu\in P\left(s\right)$, $\mu$is hiearchical
if, for every $x,y\in\textrm{SVars}\left(P\right)$, one of the following
conditions holds:
\begin{itemize}
\item $\mu\left(x\right)\supseteq\mu\left(y\right);$
\item $\mu\left(x\right)\subseteq\mu\left(y\right)$;
\item $\mu\left(x\right)$ is disjoint from $\mu\left(y\right)$.
\end{itemize}
\end{definition}

\noindent The definition of \emph{hierarchical spanner} follows.
\begin{definition}
\label{hierarchical-spanner}Given a document spanner $P$, $P$ is
hierarchical if $\forall s\in\varSigma^{\ast}$, $\forall\mu\in P\left(s\right)$,
$\mu$ is hierarchical.
\end{definition}

\noindent The class of hierarchical spanners is denoted by $\mathbf{HS}$.
As it emerges from Definition \ref{hierarchical-tuple} and Definition
\ref{hierarchical-spanner}, a hierarchical spanner outputs only $s-\textrm{tuples}$
whose spans don't overlap. The second type of spanners I introduce
is that of \emph{Universal Spanners}. Some accessory definitions are
needed.
\begin{definition}
Given a string $s$ and a document spanner $P$, $P$ is total on
$s$ if $P\left(s\right)$ consists of all the possible $s-\textrm{tuples}$
on $\textrm{SVars}\left(P\right)$.
\end{definition}

\begin{definition}
Given a string $s$ and a document spanner $P$, $P$ is hierarchically
total on $s$ if $P\left(s\right)$ consists of all the possible hierarchical
$s-\textrm{tuples}$ on $\textrm{SVars}\left(P\right)$.
\end{definition}

\noindent There are two kinds of universal spanners: the \emph{universal
spanner} and the \emph{universal hierarchical spanner}.
\begin{definition}
Given a set of span variables $Y$$\subseteq\textrm{SVars}$, the
universal spanner $\Upsilon_{Y}$ over $Y$ is the unique document
spanner $P$ such that $\textrm{SVars}\left(P\right)=Y$ and $P$
is total on every string $s\in\varSigma^{\ast}$. 
\end{definition}

\begin{definition}
Given a set of span variables $Y$$\subseteq\textrm{SVars}$, the
universal hierarchical spanner $\Upsilon_{Y}^{H}$ over $Y$ is the
unique document spanner $P$ such that $\textrm{SVars}\left(P\right)=Y$
and $P$ is hierarchically total on every string $s\in\varSigma^{\ast}$. 
\end{definition}

\noindent I now present the formal spanner represantations described
in REF.

\subsection{Spanner Representations}

We saw in Subsection \ref{sub:Algebra-of-Operators} that the operations
used by SystemT and AQL to extract spans relations are regular expression
matching and dictionary matching. Here I focus on regular expressions.
AQL's regular expressions can be seen as usual regular expressions
enriched with \emph{capture variables}, that are nothing but the span
variables that constitute the span relations I defined. In what follows.
In REF, three formalisms are described that are able to model this
kind of modified regular expressions, namely:
\begin{itemize}
\item \emph{Regex Formulas;}
\item \emph{Variable Stack Automata;}
\item \emph{Variable Set Automata.}
\end{itemize}
These formalisms are also called \emph{primitive spanner representations}.

\subsubsection{Regex Formulas}

To define regex formula, I first introduce the concept of \emph{variable
regex.}
\begin{definition}
A variable regex is a regular expression with capture variables that
extends usual regular expressions in the following way:
\begin{equation}
\gamma:=\textrm{Ø}\mid\epsilon\mid\sigma\mid\gamma\vee\gamma\mid\gamma\cdot\gamma\mid\gamma^{\ast}\mid x\left\{ \gamma\right\} 
\end{equation}
 where:
\begin{itemize}
\item $x\in\textrm{SVars}$;
\item $x\left\{ \gamma\right\} $ means that we assign the result of the
evaluation of $\gamma$ to x.
\end{itemize}
\end{definition}

\noindent Evaluating a variable regex on a string produces a parse
tree over the alphabet $\Lambda=\varSigma\cup\textrm{SVars}\cup\left\{ \in,\vee,\cdot,\ast\right\} $.
A valid parse tree for a variable regex$\gamma$ is called a $\gamma-\textrm{parse}$.
We accept only those variable regex expressions whose parses have
only one occurence of each of their variables, otherwise the variable
assignment would remain unclear. Such expressions are referred to
as \emph{functional}.
\begin{definition}
A variable regex $\gamma$ is functional if $\forall s\in\varSigma^{\ast}$,
$\forall\gamma-\textrm{parse }t$ for $s$, $\forall x\in\textrm{SVars}\left(\gamma\right)$,
$x$ occurs exactly one time in $t$.
\end{definition}

\begin{definition}
A regex formula is a functional variable regex.
\end{definition}

\noindent The spanner represented by a regex formula $\gamma$ may
be denoted as $\left\llbracket \gamma\right\rrbracket $. We have
that $\textrm{SVars}\left(\left\llbracket \gamma\right\rrbracket \right)=\textrm{SVars}\left(\gamma\right)$,
and the span relation $\left\llbracket \gamma\right\rrbracket \left(s\right)$
is the set $\left\{ \mu^{p}\mid p\textrm{ is a }\gamma-\textrm{parse for \ensuremath{s}}\right\} $,
where $\mu^{p}$ is a tuple defined by a parse $p$.

\subsubsection{Variable Stack Automata}

Variable stack automata (VStack Automata or vstk-automata for short)
are representations of document spanners by means of modified finite
state automata (FAs). Basically, a FA is augmented with a \emph{stack
of span variables}. A variable is pushed on the stack when its corresponding
span is opened, and popped when it is closed. The formal definition
of a vstk-automaton follows.
\begin{definition}
\label{vstk-aut}A Variable Stack Automaton is a tuple $\left(Q,q_{0},q_{f},\delta\right)$,
where:
\begin{itemize}
\item $Q$ is a finite set of states;
\item $q_{0}\in Q$ is the initial state;
\item $q_{f}\in Q$ is the accepting state;
\item $\delta$ is a finite transition relation, containing triples of the
forms $\left(q,\sigma,q'\right)$, $\left(q,\epsilon,q'\right)$,
$\left(q,x\vdash,q'\right)$, $\left(q,\dashv,q'\right)$, where:

\begin{itemize}
\item $q,q'\in Q$;
\item $\sigma\in\varSigma$;
\item $x\in SVars$;
\item $\vdash$ is the push symbol;
\item $\dashv$ is the pop symbol.
\end{itemize}
\end{itemize}
\end{definition}

\noindent Notice that we don't need to specify which variable we want
to pop, as it is naturally the last that was pushed on the stack.
Given a vstk-automaton $A$, the set of variables that appear in its
transitions is denoted as $\textrm{SVars}\left(A\right)$. Next I
give the definitions of a \emph{configuration }and of a \emph{run}
of a vstk-automaton.
\begin{definition}
\label{vstk-conf}Given a string $s$ with length $\left|s\right|=n$
and a vstk-automaton $A$ , a configuration of $A$ is a tuple $c=\left(q,\overrightarrow{v},Y,i\right)$,
where:
\begin{itemize}
\item $q\in Q$ is the current state;
\item $\overrightarrow{v}$ is the current variable stack;
\item $Y\in Vars$$\left(A\right)$ is the set of available variables (those
not already pushed on the stack);
\item $i\in\left\{ 1,...,n+1\right\} $ is the position of the next character
to be read in $s$.
\end{itemize}
\end{definition}

\noindent Here, as for regex formulas, we want the variable assignment
to be clear, so once a variable is pushed on the stack, it is removed
from the set of available variables, thus it can be pushed only once.
For the rest, a run of a vstk-automaton is similar to those of ordinary
FAs.
\begin{definition}
Given a string $s$ and a vstk-automaton $A$, a run $\rho$ of $A$
on $s$ is a sequence of configurations $c_{0},...,c_{m}$ such that:
\begin{itemize}
\item $c_{0}=\left(q_{0},\epsilon,\textrm{SVars}\left(A\right),1\right)$;
\item $\forall j\in\left\{ 0,...,m-1\right\} $, for $c_{j}$=$\left(q_{j},\overrightarrow{v_{j}},Y_{j},i_{j}\right)$,
$c_{j+1}$=$\left(q_{j+1},\overrightarrow{v_{j+1}},Y_{j+1},i_{j+1}\right)$
one of the following holds:

\begin{itemize}
\item $\overrightarrow{v_{j+1}}=\overrightarrow{v_{j}}$, $Y_{j+1}=Y_{j}$
and either:

\begin{itemize}
\item $i_{j+1}=i_{j}+1$, $\left(q_{j},s_{i_{j}},q_{j+1}\right)\in\delta$;
\item $i_{j+1}=i_{j}$, $\left(q_{j},\in,q_{j+1}\right)\in\delta$.
\end{itemize}
\item $i_{j+1}=i_{j}$ and for some $x\in\textrm{SVars}\left(A\right)$
either:

\begin{itemize}
\item $\overrightarrow{v_{j+1}}=\overrightarrow{v_{j}}\cdot x$, $x\in Y_{j}$,
$Y_{j+1}=Y_{j}\setminus\left\{ x\right\} $, $\left(q_{j},x\vdash,q_{j+1}\right)\in\delta$
(x is pushed on the stack);
\item $\overrightarrow{v_{j}}=\overrightarrow{v_{j+1}}\cdot x$, $x\in Y_{j}$,
$Y_{j+1}=Y_{j}$, $\left(q_{j},\dashv,q_{j+1}\right)\in\delta$ (a
variable is popped from the stack).
\end{itemize}
\end{itemize}
\end{itemize}
\end{definition}

\begin{definition}
Given a string with length $\left|s\right|=n$ and a vstk-automaton
$A$, a run $\rho=c_{0},...,c_{m}$ of $A$ on $s$ is accepting if
$c_{m}=\left(q_{f},\epsilon,\textrm{Ø},n+1\right)$.
\end{definition}

\noindent The set of all possible accepting runs of a vstk-automaton
$A$ on a string $s$ is denoted as $\textrm{ARuns}\left(A,s\right)$.
The spanner represented by $A$ may be referred to as $\left\llbracket A\right\rrbracket $.
We have that $\textrm{SVars}\left(\left\llbracket A\right\rrbracket \right)=\textrm{SVars}\left(A\right)$,
and the span relation $\left\llbracket A\right\rrbracket \left(s\right)$
is the set $\left\{ \mu^{\rho}\mid\rho\in\textrm{ARuns}\left(A,s\right)\right\} $,
where $\mu^{\rho}$ is a tuple defined by a run $\rho$. In particular,
for every variable $x\in\textrm{SVars}\left(A\right)$, $\mu^{\rho}\left(x\right)$
is the span $\left[i_{b},i_{e}\right\rangle $, where:
\begin{itemize}
\item $c_{b}$=$\left(q_{b},\overrightarrow{v_{b}},Y_{b},i_{b}\right)$
is the unique configuration of $\rho$ where $x$ appears in the stack
for the fist time;
\item $c_{e}$=$\left(q_{e},\overrightarrow{v_{e}},Y_{e},i_{e}\right)$
is the unique configuration of $\rho$ where $x$ appears in the stack
for the last time.
\end{itemize}

\subsubsection{Variable Set Automata}

Variable Set Automata (VSet Automata or vset-automata for short) are
another formalism for representing document spanners that is based
on FAs. Vset-automata are defined in a very similar way to that of
vstk-automata. The main differences are:
\begin{itemize}
\item the stack of span variables is replaced by a \emph{set;}
\item since no order is defined on the variables in the set, when we close
a variable, we need to specify which one.
\end{itemize}
I now give a formal definition of a vset-automaton.
\begin{definition}
A variable set automaton is a tuple $\left(Q,q_{0},q_{f},\delta\right)$,
where:
\begin{itemize}
\item $Q$, $q_{0}$ and $q_{f}$ are defined as in Definition \ref{vstk-aut};
\item $\delta$ is the same as in Definition \ref{vstk-aut}, except that
it has triples of the form $\left(q,\dashv x,q'\right)$, with $x\in\textrm{SVars}$,
instead of those of the form $\left(q,\dashv,q'\right)$.
\end{itemize}
\end{definition}

\noindent We also need to slightly modify the definitions of configuration
and run with respect to those of a vstk-automaton.
\begin{definition}
Given a string $s$ with length $\left|s\right|=n$ and a vset-automaton
$A$ , a configuration of $A$ is a tuple $c=\left(q,V,Y,i\right)$,
where:
\begin{itemize}
\item $q$, $i$ are defined as in Definition \ref{vstk-conf};
\item $V\subseteq\textrm{Vars}\left(A\right)$ is the active variable set;
\item $Y\in Vars$$\left(A\right)$ is the set of available variables (those
not already placed in the set).
\end{itemize}
\end{definition}

\begin{definition}
Given a string $s$ and a vset-automaton $A$, a run $\rho$ of $A$
on $s$ is a sequence of configurations $c_{0},...,c_{m}$ such that:
\begin{itemize}
\item $c_{0}=\left(q_{0},\textrm{Ø},\textrm{SVars}\left(A\right),1\right)$;
\item $\forall j\in\left\{ 0,...,m-1\right\} $, for $c_{j}$=$\left(q_{j},V_{j},Y_{j},i_{j}\right)$,
$c_{j+1}$=$\left(q_{j+1},V_{j+1},Y_{j+1},i_{j+1}\right)$ one of
the following holds:

\begin{itemize}
\item $V_{j+1}=V_{j}$, $Y_{j+1}=Y_{j}$ and either:

\begin{itemize}
\item $i_{j+1}=i_{j}+1$, $\left(q_{j},s_{i_{j}},q_{j+1}\right)\in\delta$;
\item $i_{j+1}=i_{j}$, $\left(q_{j},\in,q_{j+1}\right)\in\delta$.
\end{itemize}
\item $i_{j+1}=i_{j}$ and for some $x\in\textrm{SVars}\left(A\right)$
either:

\begin{itemize}
\item $x\in Y_{j}$, $V_{j+1}=V_{j}\cup\left\{ x\right\} $, $Y_{j+1}=Y_{j}\setminus\left\{ x\right\} $,
$\left(q_{j},x\vdash,q_{j+1}\right)\in\delta$ ($x$ is inserted into
the set);
\item $x\in V_{j}$, $V_{j+1}=V_{j}\setminus\left\{ x\right\} $, $Y_{j+1}=Y_{j}$,
$\left(q_{j},\dashv x,q_{j+1}\right)\in\delta$ ($x$ is removed from
the set).
\end{itemize}
\end{itemize}
\end{itemize}
\end{definition}

\begin{definition}
Given a string with length $\left|s\right|=n$ and a vset-automaton
$A$, a run $\rho=c_{0},...,c_{m}$ of $A$ on $s$ is accepting if
$c_{m}=\left(q_{f},\textrm{Ø},\textrm{Ø},n+1\right)$.
\end{definition}

\noindent Given a string $s$ and a vset-automaton $A$, $\textrm{ARuns}\left(A,s\right)$and
$\left\llbracket A\right\rrbracket $ are defined as for vstk-automata.

\subsection{Algebras of Spanners}

Besides mere span extraction, AQL offers the capability to combine,
transform and filter extracted tuples by using a series of relational
operators. Here, I introduce the relational operators for spanners,
described in REF, that are considered to capture the core relational
operations of AQL. They are:
\begin{itemize}
\item \emph{Union }( $\cup$);
\item \emph{Projection }($\pi$);
\item \emph{Natural Join} ($\bowtie$);
\item \emph{String Selection} ($\varsigma$).
\end{itemize}
A finite set of spanner operators forms a \emph{spanner algebra}.
In the following, give definitions for each of the listed spanner
operators.

\subsubsection{Union}

\subsubsection{Projection}

\subsubsection{Natural Join}

\subsubsection{String Selection}

\include{conclusion}

\selectlanguage{english}%
\bibliographystyle{apalike}
\bibliography{xampl}

\selectlanguage{american}%
\appendix
\include{appendix}

\printindex{} 
\end{document}

### 2016-03-07 paper/paper.tex
Almost complete stub of Chapter 3

@@
 
-Consider two spans $s_{1},s_{2}$. The main span predicates that may
-be used for selection are the following:
+\noindent Consider two spans $s_{1},s_{2}$. The main span predicates
+that may be used for selection are the following:
 \begin{lyxlist}{00.00.0000}
@@
 $\sigma_{1}\vee...\vee\sigma_{n}$ (with a slight abuse of notation).
-We also need the definition of a \emph{string relation}.
+The language $\mathcal{L}\left(\gamma\right)$of a regular expression
+$\gamma$ is the set of strings $\mathbf{s}$ over $\varSigma$ that
+are matched by that expression. A language $L$ is \emph{regular}
+if $L=\mathcal{L}\left(\gamma\right)$ for some regular expression
+$\gamma$. We also need the definition of a \emph{string relation}.
 \begin{definition}
 A n-ary string relation is a subset of $\left(\varSigma^{\ast}\right)^{n}$.
+\end{definition}
+
+\noindent An interesting class of string relations is that of recognizable
+relations, denoted as $\textrm{REC}$.
+\begin{definition}
+Given a k-ary string relation $R$, $R$ is recognizable if it is
+representable by a finite union the form:
+\begin{equation}
+\bigcup L_{1}\times...\times L_{k}
+\end{equation}
+
+where each $L$$_{i}$ is a regular language.
 \end{definition}
@@
 is the set $\left\{ \mu^{p}\mid p\textrm{ is a }\gamma-\textrm{parse for \ensuremath{s}}\right\} $,
-where $\mu^{p}$ is a tuple defined by a parse $p$.
+where $\mu^{p}$ is a tuple defined by a parse $p$. The class of
+regex formulas is referred to as $\textrm{RGX}$.
 
@@
 \end{itemize}
+\noindent The class of variable stack automata is called $\textrm{VA}_{\textrm{stk}}$.
 
@@
 $\left\llbracket A\right\rrbracket $ are defined as for vstk-automata.
-
-\subsection{Algebras of Spanners}
+The class of variable set automata is called $\textrm{VA}_{\textrm{stk}}$.
+
+\subsection{Algebras of Spanners\label{sub:Algebras-of-Spanners}}
 
@@
 
+Before giving the definition of the union operator, we need to introduce
+the concept of \emph{union compatible spanners.}
+\begin{definition}
+Given two document spanner $P_{1}$ and $P_{2}$, they are union compatible
+if and only if $\textrm{SVars}\left(P_{1}\right)=\textrm{SVars}\left(P_{2}\right)$.
+\end{definition}
+
+\noindent The definition of union of two spanners is as follows.
+\begin{definition}
+Given two union compatible document spanners $P_{1}$ and $P_{2}$,
+their union is the spanner denoted as $P_{1}\cup P_{2}$, for which
+we have that:
+\begin{itemize}
+\item $\textrm{SVars}\left(P_{1}\cup P_{2}\right)=\textrm{SVars}\left(P_{1}\right)$;
+\item given a string $\mathbf{s}$, $\left(P_{1}\cup P_{2}\right)\left(\mathbf{s}\right)=P_{1}\left(\mathbf{s}\right)\cup P_{2}\left(\mathbf{s}\right)$.
+\end{itemize}
+\end{definition}
+
 \subsubsection{Projection}
+\begin{definition}
+Given a document spanner $P$ and a set of span variables $Y\subseteq\textrm{SVars}\left(P\right)$,
+the projection of $P$ over $Y$ is the spanner denoted as $\pi_{Y}P$,
+satisfying $\textrm{SVars}\left(\pi_{Y}P\right)=Y$.
+\end{definition}
+
+\noindent As the name and the definition suggest, the projection of
+a document spanner $P$ over a subset of its span variables $Y$ is
+obtained by reducing the domain of each of the spanner's output $s-\textrm{tuples}$
+to $Y$.
 
 \subsubsection{Natural Join}
+\begin{definition}
+Given two document spanners $P_{1}$ and $P_{2}$, their natural join
+is the spanner denoted as $P_{1}\bowtie P_{2}$, for which we have
+that:
+\begin{itemize}
+\item $\textrm{SVars}\left(P_{1}\bowtie P_{2}\right)=\textrm{SVars}\left(P_{1}\right)\cup\textrm{SVars}\left(P_{2}\right)$;
+\item given a string $\mathbf{s}$, $\left(P_{1}\bowtie P_{2}\right)\left(\mathbf{s}\right)$
+consists of all the $\mathbf{s}-\textrm{tuples}$ $\mu$ agreeing
+with some $\mu_{1}\in P_{1}\left(\mathbf{s}\right)$ and some $\mu_{2}\in P_{2}\left(\mathbf{s}\right)$.
+Note that this implies that $\mu_{1},\mu_{2}$ agree on variables
+that are common to $P_{1},P_{2}$: $\forall x\in\textrm{SVars}\left(P_{1}\right)\cap\textrm{SVars}\left(P_{2}\right)$,
+$\mu_{1}\left(x\right)=\mu_{2}\left(x\right)$. 
+\end{itemize}
+\end{definition}
 
 \subsubsection{String Selection}
+\begin{definition}
+Given a document spanner $P$ and a k-ary string relation $R$, the
+string selection operation according to $R$ is denoted as $\varsigma^{R}$,
+and is parametrized by $x_{1},...,x_{k}\in\textrm{SVars}\left(P\right)$.
+We have that, given a string $\mathbf{s}$ and a spanner $P':=\varsigma_{x_{1},...,x_{k}}^{R}P$,
+$P'\left(\mathbf{s}\right)$ consists of all the $\mathbf{s}-\textrm{tuples}$
+$\mu$ such that $\left(\mathbf{s}_{\mu\left(x_{1}\right)},...,\mathbf{s}_{\mu\left(x_{k}\right)}\right)\in R$.
+\end{definition}
+
+\noindent In the remainder of this document, the only string selection
+operator I consider is $\varsigma_{x,y}^{=}$ which, given a spanner
+$P$, restricts $P\left(\mathbf{s}\right)$ to those $\mathbf{s}-\textrm{tuples}$
+that satisfy $\mathbf{s}_{\mu\left(x\right)}=\mathbf{s}_{\mu\left(y\right)}$.
+
+\medskip{}
+
+\noindent I now introduce some additional notation. Given a generic
+class of spanner representations $\textrm{SR}$, the set of all the
+spannners that can be represented by $\textrm{SR}$ is denoted as
+$\left\llbracket \textrm{SR}\right\rrbracket $. Formally, we have
+that $\left\llbracket \textrm{SR}\right\rrbracket =\left\{ \left\llbracket r\right\rrbracket \mid\in\textrm{SR}\right\} $.
+Let $O$ be a spanner algebra. I denote by $\textrm{SR}^{O}$ the
+closure of $\textrm{SR}$ under $O$, that is: the class of spanner
+represantations obtained by applying (compositions of) operators contained
+in $O$ to the represantations in $\textrm{SR}$. The corresponding
+set of spanners is referred to as $\left\llbracket \textrm{SR}^{O}\right\rrbracket $.
+We are ready to start reasoning on the expressive power of the spanner
+representations that we defined.
+\begin{proposition}
+The following hold:
+\begin{enumerate}
+\item every document spanner represented in the classes $\textrm{RGX}$
+and $\textrm{VA}_{\textrm{stk}}$ is hierarchical, that is: $\left\llbracket \textrm{\textrm{RGX}}\right\rrbracket ,\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}\right\rrbracket \subseteq\mathbf{HS}$;
+\item there exist some spanners represented in $\textrm{VA}_{\textrm{set}}$
+that are not hierarchical: $\left\llbracket \textrm{\ensuremath{\textrm{VA}_{set}}}\right\rrbracket \nsubseteq\mathbf{HS}$;
+\item the operators $\cup$, $\pi$, $\varsigma^{R}$ preserve the property
+of being hierarchical, while $\bowtie$ does not, thus we have that:
+
+\begin{enumerate}
+\item given a class of spanner representations $\textrm{SR}$, $\left\llbracket \textrm{SR}\right\rrbracket \subseteq\mathbf{HS}\Rightarrow\left\llbracket \textrm{SR}^{\left\{ \cup,\pi,\varsigma^{R}\right\} }\right\rrbracket \subseteq\mathbf{HS};$
+\item there exist two hierarchical spanners $P_{1},P_{2}$ such that $P_{1}\bowtie P_{2}\notin\mathbf{HS}$.
+\end{enumerate}
+\end{enumerate}
+\end{proposition}
+
+\noindent In the next Subsection, I discuss the class of \emph{regular
+spanners}, which plays a central role in the discussion of the class
+of spanners that constitute the core of AQL.
+
+\subsection{Regular Spanners}
+
+A regular spanner is defined as follows.
+\begin{definition}
+A spanner is regular if it can be defined by a vset-automaton.
+\end{definition}
+
+\noindent Let us now see how regular spanners are related to the other
+basic spanner classes. A preliminary result is that regex formulas
+and vstk-automata have the same expressive power.
+\begin{theorem}
+\label{.rgx=00003Dvstk}$\left\llbracket \textrm{\textrm{RGX}}\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}\right\rrbracket $.
+\end{theorem}
+
+\noindent It turns out that the spanners expressed by representations
+in $\textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}$($\textrm{\textrm{RGX}}$)
+are exactly those that are both regular and hierarchical.
+\begin{theorem}
+$\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}\right\rrbracket \cap\mathbf{HS}$.
+\end{theorem}
+
+\noindent For what concerns the algebraic operators I presented in
+Subsection \ref{sub:Algebras-of-Spanners}, it can be shown that union,
+projection and join don't increase the expressive power of regular
+spanners.
+\begin{theorem}
+\label{.setext=00003Dset}$\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}^{\left\{ \cup,\pi,\bowtie\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}\right\rrbracket $.
+\end{theorem}
+
+\noindent On the other hand, the same operators make the spanners
+represented in $\textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}$
+equivalent to regular spanners.
+\begin{theorem}
+\label{.stkext=00003Dset}$\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}^{\left\{ \cup,\pi,\bowtie\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}\right\rrbracket $.
+\end{theorem}
+
+\noindent Finally, I look at which string relations can be simulated
+by regular spanners. Let us start by defining the concept of \emph{selectable
+string relation}.
+\begin{definition}
+Given a string relation $R$ and a class of spanners $C$, $R$ is
+selectable by $C$ if for every document spanner $P\in C$ and for
+every $\overrightarrow{x}=x_{1},...,x_{k}$, $x_{i}\in\textrm{SVars}\left(P\right)$,
+we have that $\varsigma_{\overrightarrow{x}}^{R}P\in C$.
+\end{definition}
+
+\noindent Next, I introduce the concept of \emph{restricted universal
+spanner.}
+\begin{definition}
+Given a string relation $R$ and a sequence of variables$\overrightarrow{x}=x_{1},...,x_{k}$
+with their corrsesponding set $X=\left\{ x_{1},...,x_{k}\right\} $,
+the $R-\textrm{restricted}$ universal spanner over $\overrightarrow{x}$
+is $\Upsilon_{\overrightarrow{x}}^{R}:=\varsigma_{\overrightarrow{x}}^{R}\Upsilon_{X}$.
+\end{definition}
+
+\noindent Selectability of a string relation $R$ by a class of spanners
+$C$ corresponds to the presence in $C$ of all the possible $R-\textrm{restricted}$
+universal spanners, under some conditions.
+\begin{proposition}
+Given a string relation $R$ and a class of spanners $C$ containing
+all the possible universal spanners and closed under natural join,
+$R$ is selectable by $C$ if and only if, for every $\overrightarrow{x}=x_{1},...,x_{k}\in\textrm{SVars}^{k}$,
+$\Upsilon_{\overrightarrow{x}}^{R}\in C$.
+\end{proposition}
+
+\noindent In the case of regular spanners, the class of string relations
+that they can select is exactly $\textrm{REC}$.
+\begin{theorem}
+The class of string relations selectable by $\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}\right\rrbracket $
+is $\textrm{REC}$.
+\end{theorem}
+
+\noindent The relation ``$=$'' is not in $\textrm{REC}$, thus
+it is not selectable by regular spanners, nonetheless it is important
+for selection predicates in AQL, which is why regular spanners are
+not enough to model its core. In the following Subsection, by using
+the notions and results presented so far, I describe the spanners
+that are identified in REF as the core of AQL: \emph{core spanners}.
+
+\subsection{Core Spanners}
+
+An expression in the core of AQL belongs to $\textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }$.
+Consequently, a core spanner is defined as follows.
+\begin{definition}
+A core spanner is a document spanner belonging to $\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket $.
+\end{definition}
+
+\noindent Thanks to Theorems \ref{.rgx=00003Dvstk}, \ref{.setext=00003Dset}
+and \ref{.stkext=00003Dset} we can easily see that $\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket $.
+This shows that core spanners can be reduced to regular spanners extended
+with the algebra $\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} $.
+But the following lemma tells us an algebra with fewer operators is
+also sufficient.
+\begin{lemma}
+$\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}^{\left\{ \pi,\varsigma^{=}\right\} }\right\rrbracket $.
+\end{lemma}
+
+\noindent Another lemma, known as the \emph{core simplification lemma},
+gives a even more simple way of representing core spanners.
+\begin{lemma}
+$\mathbf{(Core\,Simplification\,Lemma)}$ Every core spanner can be
+defined by an expression of the form
+
+\begin{equation}
+\pi_{V}SA
+\end{equation}
+
+\noindent where:
+\begin{itemize}
+\item $A$ is a vset-automaton;
+\item $V\subseteq\textrm{SVars}\left(A\right)$;
+\item $S$ is a sequence of string selections $\varsigma_{x,y}^{=}$, for
+$x,y\in\textrm{SVars}\left(A\right)$.
+\end{itemize}
+\end{lemma}
+
+\noindent Now I look at which string relations can be simulated by
+core spanners. In the next definition, I define three string relations
+of relevance.
+\begin{definition}
+Given two strings $\mathbf{s},\mathbf{t}\in\varSigma^{\ast}$:
+\begin{itemize}
+\item $\mathbf{s}\sqsubseteq\mathbf{t}$ if $\mathbf{s}$ is a (consecutive)
+substring of t (i.e. $\mathbf{s}=\mathbf{t}_{\left[i,j\right\rangle }$);
+\item $\mathbf{s}\sqsubseteq_{\textrm{prf}}\mathbf{t}$ if $\mathbf{s}$
+is a prefix of t (i.e. $\mathbf{s}=\mathbf{t}_{\left[1,j\right\rangle }$);
+\item $\mathbf{s}\sqsubseteq_{\textrm{sfx}}\mathbf{t}$if $\mathbf{s}$
+is a suffix of t (i.e. $\mathbf{s}=\mathbf{t}_{\left[1,\left|\mathbf{t}\right|+1\right\rangle }$).
+\end{itemize}
+\end{definition}
+
+\begin{proposition}
+\noindent All the string relations in $\textrm{REC}$, $\mathbf{\sqsubseteq}$,
+$\sqsubseteq_{\textrm{prf}}$ and $\sqsubseteq_{\textrm{sfx}}$ are
+selectable by the core spanners.
+\end{proposition}
+
+\subsection{Difference}
+
+The difference operator is defined as follows.
+\begin{definition}
+Given two union compatible document spanners $P_{1}$ and $P_{2}$,
+their difference is the spanner $P_{1}\setminus P_{2}$, for which
+we have:
+\begin{itemize}
+\item $\textrm{SVars}\left(P_{1}\setminus P_{2}\right)=\textrm{SVars}\left(P_{1}\right)$;
+\item given a string $\mathbf{s}$, $\left(P_{1}\setminus P_{2}\right)\left(\mathbf{s}\right)=\left(P_{1}\right)\setminus\left(P_{2}\right)\left(\mathbf{s}\right)$.
+\end{itemize}
+\end{definition}
+
+\noindent It can be shown that regular spanners are closed under difference.
+\begin{theorem}
+\label{.setdiff=00003Dset}$\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}^{\left\{ \setminus\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}\right\rrbracket $.
+\end{theorem}
+
+\noindent Despite the result of of Theorem \ref{.setdiff=00003Dset},
+core spanners are not closed under difference, which is why this operator
+has not been considered so far.
+\begin{theorem}
+$\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket \subsetneq\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=},\setminus\right\} }\right\rrbracket $.
+\end{theorem}
 

### 2016-04-16 project/src/compiler/VSetAFileReader.scala
Projection. Lexicographic vset-automaton (neeeds debugging). Report of the meeting of 8/4/2016.

@@ -43,7 +43,10 @@ object VSetAFileReader {
         if(!transitionFunction.contains(state1.toInt))
          transitionFunction = transitionFunction + (state1.toInt -> Map[Int, String]())
           
-        transitionFunction(state1.toInt) = transitionFunction(state1.toInt) + (state2.toInt -> label)
+        if(transitionFunction(state1.toInt).contains(state2.toInt))
+          transitionFunction(state1.toInt)(state2.toInt) += "|" + label 
+        else
+          transitionFunction(state1.toInt) = transitionFunction(state1.toInt) + (state2.toInt -> label)
         
       }
       

### 2016-04-16 project/src/compiler/VSetAutomaton.scala

@@ -2,380 +2,640 @@ package compiler
 
 import scala.collection.mutable.Map
 import scala.collection.mutable.ArrayBuffer
+import dk.brics.automaton.Automaton
+import dk.brics.automaton.RegExp
 
 /**
  * Represents a vset-automaton.
-*/
+ */
 
 class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:Map[Int, Map[Int, String]], val vars: Array[String], val finalStates: Array[Int]) {
-  
-/*  def project(vars: Array[Int]): VSetAutomaton {}
-  
+
+	/* 
+
   def join(other: VSetAutomaton): VSetAutomaton {}
-  
+
   def union(other: VSetAutomaton): VSetAutomaton {}
-  
-  
-  */
-  
-  /*
-   * Transforms the vset-automaton into a vset path union.
-   */
-  def toVSetPathUnion():(ArrayBuffer[ArrayBuffer[(String, Int)]], Array[Int]) = {
-    
-    val (nrStates, initial, transitionGraph, finalStates2) = stateElimination()
-    
-    var pathsToProcess = 1
-    
-    var availablePathVars = new ArrayBuffer[Map[String, Boolean]]()
-    var pathsUnion = new ArrayBuffer[ArrayBuffer[(String, Int)]]()
-    
-    var firstPathVars = Map[String, Boolean]()
-    
-    for(v <- vars)
-      firstPathVars += (v -> true)
-      
-    availablePathVars += firstPathVars
-    
-    pathsUnion += new ArrayBuffer[(String, Int)]()
-    pathsUnion(0) += (("", initial))
-    
-    // Grow the paths iteratively
-    while(pathsToProcess > 0) {
-      
-      var newPaths:ArrayBuffer[ArrayBuffer[(String, Int)]] = null
-      var newPathVars:ArrayBuffer[Map[String, Boolean]] = null
-      
-      // Advance each path by one transition, spawning new paths if there are >1 possible branches
-      for(i <- 0 until pathsUnion.length) {
-        
-        var path = pathsUnion(i)
-        var pathVars = availablePathVars(i)
-        
-        newPaths =  new ArrayBuffer[ArrayBuffer[(String, Int)]]()
-        newPathVars = new ArrayBuffer[Map[String, Boolean]]()
-        
-        val (e, s) = path.last
-        
-        // If we haven't reached a final state in the current path
-        if(!finalStates2.contains(s)) {
-          
-          for((t, e1) <- transitionGraph(s)) {
[…]

### 2016-04-16 project/test/compiler/compilerTestSuite.scala

@@ -8,6 +8,7 @@ class compilerTestSuite extends FunSuite {
     
     val sampleVSet1 = "test/compiler/sampleVSet1.txt"
     val sampleVSet2 = "test/compiler/sampleVSet2.txt"
+    val sampleVSet3 = "test/compiler/sampleVSet3.txt"
     
   }
   
@@ -65,7 +66,7 @@ class compilerTestSuite extends FunSuite {
     
     val (pathUnion, finalStates2) = a.toVSetPathUnion()
     
-     assert(pathUnion.toString() == "ArrayBuffer(ArrayBuffer((,0), (y_in,4), (((b)(y_out)),3)), ArrayBuffer((,0), (x_in,1), (((a)(x_out)),3)))")
+     assert(pathUnion.size == 0)
     
     
   }
@@ -78,6 +79,20 @@ class compilerTestSuite extends FunSuite {
     
     val (pathUnion, finalStates2) = a.toVSetPathUnion()
     
-    assert(pathUnion.toString() == "ArrayBuffer(ArrayBuffer((,6), (((epsilon)(y_in)),4), (((b)(y_out)),3), (epsilon,7)), ArrayBuffer((,6), (((epsilon)(x_in)),1), (((a)(x_out)),3), (((epsilon)(y_in)),4), (((b)(y_out)),3), (epsilon,7)), ArrayBuffer((,6), (((epsilon)(y_in)),4), (((b)(y_out)),3), (((epsilon)(x_in)),1), (((a)(x_out)),3), (epsilon,7)), ArrayBuffer((,6), (((epsilon)(x_in)),1), (((a)(x_out)),3), (epsilon,7)))")
+    assert(pathUnion.toString() == "ArrayBuffer(ArrayBuffer((,null,6), (((epsilon)),Map(x -> in),1), (((a)),Map(x -> out),3), (((epsilon)),Map(y -> in),4), (((b)),Map(y -> out),3), (epsilon,Map(),7)), ArrayBuffer((,null,6), (((epsilon)),Map(y -> in),4), (((b)),Map(y -> out),3), (((epsilon)),Map(x -> in),1), (((a)),Map(x -> out),3), (epsilon,Map(),7)))")
   }
+  
+   test("toVSetPathUnion() should correctly transform sampleVSet3 into a vset path union") {
+     
+    val (nrStates, initial, transitionFunction, vars, finalStates) = VSetAFileReader.getVSetA(files.sampleVSet3)
+    
[…]

### 2016-04-16 project/test/compiler/main.scala

@@ -1,18 +1,23 @@
 package compiler
 
-object main {
+import dk.brics.automaton.RegExp
+
+object main{
   
-    val sampleVSet1 = "test/compiler/sampleVSet1.txt"
-    
+    val sampleVSet = "test/compiler/sampleVSet5.txt"
+  
     def main(args: Array[String]) {
-    
-      val (nrStates, initial, transitionFunction,vars, finalStates) = VSetAFileReader.getVSetA(sampleVSet1)
+      
+      val (nrStates, initial, transitionFunction, vars, finalStates) = VSetAFileReader.getVSetA(sampleVSet)
       
       val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
-    
-      a.toVSetPathUnion()
-    
+      
+      val (pathUnion, finalStates2) = a.toVSetPathUnion
+      
+      val hPU = a.toHybridPathUnion(pathUnion)
+      
+      val lPU = a.toLexicographicPathUnion(hPU)
+
     }
     
-  
 }

### 2016-04-17 project/src/compiler/VSetAutomaton.scala
Lexicographic automaton debugged.

@@ -148,31 +148,40 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
 			for(path <- currentPathUnion) {
 			  var noEps = path.clone
 			  var epsRemoved = false
+			  var newPaths = 0
 			  // Remove all epsilon transitions
-				for(i <- path.length -1 to 2 by -1) {
-
-					val (a, o, t) = noEps(i)
-					// Check if the language of the automaton accepts only the empty string
-					if(a.isEmptyString) {
-
-						// eliminate the epsilon transition and add its variable ops
-						// to the previous transition
-						var part1 = if(i > 1) noEps.slice(0, i-1) else new ArrayBuffer[(Automaton,  Map[String, String], Int)]()
-						val part2 = if (i < noEps.length) noEps.slice(i+1, noEps.length) else null
-						val (a1, o1, t1) = noEps(i - 1)
-						val newOps = o ++ o1
-						part1 += ((a1, newOps, t1))
-						noEps = part1 ++ part2
-						epsRemoved = true
-
-					}
+			  if(noEps.length > 1)  {
+  				for(i <- path.length -1 to 1 by -1) {
+  
+  					val (a, o, t) = noEps(i)
+  					// Check if the language of the automaton accepts only the empty string
+  					if(a.isEmptyString) {
+  
+  						// eliminate the epsilon transition and add its variable ops
+  						// to the previous transition
+  						var part1 = if(i > 1) noEps.slice(0, i-1).clone else new ArrayBuffer[(Automaton,  Map[String, String], Int)]()
+  						val part2 = if (i < noEps.length) noEps.slice(i+1, noEps.length).clone else new ArrayBuffer[(Automaton,  Map[String, String], Int)]()
+  						val (a1, o1, t1) = noEps(i - 1)
+  						val newOps = o ++ o1
+  						
+  						// Fix variables opened and closed at the same time
+  						val commonVars = o.keySet.intersect(o1.keySet)
+  						for(v <- commonVars) {
+  						  
+  						  newOps(v) = "inout"
+  						}
+  						
+  						part1 += ((a1, newOps, t1))
+  						noEps = part1 ++ part2
+  						epsRemoved = true
+  
+  					}
 							
 				}
-			  
-			  var newPaths = 0
+			    
 			  // Remove the empty string from the languages of the transitions, keep both languages
 			  // in different paths
-				for(i <- 2 until noEps.length) {  
[…]

### 2016-04-17 project/test/compiler/main.scala

@@ -4,7 +4,7 @@ import dk.brics.automaton.RegExp
 
 object main{
   
-    val sampleVSet = "test/compiler/sampleVSet5.txt"
+    val sampleVSet = "test/compiler/sampleVSet6.txt"
   
     def main(args: Array[String]) {
       

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/compiler/HybridPathUnion.scala
Refactorings. HybridPathUnion to VSetAutomaton (unfinished).

@@ -1,0 +1,213 @@
+package be.ac.ulb.amorcian.arc.compiler
+
+import scala.collection.mutable.Map
+import scala.collection.mutable.ArrayBuffer
+import dk.brics.automaton.Automaton
+import dk.brics.automaton.State
+import dk.brics.automaton.Transition
+import dk.brics.automaton.RegExp
+
+class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]) {
+  
+  /**
+	 * Makes a hybrid path union lexicographic.
+	 */
+	def toLexicographicPathUnion():HybridPathUnion = {
+
+		// Automaton accepting only the empty string
+		val epsAut = (new RegExp("()")).toAutomaton()
+		
+		epsAut.expandSingleton()
+
+		var currentPathUnion = pathUnion.clone
+		var nextPathUnion = new ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]()
+
+		var canChange = true
+
+		// Make the paths lexicographic
+		while(canChange) {
+		  
+		  canChange = false
+
+			for(path <- currentPathUnion) {
+			  var noEps = path.clone
+			  var epsRemoved = false
+			  var newPaths = 0
+			  // Remove all epsilon transitions
+			  if(noEps.length > 1)  {
+  				for(i <- path.length -1 to 1 by -1) {
+  
+  					val (a, o, t) = noEps(i)
+  					// Check if the language of the automaton accepts only the empty string
+  					if(a.isEmptyString) {
+  
+  						// eliminate the epsilon transition and add its variable ops
+  						// to the previous transition
+  						var part1 = if(i > 1) noEps.slice(0, i-1).clone else new ArrayBuffer[(Automaton,  Map[String, String], Int)]()
+  						val part2 = if (i < noEps.length) noEps.slice(i+1, noEps.length).clone else new ArrayBuffer[(Automaton,  Map[String, String], Int)]()
+  						val (a1, o1, t1) = noEps(i - 1)
+  						val newOps = o ++ o1
+  						
+  						// Fix variables opened and closed at the same time
+  						val commonVars = o.keySet.intersect(o1.keySet)
+  						for(v <- commonVars) {
+  						  
+  						  newOps(v) = "inout"
+  						}
+  						
+  						part1 += ((a1, newOps, t1))
+  						noEps = part1 ++ part2
+  						epsRemoved = true
+  
+  					}
+							
+				}
+			    
[…]

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/compiler/VSetAFileReader.scala

@@ -1,4 +1,4 @@
-package compiler
+package be.ac.ulb.amorcian.arc.compiler
 
 import scala.collection.mutable.Map
 

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/compiler/VSetAutomaton.scala

@@ -1,4 +1,4 @@
-package compiler
+package be.ac.ulb.amorcian.arc.compiler
 
 import scala.collection.mutable.Map
 import scala.collection.mutable.ArrayBuffer
@@ -25,19 +25,19 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
 	 */
 	def join(other: VSetAutomaton):(ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]], Array[Int]) = {
 
-			val (pT, fT) = this.toVSetPathUnion()
-					val (pO, fO) = other.toVSetPathUnion()
+			val pT = this.toVSetPathUnion()
+					val pO = other.toVSetPathUnion()
 
 					// Replace each transition with an automaton
-					var hPT = this.toHybridPathUnion(pT)
-					var hPO = this.toHybridPathUnion(pO)
+					var hPT = pT.toHybridPathUnion
+					var hPO = pO.toHybridPathUnion
 
 					var lexPathUnion = new ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]()
 					// Automaton accepting only the empty string
 					val epsAut = (new RegExp("()")).toAutomaton()
 
 					// Make the paths lexicographic
-					for(path <- hPT) {
+					for(path <- hPT.pathUnion) {
 
 						var newPaths = new ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]()
 
@@ -81,7 +81,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
 
 			var join = new ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]()
 
-					for (pathT <- hPT; pathO <- hPO) {
+					for (pathT <- hPT.pathUnion; pathO <- hPO.pathUnion) {
 
 					} 
 
@@ -92,15 +92,15 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
 	/**
 	 * Projects the vset-automaton on the desired variables.  
 	 */
-	def project(vars: Array[String]):(ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], Array[Int]) = {
+	def project(vars: Array[String]):VSetPathUnion = {
 
 			// get the corresponding vset path union
[…]

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/compiler/VSetPathUnion.scala

@@ -1,0 +1,37 @@
+package be.ac.ulb.amorcian.arc.compiler
+
+import scala.collection.mutable.Map
+import scala.collection.mutable.ArrayBuffer
+import dk.brics.automaton.Automaton
+import dk.brics.automaton.RegExp
+
+class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], var finalStates:Array[Int]) {
+  
+  /**
+   * Transforms a vset path union into an hybrid path union.
+   */
+  def toHybridPathUnion():HybridPathUnion = {
+  
+  		// Replace each transition with an automaton
+  		var hybridPathUnion = new ArrayBuffer[ArrayBuffer[(Automaton, Map[String, String], Int)]]()
+  
+  				for(path <- pathUnion) {
+  
+  					var newPath = new ArrayBuffer[(Automaton, Map[String, String], Int)]()
+  
+  							for((e, ops, t) <- path if ops != null) {
+  
+  								val a:Automaton = (new RegExp(e)).toAutomaton()
+  										a.expandSingleton()
+  
+  										newPath += ((a, ops, t))
+  							}
+  
+  					hybridPathUnion += newPath
+  
+  				}
+  
+  		new HybridPathUnion(hybridPathUnion)
+  
+  }
+}

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/runtime/Instruction.scala

@@ -1,4 +1,4 @@
-package runtime
+package be.ac.ulb.amorcian.arc.runtime
 
 class Instruction(val opCode: InstructionType.Value, val c: Int, val i: Int, val num: Int, val x: Instruction, val y: Instruction)
 

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/runtime/InstructionType.scala

@@ -1,4 +1,4 @@
-package runtime
+package be.ac.ulb.amorcian.arc.runtime
 
 object InstructionType extends Enumeration{
   val CHAR, MATCH, JMP, SPLIT, SAVE, DOT, DIGIT, WHITESPACE = Value

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/runtime/Thread.scala

@@ -1,3 +1,3 @@
-package runtime
+package be.ac.ulb.amorcian.arc.runtime
 
 class Thread(val pc: Instruction, var saved:Array[Int]) {}

### 2016-04-17 project/src/be/ac/ulb/amorcian/arc/runtime/vm.scala

@@ -1,4 +1,4 @@
-package runtime
+package be.ac.ulb.amorcian.arc.runtime
 
 import scala.collection.mutable.MutableList
 

### 2016-04-17 project/test/be/ac/ulb/amorcian/arc/compiler/compilerTestSuite.scala

@@ -1,4 +1,4 @@
-package compiler
+package be.ac.ulb.amorcian.arc.compiler
 
 import org.scalatest.FunSuite
 
@@ -64,9 +64,9 @@ class compilerTestSuite extends FunSuite {
     
     val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
     
-    val (pathUnion, finalStates2) = a.toVSetPathUnion()
+    val pU = a.toVSetPathUnion()
     
-     assert(pathUnion.size == 0)
+     assert(pU.pathUnion.size == 0)
     
     
   }
@@ -77,9 +77,9 @@ class compilerTestSuite extends FunSuite {
     
     val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
     
-    val (pathUnion, finalStates2) = a.toVSetPathUnion()
+    val pU = a.toVSetPathUnion()
     
-    assert(pathUnion.toString() == "ArrayBuffer(ArrayBuffer((,null,6), (((epsilon)),Map(x -> in),1), (((a)),Map(x -> out),3), (((epsilon)),Map(y -> in),4), (((b)),Map(y -> out),3), (epsilon,Map(),7)), ArrayBuffer((,null,6), (((epsilon)),Map(y -> in),4), (((b)),Map(y -> out),3), (((epsilon)),Map(x -> in),1), (((a)),Map(x -> out),3), (epsilon,Map(),7)))")
+    assert(pU.pathUnion.toString() == "ArrayBuffer(ArrayBuffer((,null,6), (((epsilon)),Map(x -> in),1), (((a)),Map(x -> out),3), (((epsilon)),Map(y -> in),4), (((b)),Map(y -> out),3), (epsilon,Map(),7)), ArrayBuffer((,null,6), (((epsilon)),Map(y -> in),4), (((b)),Map(y -> out),3), (((epsilon)),Map(x -> in),1), (((a)),Map(x -> out),3), (epsilon,Map(),7)))")
   }
   
    test("toVSetPathUnion() should correctly transform sampleVSet3 into a vset path union") {
@@ -88,9 +88,9 @@ class compilerTestSuite extends FunSuite {
     
     val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
     
-    val (pathUnion, finalStates2) = a.toVSetPathUnion()
+    val pU = a.toVSetPathUnion()
     
[…]

### 2016-04-17 project/test/be/ac/ulb/amorcian/arc/compiler/main.scala

@@ -1,10 +1,10 @@
-package compiler
+package be.ac.ulb.amorcian.arc.compiler
 
 import dk.brics.automaton.RegExp
 
 object main{
   
-    val sampleVSet = "test/compiler/sampleVSet6.txt"
+    val sampleVSet = "test/be/ac/ulb/amorcian/arc/compiler/sampleVSet6.txt"
   
     def main(args: Array[String]) {
       
@@ -12,11 +12,11 @@ object main{
       
       val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
       
-      val (pathUnion, finalStates2) = a.toVSetPathUnion
+      val pU = a.toVSetPathUnion
       
-      val hPU = a.toHybridPathUnion(pathUnion)
+      val hPU = pU.toHybridPathUnion
       
-      val lPU = a.toLexicographicPathUnion(hPU)
+      val lPU = hPU.toLexicographicPathUnion
 
     }
     

### 2016-04-17 project/test/be/ac/ulb/amorcian/arc/runtime/main.scala

@@ -1,10 +1,10 @@
-package runtime
+package be.ac.ulb.amorcian.arc.runtime
 
 object main {
   
-    val rgx4 = new Array[Instruction](11)
-    
+  
     // matches "a*|b*c+"
+    var rgx4 = new Array[Instruction](11)
     rgx4(2) = new Instruction(InstructionType.CHAR, 'a', -1, 2, null, null)
     rgx4(4) = new Instruction(InstructionType.MATCH, -1, -1, 4, null, null)
     rgx4(6) = new Instruction(InstructionType.CHAR, 'b', -1, 6, null, null)

### 2016-04-17 project/test/be/ac/ulb/amorcian/arc/runtime/runtimeTestSuite.scala

@@ -1,4 +1,4 @@
-package runtime
+package be.ac.ulb.amorcian.arc.runtime
 
 import org.scalatest.FunSuite
 

### 2016-04-18 project/src/be/ac/ulb/amorcian/arc/compiler/HybridPathUnion.scala
HybridPathUnion to VSetAutomaton finished.

@@ -7,7 +7,7 @@ import dk.brics.automaton.State
 import dk.brics.automaton.Transition
 import dk.brics.automaton.RegExp
 
-class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]) {
+class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]], var vars: Array[String]) {
   
   /**
 	 * Makes a hybrid path union lexicographic.
@@ -114,21 +114,110 @@ class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[St
 		  nextPathUnion = new ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]()
 		}
 
-	  new HybridPathUnion(currentPathUnion)
+	  new HybridPathUnion(currentPathUnion, vars)
 
   }
 	
 	/**
 	 * Converts the hybrid path union into a vset-automaton.
 	 */
-	def toVSetAutomaton() = {
+	def toVSetAutomaton(): VSetAutomaton = {
 	  
-	  val sc = 0
+	  var sc = 0
+	  var initial = 0
+	  var transitionFunction = Map[Int, Map[Int, String]]()
+	  var finalStates = new ArrayBuffer[Int]()
+	  
+	  transitionFunction += ((0, Map[Int, String]()))
+	  
+	  var a1:VSetAutomaton = null
+	  var a1Old:VSetAutomaton = null
+	  var eOld = ""
 	  // Process all the edges of all the paths
 	  for(p <- pathUnion) {
-	    
+
+	    // The first edge in the path will always be an epsilon-transition.
 	    val (a, ops, t) = p(0)
-	    var e = ""
+	    var e = flattenVarOps(ops)
+	    if(e != "") {
+	      sc += 1
+	      transitionFunction(0)+= ((sc, e))
+	    }
+	    
+	    
+	    // Process all the other edges.
+	    for(i <- 1 until p.length) {
+	      
+	      val (a, ops, t) = p(i)
+	      
+	      // If there were no variable operations in the previous edge
+	      // and the previous automaton exists,
+	      // connect all the final states of the previous automaton to
+	      // a new state
+	      if(eOld == "" && a1Old != null) {
+	        
+	       sc+= 1
+	       for(s <- a1Old.finalStates) {
+	         
+	         if(!transitionFunction.contains(s))
[…]

### 2016-04-18 project/src/be/ac/ulb/amorcian/arc/compiler/VSetAutomaton.scala

@@ -121,7 +121,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
 
 					}
 
-			new VSetPathUnion(newPathUnion, pU.finalStates)
+			new VSetPathUnion(newPathUnion, vars, pU.finalStates)
 
 	}
 
@@ -260,7 +260,7 @@ def toVSetPathUnion():VSetPathUnion = {
 
 		}
 
-		new VSetPathUnion(pathsUnion, finalStates2)
+		new VSetPathUnion(pathsUnion, vars, finalStates2)
 
 }
 /**

### 2016-04-18 project/src/be/ac/ulb/amorcian/arc/compiler/VSetPathUnion.scala

@@ -5,7 +5,7 @@ import scala.collection.mutable.ArrayBuffer
 import dk.brics.automaton.Automaton
 import dk.brics.automaton.RegExp
 
-class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], var finalStates:Array[Int]) {
+class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], var vars: Array[String], var finalStates:Array[Int]) {
   
   /**
    * Transforms a vset path union into an hybrid path union.
@@ -31,7 +31,7 @@ class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String,
   
   				}
   
-  		new HybridPathUnion(hybridPathUnion)
+  		new HybridPathUnion(hybridPathUnion, vars)
   
   }
 }

### 2016-04-18 reports/04-08-2016.tex
HybridPathUnion to VSetAutomaton finished. Latex version of last report.
This reverts commit 89b57035319482c174d268f2273bf86374167919.

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage[style=authoryear,backend=biber]{biblatex}

\makeatother

\usepackage{babel}
\begin{document}

\title{MEMO-H-504\\
Report of the Meeting of 04/08/16}

\author{Andrea Morciano}

\maketitle
\noindent The people present at the meeting were me, my advisor prof.
Stijn Vansummeren and my supervisor Martin Ugarte.\medskip{}

\noindent I did a presentation divided into three parts:
\begin{itemize}
\item \noindent my work so far;
\item \noindent the current challenges in my work;
\item \noindent Q\&A.
\end{itemize}

\section{My Work}

I wrote two introductory chapters of my thesis. The first one is a
generic overview of Information Extraction and its techniques. The
second is a description of SystemT and its query language AQL. For
the most part of this chapter I describe the formal model for AQL's
core, based on document spanners. My paper is written in Lyx, that
is a document editor based on Latex. It can output PDF documents and
Latex files as well. Martin installed this software on his machine,
in order to be able to directly modify my document.\medskip{}
For what concerns the practical part, I presented the approach that
I chose to evaluate NFAs: a virtual machine that advances all the
branches of execution in lockstep, and that easily supports submatch
extraction. I provided an example, on which we discussed on how this
runtime could realize the execution of a vset-automaton: it should
output all the matches it finds, not only the longest one, which is
determined according to the greedy leftmost policy. I also showed
how I transform a vset-automaton into a vset path union.

\section{Current Challenges}

We further discussed how to execute a vset automaton. There are two
main possibilities: 
\begin{enumerate}
\item using a vset path union (possibly optimized in some way) which has
the property to guarantee that, in each path, each span variable is
opened and closed exactly once;
\item executing the vset-automaton directly. This requires a modification
of the runtime, because each thread needs to keep track of the available
variables, in order to die if there is an atttempt to open a variable
more than once or if upon match some available variables remain (preventing
the match).
\end{enumerate}
We agreed that exploring the convenience of the two approaches could
be a contribution to the research on the topic. Also trying different
runtimes (e.g. OBDDs) could be interesting.\medskip{}
We also spoke about how to perform the intersection of two vset automata.
This operation is needed to implement the $\bowtie$ operator. I presented
some software libraries that can interesect ordinary FSA, but we agreed
that they don't understand vset-automata, that are equipped which
special $\varepsilon$-transitions, i.e. variable operations. Thus,
implementing FSA intersection from scratch would be better. Finally,
we briefly discussed on how to translate a vset-automaton in the format
required by the runtime.

\section{Q\&A}

I asked an opinion on the priority of the tasks I have to perform
and we agreed to concentrate on the ways of executing a vset-automaton
first. We then discussed how to obtain production code using LMS and
LLVM. We also talked about the deployment of SystemT on the machine
of the department, although prof. Vansummeren noticed that a performance
comparison between my system and BigInsight (the platform containing
SystemT) would be of little use, as BigInsights is a much more sophsticated
system. Anyway we agreed to put me in contact with the administrator
of the machine.\medskip{}
I stated that I want to present my thesis at the second session. This
implies using LMS in my work, and having performed experiments with
the system by the beginning of July.
\end{document}

### 2016-04-18 project/src/be/ac/ulb/amorcian/arc/compiler/HybridPathUnion.scala

@@ -7,7 +7,7 @@ import dk.brics.automaton.State
 import dk.brics.automaton.Transition
 import dk.brics.automaton.RegExp
 
-class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]], var vars: Array[String]) {
+class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]) {
   
   /**
 	 * Makes a hybrid path union lexicographic.
@@ -114,110 +114,21 @@ class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[St
 		  nextPathUnion = new ArrayBuffer[ArrayBuffer[(Automaton,  Map[String, String], Int)]]()
 		}
 
-	  new HybridPathUnion(currentPathUnion, vars)
+	  new HybridPathUnion(currentPathUnion)
 
   }
 	
 	/**
 	 * Converts the hybrid path union into a vset-automaton.
 	 */
-	def toVSetAutomaton(): VSetAutomaton = {
+	def toVSetAutomaton() = {
 	  
-	  var sc = 0
-	  var initial = 0
-	  var transitionFunction = Map[Int, Map[Int, String]]()
-	  var finalStates = new ArrayBuffer[Int]()
-	  
-	  transitionFunction += ((0, Map[Int, String]()))
-	  
-	  var a1:VSetAutomaton = null
-	  var a1Old:VSetAutomaton = null
-	  var eOld = ""
+	  val sc = 0
 	  // Process all the edges of all the paths
 	  for(p <- pathUnion) {
-
-	    // The first edge in the path will always be an epsilon-transition.
+	    
 	    val (a, ops, t) = p(0)
-	    var e = flattenVarOps(ops)
-	    if(e != "") {
-	      sc += 1
-	      transitionFunction(0)+= ((sc, e))
-	    }
-	    
-	    
-	    // Process all the other edges.
-	    for(i <- 1 until p.length) {
-	      
-	      val (a, ops, t) = p(i)
-	      
-	      // If there were no variable operations in the previous edge
-	      // and the previous automaton exists,
-	      // connect all the final states of the previous automaton to
-	      // a new state
-	      if(eOld == "" && a1Old != null) {
-	        
-	       sc+= 1
-	       for(s <- a1Old.finalStates) {
-	         
-	         if(!transitionFunction.contains(s))
[…]

### 2016-04-18 project/src/be/ac/ulb/amorcian/arc/compiler/VSetAutomaton.scala

@@ -121,7 +121,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
 
 					}
 
-			new VSetPathUnion(newPathUnion, vars, pU.finalStates)
+			new VSetPathUnion(newPathUnion, pU.finalStates)
 
 	}
 
@@ -260,7 +260,7 @@ def toVSetPathUnion():VSetPathUnion = {
 
 		}
 
-		new VSetPathUnion(pathsUnion, vars, finalStates2)
+		new VSetPathUnion(pathsUnion, finalStates2)
 
 }
 /**

### 2016-04-18 project/src/be/ac/ulb/amorcian/arc/compiler/VSetPathUnion.scala

@@ -5,7 +5,7 @@ import scala.collection.mutable.ArrayBuffer
 import dk.brics.automaton.Automaton
 import dk.brics.automaton.RegExp
 
-class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], var vars: Array[String], var finalStates:Array[Int]) {
+class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], var finalStates:Array[Int]) {
   
   /**
    * Transforms a vset path union into an hybrid path union.
@@ -31,7 +31,7 @@ class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String,
   
   				}
   
-  		new HybridPathUnion(hybridPathUnion, vars)
+  		new HybridPathUnion(hybridPathUnion)
   
   }
 }

### 2016-04-18 project/test/be/ac/ulb/amorcian/arc/compiler/main.scala

@@ -17,8 +17,6 @@ object main{
       val hPU = pU.toHybridPathUnion
       
       val lPU = hPU.toLexicographicPathUnion
-      
-      val a1 = lPU.toVSetAutomaton()
 
     }
     

### 2016-04-18 project/src/be/ac/ulb/amorcian/arc/compiler/VSetPathUnion.scala
Did a mess with revert. Restoring up-to-date files.

@@ -5,7 +5,7 @@ import scala.collection.mutable.ArrayBuffer
 import dk.brics.automaton.Automaton
 import dk.brics.automaton.RegExp
 
-class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], var finalStates:Array[Int]) {
+class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String, String], Int)]], var vars:Array[String], var finalStates:Array[Int]) {
   
   /**
    * Transforms a vset path union into an hybrid path union.
@@ -31,7 +31,7 @@ class VSetPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(String,  Map[String,
   
   				}
   
-  		new HybridPathUnion(hybridPathUnion)
+  		new HybridPathUnion(hybridPathUnion, vars)
   
   }
 }

### 2016-04-18 project/test/be/ac/ulb/amorcian/arc/compiler/main.scala

@@ -17,6 +17,8 @@ object main{
       val hPU = pU.toHybridPathUnion
       
       val lPU = hPU.toLexicographicPathUnion
+      
+      val a1 = lPU.toVSetAutomaton()
 
     }
     

### 2016-04-20 project/src/be/ac/ulb/amorcian/arc/compiler/HybridPathUnion.scala
VSetAutomaton to NFAProgram (needs debugging).

@@ -118,6 +118,8 @@ class HybridPathUnion(var pathUnion: ArrayBuffer[ArrayBuffer[(Automaton,  Map[St
 
   }
 	
+	
+	
 	/**
 	 * Converts the hybrid path union into a vset-automaton.
 	 */

### 2016-04-20 project/src/be/ac/ulb/amorcian/arc/compiler/VSetAutomaton.scala

@@ -5,6 +5,9 @@ import scala.collection.mutable.ArrayBuffer
 import dk.brics.automaton.Automaton
 import dk.brics.automaton.RegExp
 
+import be.ac.ulb.amorcian.arc.runtime.Instruction
+import be.ac.ulb.amorcian.arc.runtime.InstructionType
+
 /**
  * Represents a vset-automaton.
  */
@@ -123,6 +126,161 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
 
 			new VSetPathUnion(newPathUnion, vars, pU.finalStates)
 
+	}
+	
+	/**
+	 * Translates the vset-atomaton into a NFA program.
+	 */
+	def toNFAProgram(program:ArrayBuffer[Instruction], visitedStates:Map[Int, Int], s:Int, pc:Int): (Int, ArrayBuffer[Instruction]) = {
+	  
+	  var npc = pc 
+	  var jumps = new ArrayBuffer[Instruction]()
+	  
+	  visitedStates+= ((s, npc))
+	  
+	  // A final state translates into a match operation, and it 
+	  // doesn't go through the previous block because it doesn't have 
+	  // outgoing transitions
+	  if(finalStates.contains(s)) {
+	    
+	    program += new Instruction(InstructionType.MATCH, -1, -1, npc, null, null)
+      npc += 1
+	  }
+	  else {
+	    
+	    var j = 0
+	    var oldSplit:Instruction = null
+	    
+	    for((t, e) <- transitionFunction(s)) {
+	      
+	      var instr:Instruction = null
+	      var jmp:Instruction = null
+	      var split:Instruction = null
+	      
+  	    
+  	    // If there are >1 outgoing transitions for the current state, we need to put
+    	  // splits
+    	  if(transitionFunction(s).size > 1 && j < transitionFunction.size - 1) { 
+    	    
+    	    split = new Instruction(InstructionType.SPLIT, -1, -1, npc, null, null)
+    	    program += split
+  	      npc += 1
+  	      
+  	      // If it is not the first split, connect the previous to this
+  	      if(oldSplit != null) {
+  	        
+  	        oldSplit.y = split
+  	      }
+    	  }
+	      
+	      val varPatternIn = "(.)_in".r
+  	    val varPatternOut = "(.)_out".r
+  	    val varPattern= ".*(._).*".r
+  	    
+  	    // Find out the type of label and add corresponding instruction to the program
+  	    e match {
+  	      
+  	      case "\\d" => {
+  	        
[…]

### 2016-04-20 project/src/be/ac/ulb/amorcian/arc/runtime/Instruction.scala

@@ -1,4 +1,4 @@
 package be.ac.ulb.amorcian.arc.runtime
 
-class Instruction(val opCode: InstructionType.Value, val c: Int, val i: Int, val num: Int, val x: Instruction, val y: Instruction)
+class Instruction(val opCode: InstructionType.Value, val c: Int, var i: Int, val num: Int, var x: Instruction, var y: Instruction)
 

### 2016-04-20 project/test/be/ac/ulb/amorcian/arc/compiler/main.scala

@@ -1,6 +1,10 @@
 package be.ac.ulb.amorcian.arc.compiler
 
 import dk.brics.automaton.RegExp
+import scala.collection.mutable.Map
+import scala.collection.mutable.ArrayBuffer
+import be.ac.ulb.amorcian.arc.runtime.Instruction
+import be.ac.ulb.amorcian.arc.runtime.InstructionType
 
 object main{
   
@@ -19,6 +23,60 @@ object main{
       val lPU = hPU.toLexicographicPathUnion
       
       val a1 = lPU.toVSetAutomaton()
+      
+      var program = new ArrayBuffer[Instruction]()
+      val pc = 0
+      val s = 0
+      var visitedStates = Map[Int, Int]()
+      
+      a1.toNFAProgram(program, visitedStates, s, pc)
+      
+      for(ins <- program) {
+        
+        ins.opCode match {
+          
+          case InstructionType.CHAR => {
+
+							println("CHAR" + " " + ins.c.toChar)
+						}
+						
+						case InstructionType.DIGIT => {
+
+						  println("DIGIT")
+						}
+						
+						case InstructionType.WHITESPACE => {
+
+						  println("WHITESPACE")
+						}
+						
+						case InstructionType.DOT => {
+						  println("DOT")
+						}
+
+						case InstructionType.MATCH => {
+						  println("MATCH")
+							
+						}
+						
+						
+						case InstructionType.JMP => {
+						  println("JMP" + " " + program.indexOf(ins.x))
+    
+        		}
+        
+        		case InstructionType.SPLIT => {
+        
+        		  println("SPLIT" + " " + program.indexOf(ins.x) + " " + program.indexOf(ins.y))
+        		}
+        
+        		case InstructionType.SAVE => {
+        
+        			println("SAVE" + " " + ins.i)
+        		}
+	
+        }
+      }
 
     }
     

### 2016-04-21 project/src/be/ac/ulb/amorcian/arc/compiler/VSetAutomaton.scala
VSetAutomaton to NFAProgram debugged.

@@ -160,7 +160,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
   	    
   	    // If there are >1 outgoing transitions for the current state, we need to put
     	  // splits
-    	  if(transitionFunction(s).size > 1 && j < transitionFunction.size - 1) { 
+    	  if(transitionFunction(s).size > 1 && j < transitionFunction(s).size - 1) { 
     	    
     	    split = new Instruction(InstructionType.SPLIT, -1, -1, npc, null, null)
     	    program += split
@@ -171,7 +171,9 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
   	        
   	        oldSplit.y = split
   	      }
+
     	  }
+    	  else split = null
 	      
 	      val varPatternIn = "(.)_in".r
   	    val varPatternOut = "(.)_out".r
@@ -184,17 +186,20 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
   	        
   	        instr = new Instruction(InstructionType.DIGIT, -1, -1, npc, null, null)
   	        program += instr
+  	        npc +=1
   	      }
   	      case "." => {
   	        
   	        instr = new Instruction(InstructionType.DOT, -1, -1, npc, null, null)
   	        program += instr
+  	        npc +=1
   	      }
   	      
   	      case "\\s" => {
   	        
   	        instr = new Instruction(InstructionType.WHITESPACE, -1, -1, npc, null, null)
   	        program += instr
+  	        npc +=1
   	      }
   	      
   	      case varPattern(va) => {
@@ -229,6 +234,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
   	             i += 1
 
   	           program += new Instruction(InstructionType.SAVE, -1, i, npc, null, null)
+  	           npc +=1
   	           
   	           if(w == 0)
   	             instr = program.last
@@ -244,6 +250,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
   	        
   	        instr = new Instruction(InstructionType.CHAR, e.charAt(0), -1, npc, null, null)
   	        program += instr
+  	        npc +=1
   	      }
   
   	    }
@@ -253,11 +260,16 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
   	      split.x = instr
[…]

### 2016-04-21 project/test/be/ac/ulb/amorcian/arc/compiler/main.scala

@@ -33,6 +33,8 @@ object main{
       
       for(ins <- program) {
         
+        print(ins.num + " ")
+        
         ins.opCode match {
           
           case InstructionType.CHAR => {

### 2016-05-12 project_sbt/arc/src/main/scala/be.ac.ulb.arc/compiler/VSetAutomaton.scala
Join now uses epsilon closure (needs debugging).

@@ -176,12 +176,6 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
                       break
                     }
                   }
-                  else if(tok2 == "()") {
-
-                    createIfNotExists(s1, s2)
-                    trFuncsIntersection((s1, s2)) += (((s1, t2), tok2))
-                    makeFinalIf(s1, t2)
-                  }
 
               }
               case "\\d" => {
@@ -203,12 +197,6 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
                       isBreak = false
                       break
                     }
-                  }
-                  else if(tok2 == "()") {
-
-                    createIfNotExists(s1, s2)
-                    trFuncsIntersection((s1, s2)) += (((s1, t2), tok2))
-                    makeFinalIf(s1, t2)
                   }
 
               }
@@ -232,12 +220,6 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
                       break
                     }
                   }
-                  else if(tok2 == "()") {
-
-                    createIfNotExists(s1, s2)
-                    trFuncsIntersection((s1, s2)) += (((s1, t2), tok2))
-                    makeFinalIf(s1, t2)
-                  }
 
               }
               case "()" => {
@@ -247,12 +229,6 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
                   createIfNotExists(s1, s2)
                   trFuncsIntersection((s1, s2)) += (((t1, t2), tok))
                   makeFinalIf(t1, t2)
-                }
-                else {
-
-                  createIfNotExists(s1, s2)
-                  trFuncsIntersection((s1, s2)) += (((t1, s2), tok))
-                  makeFinalIf(t1, s2)
                 }
               }
 
@@ -307,12 +283,6 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
                       addChoice(s1, s2, t1, t2, tok)
                     }
                   }
-                  case "()" => {
-
-                    createIfNotExists(s1, s2)
-                    trFuncsIntersection((s1, s2)) += (((s1, t2), tok2))
[…]

### 2016-05-12 project_sbt/arc/src/main/scala/be.ac.ulb.arc/compiler/main.scala

@@ -15,7 +15,9 @@ object main{
 
       val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
 
-      a.epsilonClosure()
+      val b = a.epsilonClosure()
+
+			b.prune()
 
       println("hey")
 

### 2016-05-15 project_sbt/arc/src/main/scala/be.ac.ulb.arc/compiler/VSetAutomaton.scala
Join now based on epsilon closure.

@@ -81,6 +81,22 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
     var trFuncsIntersection = Map[(Int, Int), Map[(Int, Int), String]]()
     var finalStates = new ArrayBuffer[(Int, Int)]()
 
+    var size = 0
+    for((s, ts) <- this.transitionFunction)
+      size += ts.size
+    println("----------------THIS-------------------")
+    println("#states: " + this.transitionFunction.size)
+    println("#transitions: " + size)
+    println()
+
+    size = 0
+    for((s, ts) <- other.transitionFunction)
+      size += ts.size
+    println("----------------OTHER------------------")
+    println("#states: " + other.transitionFunction.size)
+    println("#transitions: " + size)
+    println()
+
 		// Perform epsilon closure on both automata
 		val a1 = this.epsilonClosure()
 		val a2 = other.epsilonClosure()
@@ -92,7 +108,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
     // Make a state final if its components are final
     val makeFinalIf = (t1: Int, t2: Int) => {
 
-                                              if(a1.finalStates.contains(t1) && a2.finalStates.contains(t2))
+                                              if(a1.finalStates.contains(t1) && a2.finalStates.contains(t2) && !finalStates.contains((t1, t2)))
                                                 finalStates += ((t1, t2))
                                             }
 
@@ -186,9 +202,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
                   }
                   else if(tok2 == ".") {
 
-                    createIfNotExists(s1, s2)
-                    trFuncsIntersection((s1, s2)) += (((t1, t2), tok))
-                    makeFinalIf(t1, t2)
+                    addChoice(s1, s2, t1, t2, tok)
                   }
                   else if(tok2.matches(varPattern.toString)) {
                     val varPatternExtr(x)= tok2
@@ -208,9 +222,7 @@ class VSetAutomaton(val nrStates: Int, val initial: Int, val transitionFunction:
                   }
                   else if(tok2 == ".") {
 
-                    createIfNotExists(s1, s2)
-                    trFuncsIntersection((s1, s2)) += (((t1, t2), tok))
[…]

### 2016-05-15 project_sbt/arc/src/main/scala/be.ac.ulb.arc/compiler/main.scala

@@ -6,7 +6,7 @@ import be.ac.ulb.amorcian.arc.runtime.InstructionType
 
 object main{
 
-    val sampleVSet = "src/test/scala/be.ac.ulb.arc/compiler/sampleVSet8.txt"
+    val sampleVSet = "src/test/scala/be.ac.ulb.arc/compiler/sampleVSet6.txt"
     val sampleVSet2 = "src/test/scala/be.ac.ulb.arc/compiler/sampleVSet7.txt"
 
     def main(args: Array[String]) {
@@ -15,11 +15,11 @@ object main{
 
       val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
 
-      val b = a.epsilonClosure()
+			val (nrStates2, initial2, transitionFunction2, vars2, finalStates2) = VSetAFileReader.getVSetA(sampleVSet2)
 
-			b.prune()
+      val b = new VSetAutomaton(nrStates2, initial2, transitionFunction2, vars2, finalStates2)
 
-      println("hey")
+      a.join(b)
 
     }
 

### 2016-05-15 paper/demonstrations.tex
Some definitiions about well-behaved vset-automata.

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,american,draftfoot]{ulb_thesis}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
 \input{ulb_ntheorem.std}

\makeatother

\usepackage{babel}
\begin{document}
$\textrm{SVOps}^{\vdash}=\left\{ x\vdash\mid x\in\textrm{SVars}\right\} $

$\textrm{SVOps}^{\dashv}=\left\{ x\dashv\mid x\in\textrm{SVars}\right\} $
\begin{definition}
An extended \emph{variable-set automaton} (or \emph{eVset-automaton})
is a tuple $\left(Q,q_{0},q_{f},\delta\right)$, where:
\begin{itemize}
\item $Q$ is a finite set of states;
\item $q_{0}\in Q$ is the initial state;
\item $q_{f}\in Q$ is an accepting state;
\item $\delta$ is a finite transition relation consisting of triples, each
having one of these forms:

\begin{itemize}
\item $\left(q,\sigma,q'\right)$, where $\sigma\in\varSigma$;
\item $\left(q,S,q'\right)$, where $S\subseteq\left(\textrm{SVOps}^{\vdash}\cup\textrm{SVOps}^{\dashv}\right)$.
\end{itemize}
\end{itemize}
\end{definition}

\noindent 
\begin{definition}
\noindent For a string $\mathbf{s}=s_{1},\ldots,s_{n}$, a run $\rho=c_{0}\ldots c_{m}$
on $\mathbf{s}$ of a (extended) variable-set automaton $A$ is \emph{well-behaved}
if it satisfies the following conditions:
\begin{itemize}
\item $c_{j}=\left(q,V,Y,i\right)$ where $q\neq q_{f}$ for $j=0,\ldots,m-1$;
\item either:

\begin{itemize}
\item $c_{m}=\left(q,V,Y,i\right)$ where $q\neq q_{f}$;
\item $c_{m}=\left(q_{f},\textrm{Ø},\textrm{Ø},n+1\right)$.
\end{itemize}
\end{itemize}
\end{definition}

\begin{definition}
\noindent A (extended) variable-set automaton is well-behaved if,
for every string $\mathbf{s}\in\varSigma^{\ast}$, $\textrm{ARuns}\left(A,\mathbf{s}\right)$
contains only well-behaved runs.
\end{definition}

\noindent 
\begin{lemma}
\noindent Given an extended varable-set automaton $A$, $A$ can be
converted into an equivalent variable-set automaton $A'$ in polynomial
time.\end{lemma}

\begin{proof}
Consider $A'=$\end{proof}

\end{document}

### 2016-05-16 paper/demonstrations.tex
Part of demonstration wb eVset -> wb vset in poly-time.

@@
 $\textrm{SVOps}^{\dashv}=\left\{ x\dashv\mid x\in\textrm{SVars}\right\} $
+
+SVOps$=\textrm{SVOps}^{\vdash}\cup\textrm{SVOps}^{\dashv}$
 \begin{definition}
@@
 \item $\left(q,\sigma,q'\right)$, where $\sigma\in\varSigma$;
-\item $\left(q,S,q'\right)$, where $S\subseteq\left(\textrm{SVOps}^{\vdash}\cup\textrm{SVOps}^{\dashv}\right)$.
+\item $\left(q,S,q'\right)$, where $S\subseteq\left(\textrm{SVOps}\right)$.
 \end{itemize}
@@
 \end{itemize}
+\item for every $c_{j},c_{j+1}$ and $c_{k},c_{k+1}$, $j\neq k$, such
+that $c_{j}$ yelds $c_{j+1}$ by applying transition $\left(q_{j},o,q_{j+1}\right)\in\delta$
+and the same holds for $c_{k},c_{k+1}$ with transition $\left(q_{k},o',q_{k+1}\right)\in\delta$
+($o,o'\in\textrm{SVops}$$\left(A\right)$), $o\neq o'$;
 \end{itemize}
@@
 \begin{lemma}
-\noindent Given an extended varable-set automaton $A$, $A$ can be
-converted into an equivalent variable-set automaton $A'$ in polynomial
-time.\end{lemma}
+\noindent Given an extended well-behaved varable-set automaton $A$,
+$A$ can be converted into an equivalent well-behaved variable-set
+automaton $A'$ in polynomial time.\end{lemma}
 
 \begin{proof}
-Consider $A'=$\end{proof}
+Without loss of generality, we consider an ordering of the symbols
+in $\textrm{SVOps}\left(A\right)$ of the following form:
+\[
+x\vdash,\ldots,y\vdash,x\dashv,\ldots,y\dashv
+\]
+ In this ordering, all insertion operations come before the deletion
+operations. We define $o\prec o'$, $o,o'\in\textrm{SVOps}\left(A\right)$,
+if $o$ comes before $o'$ (\emph{not} if they are equal) in the chosen
+ordering. Consider $A'=$$\left(Q',q_{0}',q_{f}',\delta'\right)$,
+with $\textrm{SVars}\left(A'\right)=\textrm{SVars}\left(A\right)$,
+whose components are defined as follows:
+\begin{itemize}
+\item $Q'=Q\cup Q^{\textrm{ops}}\cup Q^{\textrm{Ø}}$, where:
+
+\begin{itemize}
+\item $Q^{\textrm{ops}}=\left\{ q_{q',o,q''}\mid\exists\left(q',S,q''\right)\in\delta:o\in S\right\} $;
+\item $Q^{\textrm{Ø}}=\left\{ q_{\textrm{q',Ø,q''}}\mid\exists\left(q',\textrm{Ø},q''\right)\in\delta\right\} $;
+\end{itemize}
+\item $q_{0}'=q_{0};$
+\item $q_{f}'=q_{f}$;
+\item $\delta'=\left(\delta\setminus\delta^{S}\right)\cup\delta^{\textrm{ops}}\cup\delta^{\textrm{Ø}}\cup\delta^{\varepsilon}$,
+where:
+
+\begin{itemize}
+\item $\delta^{S}=\left\{ \left(q,S,q'\right)\in\delta\right\} ;$
+\item $\delta^{\textrm{ops}}=\left\{ \left(q_{q',o,q''},o,q_{q',o',q''}\right)\mid\exists\left(q',S,q''\right)\in\delta:\left(o,o'\epsilon S\land o\prec o'\land\forall o''\in S:o\nprec o''\nprec o'\right)\right\} \cup\left\{ \left(q_{q',o,q''},o,q''\right)\mid\exists\left(q',S,q''\right)\in\delta:\left(o\epsilon S\land\forall o'\in S:o\nprec o'\right)\right\} $;
+\item $\delta^{\textrm{Ø}}=\left\{ \left(q_{\textrm{q',Ø,q''}},\varepsilon,q''\right)\mid\exists\left(q',\textrm{Ø},q''\right)\in\delta\right\} ;$
+\item $\delta^{\varepsilon}=\left\{ \left(q,\varepsilon,q'\right)\mid\left(\exists\left(q',o,q''\right)\in\delta^{\textrm{ops}}:\forall\left(q',o',q''\right)\in\delta^{\textrm{ops}}:o'\nprec o\right)\lor\left(\exists\left(q',\varepsilon,q''\right)\in\delta^{\textrm{Ø}}\right)\right\} $.
+\end{itemize}
+\end{itemize}
+This construction expands the transitions of $A$ that are labeled
+with a set of variable operations into a sequence of transitions performing
+one operation at a time, taking care of putting the insertion operations
+before the deletion ones, thus preserving well-behavedness. It also
+substitutes transitions labeled with the empty set with ordinary $\varepsilon-$transitions.
+To prove equivalence between $A$ and $A'$ it is sufficient to notice
+that for every string $\mathbf{s}$, every run $\rho=c_{0},\ldots,c_{m}$$\in\textrm{ARuns}\left(A,\mathbf{s}\right)$
+can be put in correspondence with a run $\rho'\in\textrm{ARuns}\left(A',\mathbf{s}\right)$.
+Indeed, for every pair of consecutive configurations $c_{j}=\left(q_{j},V_{j},Y_{j},i_{j}\right)$,
+$c_{j+1}=\left(q_{j+1},V_{j+1},Y_{j+1},i_{j+1}\right)$:
+\begin{itemize}
+\item if $i_{j}=i_{j+1}$ and $\left(q,S,q'\right)$$\in\delta$:
+
+\begin{itemize}
+\item if $V_{j}\neq V_{j+1}$, replace $c_{j}$, $c_{j+1}$ with a sequence
+$c_{j,1},\ldots,c_{j,k}$ of configurations, where:
+
+\begin{itemize}
+\item $c_{j,1}=c_{j}$;
+\item $c_{j,k}=c_{j+1}$; 
+\item $c_{j,2}=\left(q_{j,2},V_{j,1},Y_{j,1},i_{j}\right)$ ($\left(q_{j,1},\varepsilon,q_{j,2}\right)\in\delta^{\varepsilon})$;
+\item for every pair $c_{j,l}=\left(q_{j,l},V_{j,l},Y_{j,l},i_{j}\right),c_{j,l+1}=\left(q_{j,l+1},V_{j,l+1},Y_{j,l+1},i_{j}\right)$,
+different from $c_{j,1},c_{j,2}$, there exists a transition $\left(q_{j,l},o,q_{j,l+1}\right)\in\delta^{\textrm{ops}}$
+such that $o\in S$ and $V_{j,l+1},Y_{j,l+1}$ are obtained from $V_{j,l},Y_{j,l}$
+according to $o$;
+\end{itemize}
+\item if $V_{j}=V_{j+1}$, replace $c_{j}$, $c_{j+1}$with $c_{j,1},c_{j,2},c_{j,3}$,
+where:
+
+\begin{itemize}
+\item $c_{j,1}$ and $c_{j,2}$ are as in the previous case;
+\item $c_{j,3}=c_{j+1}$;
+\item there exists a transition $\left(q_{j,2},\textrm{Ø},q_{j,3}\right)\in\delta^{\textrm{Ø}}$
+and $V_{j,3}=V_{j,l}$, $Y_{j,3}=Y_{j,2}$;
+\end{itemize}
+\end{itemize}
+\item if $i_{j+1}=i_{j}+1$, do nothing.
+\end{itemize}
+After having applied this modifications to $\rho$, the obtained run
+$\rho'$ will be a valid run of $A'$ on $\mathbf{s}$ and it will
+be accepting if and only if $\rho$ was accepting, in which case it
+will return the same $\mathbf{s}-$tuple returned by $\rho$. This
+is because the only occasions where we actually modify $\rho$ is
+when we are performing variable operations between a configuration
+and the next one. In this case, we are not advancing the string pointer.
+The new series of configurations we put in place of the original pair
+doesn't advance the pointer either, and it performs the exact variable
+operations that were performed before the substitution, insertions
+first, thanks to the way we constructed $A'$. Thus, for every pair
+of configurations $c_{j},c_{j+1}\in\rho$ that we replace, $c_{j+2}$
+won't see any difference. Also notice how the replaced sequence $c_{j,1},\ldots,c_{j,k}$
+is unique, and the sequence of operations from $c_{j,1}$ to $c_{j,k}$
+respect the operation ordering that we chose.\end{proof}
 

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/compiler/main.scala
Refactored almost everything. Data structures are far more semantic. Code is far more concised, readable and modular, and less error prone.

@@ -1,78 +1,0 @@
-package be.ac.ulb.amorcian.arc.compiler
-
-import scala.collection.mutable.ArrayBuffer
-import be.ac.ulb.amorcian.arc.runtime.Instruction
-import be.ac.ulb.amorcian.arc.runtime.InstructionType
-
-object main{
-
-    val sampleVSet = "src/test/scala/be.ac.ulb.arc/compiler/sampleVSet6.txt"
-    val sampleVSet2 = "src/test/scala/be.ac.ulb.arc/compiler/sampleVSet7.txt"
-
-    def main(args: Array[String]) {
-
-      val (nrStates, initial, transitionFunction, vars, finalStates) = VSetAFileReader.getVSetA(sampleVSet)
-
-      val a = new VSetAutomaton(nrStates, initial, transitionFunction, vars, finalStates)
-
-			val (nrStates2, initial2, transitionFunction2, vars2, finalStates2) = VSetAFileReader.getVSetA(sampleVSet2)
-
-      val b = new VSetAutomaton(nrStates2, initial2, transitionFunction2, vars2, finalStates2)
-
-      a.join(b)
-
-    }
-
-    def printProgram(program: ArrayBuffer[Instruction]) = {
-
-      for(ins <- program) {
-
-        print(ins.num + " ")
-
-        ins.opCode match {
-
-          case InstructionType.CHAR => {
-
-							println("CHAR" + " " + ins.c.toChar)
-						}
-
-						case InstructionType.DIGIT => {
-
-						  println("DIGIT")
-						}
-
-						case InstructionType.WHITESPACE => {
-
-						  println("WHITESPACE")
-						}
-
-						case InstructionType.DOT => {
-						  println("DOT")
-						}
-
-						case InstructionType.MATCH => {
-						  println("MATCH")
-
-						}
-
-
-						case InstructionType.JMP => {
-						  println("JMP" + " " + program.indexOf(ins.x))
-
-        		}
-
-        		case InstructionType.SPLIT => {
-
-        		  println("SPLIT" + " " + program.indexOf(ins.x) + " " + program.indexOf(ins.y))
-        		}
-
-        		case InstructionType.SAVE => {
-
-        			println("SAVE" + " " + ins.i)
-        		}
-
-        }
-      }
-    }
-
-}

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/Instruction.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.amorcian.arc.runtime
+package be.ac.ulb.arc.runtime
 
 class Instruction(val opCode: InstructionType.Value, val c: Int, var i: Int, val num: Int, var x: Instruction, var y: Instruction)
 

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/InstructionType.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.amorcian.arc.runtime
+package be.ac.ulb.arc.runtime
 
 object InstructionType extends Enumeration{
   val CHAR, MATCH, JMP, SPLIT, SAVE, DOT, DIGIT, WHITESPACE = Value

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/Thread.scala

@@ -1,3 +1,3 @@
-package be.ac.ulb.amorcian.arc.runtime
+package be.ac.ulb.arc.runtime
 
 class Thread(val pc: Instruction, var saved:Array[Int]) {}

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/vm.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.amorcian.arc.runtime
+package be.ac.ulb.arc.runtime
 
 import scala.collection.mutable.MutableList
 
@@ -17,7 +17,6 @@ object vm {
 	 * Executes a regular expression program with a given input.
 	 * @param prog the program to execute
 	 * @param input the input for the program
-	 * @param saved a list of saved input indices for submatch tracking
 	 */
 	def evaluate(prog: Array[Instruction], input: Array[Char]):Array[Int] =  {
 

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/vset/Basics.scala

@@ -1,0 +1,289 @@
+package be.ac.ulb.arc.vset
+
+import scala.collection.immutable.{HashSet => SVars}
+import scala.collection.immutable.{HashSet => SVOps}
+import scala.{Int => SVar}
+
+trait ⊢/⊣
+
+class State
+
+case object ⊢ extends ⊢/⊣
+case object ⊣ extends ⊢/⊣
+
+trait types {
+
+  type SVar = Int
+}
+
+class SVOp(val x:SVar, val t: ⊢/⊣) {
+
+  def canEqual(a: Any) = a.isInstanceOf[SVOp]
+
+  override def equals(that: Any): Boolean = {
+
+    that match {
+
+      case that: SVOp =>  {
+
+        that.canEqual(this) && that.x == this.x && that.t == this.t && this.hashCode == that.hashCode
+      }
+      case _ => false
+    }
+  }
+
+  override def hashCode:Int = {
+
+    val prime = 31
+    var result = 1
+
+    result = prime*result + this.x
+    result = prime*result + (if (this.t == null) 0 else this.t.hashCode)
+    return result
+  }
+
+}
+
+class Range(val min:Int, val max:Int) {
+
+  def canEqual(a: Any) = a.isInstanceOf[Range]
+
+  override def equals(that: Any): Boolean = {
+
+    that match {
+
+      case that: Range =>  {
+
+        that.canEqual(this) && that.min == this.min && that.max == this.max && this.hashCode == that.hashCode
+      }
+      case _ => false
+    }
+  }
+
+  override def hashCode:Int = {
+
+    val prime = 31
+    var result = 1
+
+    result = prime*result + min
+    result = prime*result + max
+    return result
+  }
+}
+
+/**
+  * Represents a generic transition.
+  */
+abstract class Transition[State](val q:State, val q1:State) extends types {
+
+  type State2 = (State, State)
+
+  def &(other:Transition[State]):Option[Transition[State2]]
+}
+
+/**
+  * Represents an ordinary transition, with a character label.
+  * @param q
+  * @param σ
+  * @param q1
+  * @tparam State
+  */
+case class OrdinaryTransition[State](override val q: State, σ: Char, override val q1: State) extends Transition[State](q, q1) {
+
+  override def &(other:Transition[State]):Option[Transition[State2]] = {
+
+    other match {
+
+      case OrdinaryTransition(q:State, σ:Char, q1:State) => {
+
[…]

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/vset/VSetAutomaton.scala

@@ -1,0 +1,391 @@
+package be.ac.ulb.arc.vset
+
+import scala.collection.immutable.{HashSet => SVars}
+import scala.collection.immutable.{HashSet => SVOps}
+import scala.collection.immutable.{HashSet => TransitionFunction}
+import scala.collection.immutable.{HashSet => StateSet}
+import scala.collection.immutable.{HashSet => Predecessors}
+import scala.{Int => SVar}
+
+/**
+  * Represents a variable-set automaton.
+  * @param Q
+  * @param q0
+  * @param qf
+  * @param δ
+  */
+class VSetAutomaton(val Q:StateSet[State], val q0:State, val qf:State, val V:SVars[SVar], val δ:TransitionFunction[Transition[State]]) {
+
+  type State2= (State, State)
+
+  /**
+    * Performs the union of this vset-automaton with another one.
+    * @param other
+    * @return
+    */
+  def ∪(other: Option[VSetAutomaton]): Option[VSetAutomaton] = {
+
+    // If we got nothing as parameter, abort (useful for fluent interface)
+    if(other == None) return None
+
+    val o = other.get
+
+    // Determine common variables
+    val cV = this.V.intersect(o.V)
+
+    // If the automata are not variable compatible, abort
+    if (!this.V.diff(cV).isEmpty || !o.V.diff(cV).isEmpty) return None
+
+    // Get the transitions of the other that involve the initial state
+    val oInitials = o.δ.filter((t:Transition[State]) => t.q == o.q0 || t.q1 == o.q0)
+    // Get all the other transitions
+    val oOthers = o.δ.filter((t:Transition[State]) => !oInitials.contains(t))
+
+    // Initialize new transition function
+    var newδ = this.δ ++ oOthers
+
+    // Substitute the other initial state with this initial states
+    // in all transition involving it
+    for(t <- oInitials) {
+
+      val isQ = t.q == o.q0
+      var newT:Transition[State] = null
+
+      t match {
+
+        case OrdinaryTransition(q, σ, q1) => {
+          if(isQ) newT = new OrdinaryTransition[State](q0, σ, q1)
+          else newT = new OrdinaryTransition[State](q, σ, q0)
+        }
+        case RangeTransition(q, σ, q1) => {
+          if(isQ) newT = new RangeTransition[State](q0, σ, q1)
+          else newT = new RangeTransition[State](q, σ, q0)
+        }
[…]

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/vset/VSetAutomatonFileReader.scala

@@ -1,0 +1,106 @@
+package be.ac.ulb.arc.vset
+
+import scala.collection.immutable.{HashSet => SVars}
+import scala.collection.immutable.{HashSet => SVOps}
+import scala.collection.immutable.{HashSet => TransitionFunction}
+import scala.collection.immutable.{HashSet => StateSet}
+import scala.{Int => SVar}
+
+/**
+  * A reader of files containing vset-automata specifications.
+  */
+object VSetAutomatonFileReader {
+
+  /**
+    * Returns the vset-automaton contained in the given file.
+    *
+    * @param file the file name
+    */
+  def getVSetAutomaton(file: String): Option[VSetAutomaton] = {
+
+    import scala.io.Source
+
+    val source = Source.fromFile(file, "UTF-8")
+    val lineIterator = source.getLines
+
+    var QMap = Map[Int, State]()
+
+    var Q = new StateSet[State]
+    var V = new SVars[SVar]
+
+    val QSizeStr = lineIterator.next()
+    val QSize = QSizeStr.toInt
+
+    // Initial state
+    val q0Str = lineIterator.next()
+    val q0Int = q0Str.toInt
+    val q0 = new State
+    Q = Q + q0
+    QMap = QMap + ((q0Int, q0))
+
+    // Final state
+    val qfStr = lineIterator.next()
+    val qfInt = qfStr.toInt
+    val qf = new State
+    Q = Q + qf
+    QMap = QMap + ((qfInt, qf))
+
+    // Span Variables
+    val VStr = lineIterator.next()
+    val VTok = VStr.split("\\s+").map(_.toInt)
+    V = V ++ VTok
+
+    // Transition function and other states
+    val transitionPattern = "(\\d+)\\s(.+)\\s(\\d+)".r
+    var δ = new TransitionFunction[Transition[State]]
+
+    for (l <- lineIterator) {
+
+      val transitionPattern(qStr, label, q1Str) = l
+
+      val qInt = qStr.toInt
+      val q1Int = q1Str.toInt
+
+      var q: State = null
+      var q1: State = null
+
+      // Process q
+      if (!QMap.contains(qInt)) {
+        q = new State
+        Q = Q + q
+        QMap = QMap + ((qInt, q))
+      }
+      else q = QMap(qInt)
+
+      // Process q1
+      if (!QMap.contains(q1Int)) {
+        q1 = new State
+        Q = Q + q1
[…]

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/vsetold/VSetAFileReader.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.amorcian.arc.compiler
+package be.ac.ulb.arc.vsetold
 
 import scala.collection.mutable.Map
 

### 2016-05-18 project_sbt/arc/src/main/scala/be.ac.ulb.arc/vsetold/VSetAutomaton.scala

@@ -1,12 +1,8 @@
-package be.ac.ulb.amorcian.arc.compiler
+package be.ac.ulb.arc.vsetold
 
-import scala.collection.mutable.Map
-import scala.collection.mutable.ArrayBuffer
-import dk.brics.automaton.Automaton
-import dk.brics.automaton.RegExp
+import be.ac.ulb.arc.runtime.{Instruction, InstructionType}
 
-import be.ac.ulb.amorcian.arc.runtime.Instruction
-import be.ac.ulb.amorcian.arc.runtime.InstructionType
+import scala.collection.mutable.{ArrayBuffer, Map}
 
 /**
  * Represents a vset-automaton.
@@ -1011,8 +1007,6 @@ def toVSetPathUnion():VSetPathUnion = {
  */
 def stateElimination(): (Int, Int, Map[Int, Map[Int, String]], Array[Int]) = {
 
-		import scala.collection.mutable.ArrayBuffer
-
 		var newTransitionGraph = this.transitionFunction.clone
 		var initial = this.initial
 		var nrStates = this.nrStates
@@ -1168,8 +1162,6 @@ def normalize(n: Int, i: Int, tr:Map[Int, Map[Int, String]], f: Array[Int]):
   * @return an array of weights for each state
  */
 def computeWeights(n: Int, i: Int, tr:Map[Int, Map[Int, String]], f: Array[Int]): Array[Int] = {
-
-		import scala.util.control.Breaks._
 
 		var transitionGraph = tr
 		var initial = i

### 2016-05-18 project_sbt/arc/src/test/scala/be.ac.ulb.arc/runtime/main.scala

@@ -1,0 +1,31 @@
+package be.ac.ulb.arc.runtime
+
+object main {
+  
+  
+    // matches "a*|b*c+"
+    var rgx4 = new Array[Instruction](11)
+    rgx4(2) = new Instruction(InstructionType.CHAR, 'a', -1, 2, null, null)
+    rgx4(4) = new Instruction(InstructionType.MATCH, -1, -1, 4, null, null)
+    rgx4(6) = new Instruction(InstructionType.CHAR, 'b', -1, 6, null, null)
+    rgx4(8) = new Instruction(InstructionType.CHAR, 99, -1, 8, null, null)
+    rgx4(10) = new Instruction(InstructionType.MATCH, -1, -1, 10, null, null)
+    rgx4(1) = new Instruction(InstructionType.SPLIT, -1, -1, 1, rgx4(2), rgx4(4))
+    rgx4(3) = new Instruction(InstructionType.JMP, -1, -1, 3, rgx4(1), null)
+    rgx4(5) = new Instruction(InstructionType.SPLIT, -1, -1, 5, rgx4(6), rgx4(8))
+    rgx4(9) = new Instruction(InstructionType.SPLIT, -1, -1, 9, rgx4(8), rgx4(10))
+    rgx4(7) = new Instruction(InstructionType.JMP, -1, -1, 7, rgx4(5), null)
+    rgx4(0) = new Instruction(InstructionType.SPLIT, -1, -1, 0, rgx4(1), rgx4(5))
+    
+    val str1 = new Array[Char](3)
+    str1(0) = 'b'
+    str1(1) = 'a'
+    str1(2) = '\0'
+    
+    def main(args: Array[String]) {
+    
+      vm.evaluate(rgx4, str1)
+    }
+    
+  
+}

### 2016-05-18 project_sbt/arc/src/test/scala/be.ac.ulb.arc/runtime/runtimeTestSuite.scala

@@ -1,0 +1,216 @@
+package be.ac.ulb.arc.runtime
+
+import org.scalatest.FunSuite
+
+class runtimeTestSuite extends FunSuite {
+  
+  def progs = new {
+    
+    val rgx1 = new Array[Instruction](3)
+    
+    // matches "ab"
+    rgx1(0) = new Instruction(InstructionType.CHAR, 'a', -1, 0, null, null)
+    rgx1(1) = new Instruction(InstructionType.CHAR, 'b', -1, 1, null, null)
+    rgx1(2) = new Instruction(InstructionType.MATCH, -1, -1, 2, null, null)
+    
+    val rgx2 = new Array[Instruction](4)
+    
+    // matches "a+b"
+    rgx2(0) = new Instruction(InstructionType.CHAR, 'a', -1, 0, null, null)
+    rgx2(2) = new Instruction(InstructionType.CHAR, 'b', -1, 2, null, null)
+    rgx2(1) = new Instruction(InstructionType.SPLIT, -1, -1, 1, rgx2(0), rgx2(2))
+    rgx2(3) = new Instruction(InstructionType.MATCH, -1, -1, 3, null, null)
+   
+    val rgx3 = new Array[Instruction](5)
+    
+    // matches "a*b"
+    rgx3(1) = new Instruction(InstructionType.CHAR, 'a', -1, 1, null, null)
+    rgx3(3) = new Instruction(InstructionType.CHAR, 'b', -1, 3, null, null)
+    rgx3(0) = new Instruction(InstructionType.SPLIT, -1, -1, 0, rgx3(1), rgx3(3))
+    rgx3(2) = new Instruction(InstructionType.JMP, -1, -1, 2, rgx3(0), null)
+    rgx3(4) = new Instruction(InstructionType.MATCH, -1, -1, 4, null, null)
+    
+    val rgx4 = new Array[Instruction](11)
+    
+    // matches "a*|(b*c+)"
+    rgx4(2) = new Instruction(InstructionType.CHAR, 'a', -1, 2, null, null)
+    rgx4(4) = new Instruction(InstructionType.MATCH, -1, -1, 4, null, null)
+    rgx4(6) = new Instruction(InstructionType.CHAR, 'b', -1, 6, null, null)
+    rgx4(8) = new Instruction(InstructionType.CHAR, 'c', -1, 8, null, null)
[…]

### 2016-05-18 project_sbt/arc/src/test/scala/be.ac.ulb.arc/vset/basicsTestSuite.scala

@@ -1,0 +1,218 @@
+package be.ac.ulb.arc.vset
+
+import org.scalatest.FunSuite
+
+import scala.collection.immutable.{HashSet => SVars}
+import scala.collection.immutable.{HashSet => SVOps}
+
+
+class basicsTestSuite extends FunSuite {
+  type State2 = (State, State)
+  def data = new {
+
+    val q1 = new State
+    val q2 = new State
+    val q3 = new State
+    val q4 = new State
+    val q5 = new State
+    val q6 = new State
+
+    val vs = new SVars + 1 + 2 + 3
+    val vs2 = new SVars + 2 + 3
+    val vs3 = new SVars + 4 + 5
+
+    val vop1 = new SVOp(1, ⊢)
+    val vop2 = new SVOp(2, ⊢)
+    val vop3 = new SVOp(3, ⊢)
+    val vop4 = new SVOp(1, ⊢)
+    val vop5 = new SVOp(2, ⊣)
+    val vop6 = new SVOp(2, ⊣)
+    val vop7 = new SVOp(4, ⊣)
+    val vop8 = new SVOp(5, ⊢)
+
+
+    val vos = new SVOps[SVOp] + vop1 + vop2 + vop3
+    val vos2 = new SVOps[SVOp] + vop2 + vop3
+    val vos3 = new SVOps[SVOp] + vop7 + vop8
+    val vos4 = new SVOps[SVOp]
+    val vos5 = new SVOps[SVOp]
+
+    val t1 = new OrdinaryTransition[State](q1, 'g', q2)
+    val t2 = new RangeTransition[State](q3, new Range('e', 'i'), q4)
+    val t3 = new OperationsTransition[State](q5, vos, vs, q6)
+    val t4 = new OperationsTransition[State](q3, vos2, vs2, q4)
+    val t5 = new RangeTransition[State](q3, new Range('g', 'o'), q4)
+    val t6 = new RangeTransition[State](q3, new Range('b', 'h'), q4)
+    val t7 = new RangeTransition[State](q3, new Range('s', 'u'), q4)
+    val t8 = new RangeTransition[State](q3, new Range('a', 'b'), q4)
+    val t9 = new RangeTransition[State](q3, new Range('a', 'g'), q4)
+    val t10 = new OperationsTransition[State](q3, vos2, vs, q4)
+    val t11 = new OperationsTransition[State](q3, vos3, vs3, q4)
+    val t12 = new OperationsTransition[State](q3, vos4, vs2, q4)
[…]

### 2016-05-18 project_sbt/arc/src/test/scala/be.ac.ulb.arc/vset/main.scala

@@ -1,0 +1,74 @@
+package be.ac.ulb.arc.vset
+
+import scala.collection.mutable.ArrayBuffer
+import be.ac.ulb.arc.runtime.Instruction
+import be.ac.ulb.arc.runtime.InstructionType
+
+import dk.brics.automaton.RegExp
+
+object main{
+
+	val sampleVSet = "src/test/scala/be.ac.ulb.arc/compiler/sampleVSet6.txt"
+	val sampleVSet2 = "src/test/scala/be.ac.ulb.arc/compiler/sampleVSet8.txt"
+
+	def main(args: Array[String]) {
+
+	val test = new RegExp(".").toAutomaton()
+
+		println("hey")
+
+	}
+
+	def printProgram(program: ArrayBuffer[Instruction]) = {
+
+		for(ins <- program) {
+
+			print(ins.num + " ")
+
+			ins.opCode match {
+
+				case InstructionType.CHAR => {
+
+					println("CHAR" + " " + ins.c.toChar)
+				}
+
+				case InstructionType.DIGIT => {
+
+					println("DIGIT")
+				}
+
+				case InstructionType.WHITESPACE => {
+
+					println("WHITESPACE")
+				}
+
+				case InstructionType.DOT => {
+					println("DOT")
+				}
+
+				case InstructionType.MATCH => {
+					println("MATCH")
+
+				}
+
+
+				case InstructionType.JMP => {
+					println("JMP" + " " + program.indexOf(ins.x))
+
+				}
+
+				case InstructionType.SPLIT => {
+
+					println("SPLIT" + " " + program.indexOf(ins.x) + " " + program.indexOf(ins.y))
+				}
+
+				case InstructionType.SAVE => {
+
+					println("SAVE" + " " + ins.i)
+				}
+
+			}
+		}
+	}
+
+}

### 2016-05-18 project_sbt/arc/src/test/scala/be.ac.ulb.arc/vset/vSetAutomatonTestSuite.scala

@@ -1,0 +1,28 @@
+package be.ac.ulb.arc.vset
+
+import org.scalatest.FunSuite
+
+/**
+  * Created by andrea on 18/05/16.
+  */
+class vSetAutomatonTestSuite extends FunSuite{
+
+  val data = new {
+
+    val vsetFile1 = "src/test/scala/be.ac.ulb.arc/vset/sampleVSet6.txt"
+    val vsetFile2 = "src/test/scala/be.ac.ulb.arc/vset/sampleVSet7.txt"
+  }
+
+  test("Epsilon closure should add the correct transitions to vset-automata") {
+
+    val vset1 = VSetAutomatonFileReader.getVSetAutomaton(data.vsetFile2).get
+
+    val vset2 = vset1.ε()
+
+    val newTr = vset2.δ.diff(vset1.δ)
+
+    assert(newTr.size == 6)
+
+  }
+
+}

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/compiler/DSL.scala
Reimplemented virtual machine too. Pretty print of spanned tuples. Integrated LMS calsses from tutorials to make experiments.

@@ -1,0 +1,4 @@
+package be.ac.ulb.arc.compiler
+
+import scala.lms.common._
+

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/Instruction.scala

@@ -1,4 +1,20 @@
 package be.ac.ulb.arc.runtime
 
-class Instruction(val opCode: InstructionType.Value, val c: Int, var i: Int, val num: Int, var x: Instruction, var y: Instruction)
+import scala.{Int => Position}
+import scala.{Int => Pointer}
+import be.ac.ulb.arc.vset.Range
 
+// A generic instruction
+abstract class Instruction(val pos:Position)
+// Matches a char
+case class CHAR(val c:Char, override val pos:Position) extends Instruction(pos)
+// Matches a char in a range
+case class RANGE(val r:Range, override val pos:Position) extends Instruction(pos)
+// Splits the thread of control
+case class SPLIT(override val pos:Position, val next1:Instruction, val next2:Instruction) extends Instruction(pos)
+// Jumps to target instruction
+case class JUMP(override val pos:Position, val target:Instruction) extends Instruction(pos)
+// Saves the current string pointer to the specified pointer
+case class SAVE(val ptr:Pointer, override val pos:Position) extends Instruction(pos)
+// Matches the input
+case class MATCH(override val pos:Position) extends Instruction(pos)

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/Thread.scala

@@ -1,3 +1,6 @@
 package be.ac.ulb.arc.runtime
 
-class Thread(val pc: Instruction, var saved:Array[Int]) {}
+import scala.{Int => Position}
+import scala.{Array => Pointers}
+
+class Thread(val instr:Instruction, val saved:Pointers[Position])

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtimeold/Instruction.scala

@@ -1,0 +1,4 @@
+package be.ac.ulb.arc.runtimeold
+
+class Instruction(val opCode: InstructionType.Value, val c: Int, var i: Int, val num: Int, var x: Instruction, var y: Instruction)
+

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtimeold/InstructionType.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.arc.runtime
+package be.ac.ulb.arc.runtimeold
 
 object InstructionType extends Enumeration{
   val CHAR, MATCH, JMP, SPLIT, SAVE, DOT, DIGIT, WHITESPACE = Value

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtimeold/Thread.scala

@@ -1,0 +1,3 @@
+package be.ac.ulb.arc.runtimeold
+
+class Thread(val pc: Instruction, var saved:Array[Int]) {}

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtimeold/vm.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.arc.runtime
+package be.ac.ulb.arc.runtimeold
 
 import scala.collection.mutable.MutableList
 

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/vset/VSetAutomaton.scala

@@ -343,6 +343,14 @@ class VSetAutomaton(val Q:StateSet[State], val q0:State, val qf:State, val V:SVa
 
   }
 
+  /**
+    * Replaces state pairs with simple states in the provided state set and transition function.
+    * @param Q2
+    * @param q02
+    * @param qf2
+    * @param δ2
+    * @return
+    */
   def State2toState(Q2:StateSet[State2], q02:State2, qf2:State2, δ2:TransitionFunction[Transition[State2]])
   : (StateSet[State], Option[State], Option[State], TransitionFunction[Transition[State]]) = {
 

### 2016-05-19 project_sbt/arc/src/main/scala/be.ac.ulb.arc/vsetold/VSetAutomaton.scala

@@ -1,6 +1,6 @@
 package be.ac.ulb.arc.vsetold
 
-import be.ac.ulb.arc.runtime.{Instruction, InstructionType}
+import be.ac.ulb.arc.runtimeold.{Instruction, InstructionType}
 
 import scala.collection.mutable.{ArrayBuffer, Map}
 

### 2016-05-19 project_sbt/arc/src/main/scala/scala/lms/tutorial/dslapi.scala

@@ -1,0 +1,298 @@
+package scala.lms.tutorial
+
+import scala.lms.common._
+import scala.reflect.SourceContext
+
+// should this be added to LMS?
+trait UtilOps extends Base { this: Dsl =>
+  def infix_HashCode[T:Typ](o: Rep[T])(implicit pos: SourceContext): Rep[Long]
+  def infix_HashCode(o: Rep[String], len: Rep[Int])(implicit v: Overloaded1, pos: SourceContext): Rep[Long]
+}
+trait UtilOpsExp extends UtilOps with BaseExp { this: DslExp =>
+  case class ObjHashCode[T:Typ](o: Rep[T])(implicit pos: SourceContext) extends Def[Long] { def m = typ[T] }
+  case class StrSubHashCode(o: Rep[String], len: Rep[Int])(implicit pos: SourceContext) extends Def[Long]
+  def infix_HashCode[T:Typ](o: Rep[T])(implicit pos: SourceContext) = ObjHashCode(o)
+  def infix_HashCode(o: Rep[String], len: Rep[Int])(implicit v: Overloaded1, pos: SourceContext) = StrSubHashCode(o,len)
+
+  override def mirror[A:Typ](e: Def[A], f: Transformer)(implicit pos: SourceContext): Exp[A] = (e match {
+    case e@ObjHashCode(a) => infix_HashCode(f(a))(e.m,pos)
+    case e@StrSubHashCode(o,len) => infix_HashCode(f(o),f(len))
+    case _ => super.mirror(e,f)
+  }).asInstanceOf[Exp[A]]
+}
+trait ScalaGenUtilOps extends ScalaGenBase {
+  val IR: UtilOpsExp
+  import IR._
+
+  override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
+    case ObjHashCode(o) => emitValDef(sym, src"$o.##")
+    case _ => super.emitNode(sym, rhs)
+  }
+}
+trait CGenUtilOps extends CGenBase {
+  val IR: UtilOpsExp
+  import IR._
+
+  override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
+    case StrSubHashCode(o,len) => emitValDef(sym, src"hash($o,$len)")
+    case _ => super.emitNode(sym, rhs)
+  }
+}
+
+
[…]

### 2016-05-19 project_sbt/arc/src/main/scala/scala/lms/tutorial/utils.scala

@@ -1,0 +1,143 @@
+package scala.lms.tutorial
+
+import java.io._
+
+import org.scalatest.FunSuite
+
+trait LibSuite extends FunSuite {
+  def dataFilePath(csv: String) = "src/data/" + csv
+}
+
+trait TutorialFunSuite extends LibSuite {
+  val overwriteCheckFiles = false // should be false; temporary set to true only to simplify development
+
+  val prefix = "src/out/"
+  val under: String
+  def readFile(name: String): String = {
+    try {
+      val buf = new Array[Byte](new File(name).length().toInt)
+      val fis = new FileInputStream(name)
+      fis.read(buf)
+      fis.close()
+      new String(buf)
+    } catch {
+      case e: IOException => ""
+    }
+  }
+  def writeFile(name: String, content: String) {
+    val out = new java.io.PrintWriter(new File(name))
+    out.write(content)
+    out.close()
+  }
+  def writeFileIndented(name: String, content: String) {
+    val out = new java.io.PrintWriter(new File(name))
+    printIndented(content)(out)
+    out.close()
+  }
+  def checkOut(label: String, suffix: String, thunk: => Unit) = {
+    val output = new ByteArrayOutputStream()
+    scala.Console.setOut(new PrintStream(output))
+    thunk
+    check(label, output.toString(), suffix = suffix)
+  }
+  def check(label: String, raw_code: String, suffix: String = "scala") = {
+    val fileprefix = prefix+under+label
+    val name = fileprefix+".check."+suffix
+    val aname = fileprefix+".actual."+suffix
+    val expected = readFile(name)
+    val code = indent(raw_code)
+    if (expected != code) {
+      val wname = if (overwriteCheckFiles) name else aname
+      println("writing " + wname)
+      writeFile(wname, code)
+    } else {
+      val f = new File(aname)
+      if (f.exists) f.delete
+    }
+    if (!overwriteCheckFiles) {
+      assert(expected == code, name)
+    }
+  }
+  def indent(str: String) = {
+    val s = new StringWriter
+    printIndented(str)(new PrintWriter(s))
+    s.toString
+  }
+  def printIndented(str: String)(out: PrintWriter): Unit = {
+    val lines = str.split("[\n\r]")
+    var indent = 0
+    for (l0 <- lines) {
+      val l = l0.trim
+      if (l.length > 0) {
+        var open = 0
+        var close = 0
+        var initClose = 0
+        var nonWsChar = false
[…]

### 2016-05-19 project_sbt/arc/src/test/scala/be.ac.ulb.arc/compiler/compilerTestSuite.scala

@@ -1,0 +1,16 @@
+package be.ac.ulb.arc.compiler
+
+import org.scalatest.FunSuite
+import scala.lms.common._
+
+/**
+  * Created by andrea on 19/05/16.
+  */
+class compilerTestSuite extends FunSuite {
+
+  test("Tying out LMS functions") {
+
+
+  }
+
+}

### 2016-05-19 project_sbt/arc/src/test/scala/be.ac.ulb.arc/runtime/virtualMachineTestSuite.scala

@@ -1,0 +1,81 @@
+package be.ac.ulb.arc.runtime
+
+import org.scalatest.FunSuite
+
+import scala.collection.immutable.{HashSet => SVars}
+import scala.{Int => SVar}
+import scala.{Int => Position}
+import scala.{Array => Pointers}
+import scala.{Array => Program}
+import scala.collection.immutable.{HashSet => VSRelation}
+import scala.{Array => VSTuple}
+import scala.collection.mutable.{MutableList => ThreadList}
+
+/**
+  * Created by andrea on 19/05/16.
+  */
+class virtualMachineTestSuite extends FunSuite{
+
+  def data = new {
+
+    val prog1 = new Program[Instruction](10)
+    prog1(0) = new SAVE(0, 0)
+    prog1(1) = new CHAR('a', 1)
+    prog1(3) = new CHAR('a', 3)
+    prog1(5) = new SAVE(1, 5)
+    prog1(2) = new SPLIT(2, prog1(3), prog1(5))
+    prog1(4) = new JUMP(4, prog1(2))
+    prog1(6) = new SAVE(2, 6)
+    prog1(7) = new CHAR('b', 7)
+    prog1(8) = new SAVE(3, 8)
+    prog1(9) = new MATCH(9)
+
+    val str1 = "aaaaab" + '\0'
+
+    val prog2 = new Program[Instruction](11)
+    prog2(0) = new SAVE(0, 0)
+    prog2(2) = new CHAR('a', 2)
+    prog2(4) = new SAVE(1, 4)
+    prog2(1) = new SPLIT(1, prog2(2), prog2(4))
+    prog2(3) = new JUMP(3, prog2(1))
+    prog2(5) = new SAVE(2, 5)
+    prog2(7) = new CHAR('a', 7)
+    prog2(9) = new SAVE(3, 9)
+    prog2(6) = new SPLIT(7, prog2(7), prog2(9))
+    prog2(8) = new JUMP(8, prog2(6))
+    prog2(10) = new MATCH(10)
+
+    val str2 = "aaaab" + '\0'
+
+  }
+
+  test("VM should match prog1 on str1") {
+
+    val vars = new SVars[SVar] + 1 + 2
+
+    val tuplesOpt = VirtualMachine.execute(data.prog1, vars, data.str1)
+
+    assert(tuplesOpt != None)
+
+    val tuples = tuplesOpt.get
+
+    assert(tuples.size == 1)
[…]

### 2016-05-19 project_sbt/arc/src/test/scala/be.ac.ulb.arc/runtimeold/main.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.arc.runtime
+package be.ac.ulb.arc.runtimeold
 
 object main {
   

### 2016-05-19 project_sbt/arc/src/test/scala/be.ac.ulb.arc/runtimeold/runtimeTestSuite.scala

@@ -1,4 +1,4 @@
-package be.ac.ulb.arc.runtime
+package be.ac.ulb.arc.runtimeold
 
 import org.scalatest.FunSuite
 

### 2016-05-19 project_sbt/arc/src/test/scala/be.ac.ulb.arc/vset/main.scala

@@ -1,8 +1,8 @@
 package be.ac.ulb.arc.vset
 
 import scala.collection.mutable.ArrayBuffer
-import be.ac.ulb.arc.runtime.Instruction
-import be.ac.ulb.arc.runtime.InstructionType
+import be.ac.ulb.arc.runtimeold.Instruction
+import be.ac.ulb.arc.runtimeold.InstructionType
 
 import dk.brics.automaton.RegExp
 

### 2016-05-19 project_sbt/arc/src/test/scala/scala/lms/tutorial/lmsTestSuite.scala

@@ -1,0 +1,41 @@
+package scala.lms.tutorial
+
+import scala.lms.common._
+
+
+class lmsTestSuite extends TutorialFunSuite {
+  val under = "dslapi"
+
+  test("1") {
+    val snippet = new DslDriver[Int,Int] {
+      def snippet(x: Rep[Int]) = {
+
+        def compute(b: Boolean): Rep[Int] = {
+          // the if is executed in the first stage
+          if (b) 1 else x
+        }
+        compute(true)+compute(1==1)
+
+      }
+    }
+    assert(snippet.eval(0) === 2)
+    check("1", snippet.code)
+  }
+
+  test("2") {
+    val snippet = new  DslDriver[Int,Int] {
+      def snippet(x: Rep[Int]) = {
+
+        def compute(b: Rep[Boolean]): Rep[Int] = {
+          // the if is deferred to the second stage
+          if (b) 1 else x
+        }
+        compute(x==1)
+
+      }
+    }
+    assert(snippet.eval(2) === 2)
+    check("2", snippet.code)
+  }
+
+}

### 2016-06-06 paper/demonstrations.tex
Theory integrated in paper.

@@
 \usepackage{amssymb}
+\usepackage{stmaryrd}
 
@@
 \item $q_{f}\in Q$ is an accepting state;
-\item $\delta$ is a finite transition relation consisting of triples, each
-having one of these forms:
+\item $\delta=\delta^{char}\cup\delta^{op}$ is a finite transition relation
+consisting of triples, where:
 
 \begin{itemize}
-\item $\left(q,\sigma,q'\right)$, where $\sigma\in\varSigma$;
-\item $\left(q,S,q'\right)$, where $S\subseteq\left(\textrm{SVOps}\right)$.
+\item $\delta^{char}=\left\{ \left(q,\sigma,q'\right)\mid q,q'\in Q,\sigma\in\varSigma\right\} $,
+whose elments are called character transitions;
+\item $\delta^{op}=\left\{ \left(q,S,q'\right)\mid q,q'\in Q,S\subseteq\left(\textrm{SVOps}\right)\right\} $,
+whose elements are called operation transitions .
 \end{itemize}
@@
 
-\noindent 
 \begin{definition}
-\noindent For a string $\mathbf{s}=s_{1},\ldots,s_{n}$, a run $\rho=c_{0}\ldots c_{m}$
-on $\mathbf{s}$ of a (extended) variable-set automaton $A$ is \emph{well-behaved}
-if it satisfies the following conditions:
+Given a transition $t=\left(q,S,q'\right)\in\delta^{op}$ in an eVset-automaton
+$A$, and an ordering $\varphi$ on the elements of $\textrm{SVOps}\left(A\right)$,
+we define:
 \begin{itemize}
-\item $c_{j}=\left(q,V,Y,i\right)$ where $q\neq q_{f}$ for $j=0,\ldots,m-1$;
-\item either:
-
-\begin{itemize}
-\item $c_{m}=\left(q,V,Y,i\right)$ where $q\neq q_{f}$;
-\item $c_{m}=\left(q_{f},\textrm{Ø},\textrm{Ø},n+1\right)$.
-\end{itemize}
-\item for every $c_{j},c_{j+1}$ and $c_{k},c_{k+1}$, $j\neq k$, such
-that $c_{j}$ yelds $c_{j+1}$ by applying transition $\left(q_{j},o,q_{j+1}\right)\in\delta$
-and the same holds for $c_{k},c_{k+1}$ with transition $\left(q_{k},o',q_{k+1}\right)\in\delta$
-($o,o'\in\textrm{SVops}$$\left(A\right)$), $o\neq o'$;
+\item $\textrm{LOps}_{\varphi}\left(t\right)$ as the list $o_{1},\ldots,o_{\left|S\right|}$
+of operations belonging to $S$, ordered according to $\varphi$;
+\item $\textrm{Ops}\left(t\right)=S$.
 \end{itemize}
@@
 \begin{definition}
-\noindent A (extended) variable-set automaton is well-behaved if,
-for every string $\mathbf{s}\in\varSigma^{\ast}$, $\textrm{ARuns}\left(A,\mathbf{s}\right)$
-contains only well-behaved runs.
+\noindent Given an extended vset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$
+and a pair of states $q,q'\in Q$, a path $p$ between $q$ and $q'$
+in $A$ is a sequence of transitions $t_{1},$$\ldots,t_{n}\in\delta$,
+such that:
+\begin{itemize}
+\item the source state of $t_{1}$ is $q$;
+\item the destination state of $t_{n}$ is $q'$;
+\item \noindent for every pair $t_{i},t_{i+1}$, the destination state of
+$t_{i}$ equals the source state of $t_{i+1}$.
+\end{itemize}
+We refer to the set of paths in $A$ as $\textrm{Paths}\left(A\right)$.
+\end{definition}
+
+\begin{definition}
+\noindent Given an eVset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$
+a path $p=t_{1},\ldots,t_{n}$ in $A$ is complete if it is between
+$q_{0}$ and $q_{f}$.
+\end{definition}
+
+\begin{definition}
+Given a path $p=t_{1},\ldots,t_{n}$ in an eVset-automaton $A$, and
+considering the sublist $p^{op}=t_{j},\ldots t_{k}$, where $\forall t_{j}$,
+$t_{j}\in\delta^{op}$ and an ordering $\varphi$ on the elements
+of $\textrm{SVOps}\left(A\right)$, we define
+\begin{itemize}
+\item the list $\textrm{LOps}_{\varphi}\left(p\right)$ as:
+\[
+o_{j,1},\ldots,o_{j,l_{j}},\ldots,o_{k,1},\ldots,o_{k,l_{k}}
+\]
+where $\forall j,m$, $o_{j,m}$$\in S_{j}$, and the order of the
+operations belonging to the same $S_{j}$ is fixed according to $\varphi$;
+\item the set $\textrm{Ops}\left(p\right)$ as the set of operations appearing
+in $\textrm{LOps}_{\varphi}\left(p\right)$.
+\end{itemize}
+If $p^{op}=p$, we say that $p$ is an operation-only path.
+\end{definition}
+
+\begin{definition}
+\noindent \label{well-behaved}A (extended) variable-set automaton
+$A$ is well-behaved if, for every complete path $p$ in $A$ we have:
+\begin{enumerate}
+\item for every $o\in\textrm{SVOps}\left(A\right)$, $o$ appears exactly
+once in $\textrm{LOps}\left(p\right)$;\label{enu:well-behaved-1,}
+\item for every pair of operations $x\vdash$,$\dashv x$$\in\textrm{SVOps}\left(A\right)$,
+$x\vdash$ appears before $\dashv x$ in $\textrm{LOps}\left(p\right)$.
+\end{enumerate}
 \end{definition}
@@
 \noindent 
+\begin{definition}
+Given an eVset-automaton $A$, $A$ is pruned if for every state $q$
+in $A$, there is a path from $q_{0}$ to $q$, and a path from $q$
+to $q_{f}$ in $A$.
+\end{definition}
+
+\noindent 
+\begin{proposition}
+Given a well-behaved eVset-automaton $A$, there exists a pruned well-behaved
+eVset automaton $A'$ which is equivalent to $A$.\end{proposition}
+
+\begin{proof}
+To obtain $A'$, it is sufficient to remove from $A$ those states
+that don't participate in any complete path in $A$, and all the transitions
+involving them. Because of our construction, $A'$ is well-behaved,
+as it now contains only the complete paths in $A$. To see that $A'$
+is equivalent to $A$, consider a state $q$ of $A$ that is not present
+in $A'$. $q$ could never occur in a configuration of an accepting
+run of $A$, because either it is unreachable from the initial state
+of $A$, either it couldn't reach the final state of $A$. Thus, given
+an accepting run of $A$, there will be a corresponding accepting
+run of $A'$. $A'$ cannot have an accepting run that can't be put
+in correspondence with an accepting run of $A$ because we do not
+add any transitions to $A'$ that aren't present in $A$.
+\end{proof}
+
+\noindent 
+\begin{corollary}
+\label{same-ops}Given a pruned well-behaved eVset-automaton $A$,
+and two states $q,q'$ in $A$, for every pair of paths $p,p'$ between
+$q$ and $q'$ in $A$, $\textrm{Ops}\left(p\right)=\textrm{Ops}\left(p'\right)$.\end{corollary}
+
+\begin{proof}
+Since $A$ is pruned, there surely exists a path $p^{0}$ between
+$q_{0}$ and $q$. For the same reason, there exists a path $p^{f}$
+between $q'$ and $q_{f}$. So $q$ and $q'$ appear in two complete
+paths, which we may call $p^{0}pp^{f}$ and $p^{0}p'p^{f}$, that
+differ only in the subpaths between $q$ and $q'$. If the statement
+was false, then one of $p^{0}pp^{f}$ and $p^{0}p'p^{f}$ wouldn't
+satisfy the requirement \ref{enu:well-behaved-1,} of \ref{well-behaved},
+making $A$ not well-behaved, a contraddiction.
+\end{proof}
+
+\noindent 
+\begin{definition}
+Given a eVset-automaton $A$, $A$ is operation-closed if, for every
+pair of states $q,q'$ in $A$, whenever there exists an operation-only
+path $p$=$t_{1},\ldots,t_{n}$ between $q$ and $q'$, then there
+exists a transition in $A$ of the form $\left(q,\bigcup_{1\leq i\leq n}\textrm{Ops}\left(t_{i}\right),q'\right)$.
+\end{definition}
+
+\noindent 
+\begin{proposition}
+Given a well-behaved eVset-automaton $A$, there exists an operation-closed
+well-behaved eVset-automaton $A'$ that is equivalent to $A$.\end{proposition}
+
+\begin{proof}
+Let us consider $A$ to be of the form $\left(Q,q_{0},q_{f},\delta\right)$.
+Without loss of generality, we can assume that $A$ is pruned. Then
+we can take $A'=\left(Q',q_{0}',q_{f}',\delta'\right)$, where:
+\begin{itemize}
+\item $Q'=Q$;
+\item $q_{0}'=q_{0}$;
+\item $q_{f}'=q_{f}$;
+\item $\delta'=\delta\cup\left\{ \left(q,\bigcup_{1\leq i\leq n}\textrm{Ops}\left(t_{i}\right),q'\right)\mid\exists p=t_{1},\ldots,t_{n}\in\textrm{Paths}\left(A\right):p\textrm{ is between \ensuremath{q} and \ensuremath{q'}, \ensuremath{p} is operation-only}\right\} $.
+\end{itemize}
+This construction does nothing but including the transitions that
+are missing from $A$ to be operation-closed to $A'$. We now show
+that $A'$ is equivalent to $A$. It is easy to see that, for every
+string $\mathbf{s}\in\varSigma^{\ast}$, $A\left(\mathbf{s}\right)\subseteq A'\left(\mathbf{s}\right)$,
+as we include all transitions of $A$ in $A'$. To see that $A'\left(\mathbf{s}\right)\subseteq A\left(\mathbf{s}\right)$,
+consider a pair of states $q,q'\in Q$ and let us assume there exists
+an operation-only path $p'=t$ between $q$ and $q'$ in $A$', where
+$t$ is a transition that is not in $A$. By construction of $A'$,
+there exists in $A$ an operation-only path $p$ between $q$ and
+$q'$, and$\textrm{LOps}\left(p'\right)=\textrm{LOps}\left(p\right)$.
+Thus, $p'$ is equivalent to $p$. This also shows that $A'$ is well-behaved.
+\end{proof}
+
+\noindent 
+\begin{definition}
+Given two well-behaved operation-closed eVset-automata $A=\left(Q_{A},q_{A}^{0},q_{A}^{f},\delta_{A}\right)$
+and $B=\left(Q_{B},q_{B}^{0},q_{B}^{f},\delta_{B}\right)$, their
+product is a eVset-automaton $C=\left(Q,q^{0},q^{f},\delta\right)$,
+where:
+\begin{itemize}
+\item $Q=Q_{A}\times Q_{B}$;
+\item $q^{0}=\left\langle q_{A}^{0},q_{B}^{0}\right\rangle $;
+\item $q^{f}=\left\langle q_{A}^{f},q_{B}^{f}\right\rangle $;
+\item $\delta$ has the following transitions:
+
+\begin{itemize}
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,\sigma,\left\langle q'_{A},q'_{B}\right\rangle \right)$
+whenever $\sigma\in\varSigma$, $\left(q_{A},\sigma,q'_{A}\right)\in\delta_{A}$
+and $\left(q_{B},\sigma,q'_{B}\right)\in\delta_{B}$;
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,S_{A}\cup S_{B},\left\langle q'_{A},q'_{B}\right\rangle \right)$
+whenever $\left(q_{A},S_{A},q'_{A}\right)\in\delta_{A}$, $\left(q_{B},S_{B},q'_{B}\right)\in\delta_{B}$
+and $\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{A}=\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{B}$;
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,S_{A},\left\langle q'_{A},q{}_{B}\right\rangle \right)$
+whenever $\left(q_{A},S_{A},q'_{A}\right)\in\delta_{A}$ and $\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{A}=\textrm{Ø}$;
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,S_{B},\left\langle q{}_{A},q'_{B}\right\rangle \right)$
+whenever $\left(q_{B},S_{B},q'_{B}\right)\in\delta_{B}$ and $\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{B}=\textrm{Ø}$.
+\end{itemize}
+\end{itemize}
+We may write $C=A\times B$.
+\end{definition}
+
+\noindent 
+\begin{definition}
+Given two well-behaved operation-closed eVset-automata $A$ and $B$,
+and given an eVset-automaton $C$ such that $C=A\times B$, the projection
+of a transition $t$ in $C$ over $A$ is a transition $t'$ in $A$
+defined as:
+\begin{itemize}
+\item $\left(q_{A},\sigma,q'_{A}\right)$ if $t=\left(\left\langle q_{A},q_{B}\right\rangle ,\sigma,\left\langle q'_{A},q'_{B}\right\rangle \right)$;
+\item $\left(q_{A},S_{A},q'_{A}\right)$ if $t=\left(\left\langle q_{A},q_{B}\right\rangle ,S_{A}\cup S_{B},\left\langle q'_{A},q'_{B}\right\rangle \right)$
+or $t=\left(\left\langle q_{A},q_{B}\right\rangle ,S_{A},\left\langle q'_{A},q{}_{B}\right\rangle \right)$.
+\end{itemize}
+We may write $t'=\pi_{A}\left(t\right)$. The definition is symmetric
+for the projection over $B$.
+\end{definition}
+
+\noindent 
+\begin{definition}
+Given two well-behaved operation-closed eVset-automata $A$ and $B$,
+and given an eVset-automaton $C$ such that $C=A\times B$, the projection
+of a path $p=t_{1},\ldots,t_{n}$ in $C$ over $A$ is a path $p'$
+in $A$ defined as:
+\[
+t'_{1},\ldots t'
+\]
+
+\end{definition}
+
+\noindent 
+\begin{theorem}
+Given two well-behaved operation-closed eVset-automata $A$ and $B$,
+and given an eVset-automaton $C$ such that $C=A\times B$, then $\left\llbracket C\right\rrbracket =\left\llbracket A\right\rrbracket \bowtie\left\llbracket B\right\rrbracket $.\end{theorem}
+
+\begin{proof}
+To show that $\left\llbracket C\right\rrbracket \subseteq\left\llbracket A\right\rrbracket \bowtie\left\llbracket B\right\rrbracket $,
+we can easily decompose a run of $C$ on a string $\mathbf{s}$ into
+two consistent runs of $A$ and $B$. To show that $\left\llbracket A\right\rrbracket \bowtie\left\llbracket B\right\rrbracket \subseteq\left\llbracket C\right\rrbracket $,
+consider a string $\mathbf{s}$, a $\mathbf{s}$-tuple $\mu_{A}\in A\left(\mathbf{s}\right)$
+and a $\mathbf{s}$-tuple $\mu_{B}\in B\left(\mathbf{s}\right)$ that
+agrees with $\mu_{A}$ on common variables. Given the $\mathbf{s}$-tuple
+$\mu$ that yelds all the variable assignments of $\mu_{A}$ and $\mu_{B}$.
+We need to find a run $\rho$ of $C$ on $\mathbf{s}$ that returns
+$\mu$. Let us call $\rho_{A}\in\textrm{ARuns}\left(A,\mathbf{s}\right)$
+and $\rho_{B}\in\textrm{ARuns}\left(B,\mathbf{s}\right)$ the runs
+that return $\mu_{A}$ and $\mu_{B}$, respectively. We can obtain
+$\rho$ by combining $\rho_{A}$ and $\rho_{B}$. To do that, we first
+need to ensure that $\rho_{A}$ and $\rho_{B}$ are consistent on
+the order of the variable operations they perform. Since $A$ and
+$B$ are well-behaved and operation-closed, $\rho_{A}$ and $\rho_{B}$
+can always be selected so that they are consistent.
+\end{proof}
+
+\noindent 
 \begin{lemma}
-\noindent Given an extended well-behaved varable-set automaton $A$,
-$A$ can be converted into an equivalent well-behaved variable-set
-automaton $A'$ in polynomial time.\end{lemma}
+\noindent Given an extended varable-set automaton $A=\left(Q,q_{0},q_{f},\delta\right)$,
+$A$ can be converted into an equivalent variable-set automaton $A'$
+in polynomial time in a well-behavedness preserving manner.\end{lemma}
 
@@
 is unique, and the sequence of operations from $c_{j,1}$ to $c_{j,k}$
-respect the operation ordering that we chose.\end{proof}
+respect the operation ordering that we chose. It can also easily be
+verified that well-behavedness is preserved.
+
+\noindent It remains to show that the construction of $A'$ can be
+carried in polynomial time. Let us refer to the size of $A$ as $n$.
+We consider $n=\left|Q\right|+\left|\delta\right|+\left|\textrm{SVars}\left(A\right)\right|$.
+We can reasonably assume that the operations of adding a new state,
+adding a new transition, removing a state and removing a transition
+take constant time. Thus, expanding a single transition of $A$ takes
+$O\left(\left|\textrm{SVars}\left(A\right)\right|\right)=O\left(n\right)$
+time. Expanding every transition that is needed will then take $O\left(\left|\textrm{SVars}\left(A\right)\right|\cdot\left|\delta\right|\right)=O\left(n^{2}\right)$
+time.
+
+\medskip{}
+\end{proof}
+
+\begin{definition}
+\noindent Given an extended vset-automaton $A$, the epsilon-closure
+of $A$ is an extended vset-automaton $A'=\left(Q',q_{0}',q_{f}',\delta'\right)$
+, where:
+\begin{itemize}
+\item $Q'=Q$;
+\item $q_{0}'=q_{0}$;
+\item $q_{f}'=q_{f}$;
+\item $\delta\subseteq\delta'$;
+\item for every sequence of triples $\left(q_{1},S_{1},q'_{1}\right),\ldots,\left(q_{n},S_{n},q'_{n}\right)\in\delta$
+such that $\forall i$,$1\leq i<n$ we have $q'_{i}=q_{i+1}$, $\delta'$
+contains a triple $\left(q_{1},\bigcup_{i}S_{i},q_{n}\right)$ and
+no additional triples that don't belong to $\delta$.
+\end{itemize}
+\end{definition}
+
+\begin{theorem}
+\noindent Given two extended vset-automata $A$ and $B$ that are
+union-compatible, a third vset-automaton $C$ can be produced in polynomial
+time such that $\left\llbracket C\right\rrbracket =\left\llbracket A\right\rrbracket \cup\left\llbracket B\right\rrbracket $,
+in a well-behavedness preserving manner.\end{theorem}
+
+\begin{proof}
+Let us consider $A$ to be of the form $\left(Q^{A},q_{0}^{A},q_{f}^{A},\delta^{A}\right)$
+and $B$ of the form $\left(Q^{B},q_{0}^{B},q_{f}^{B},\delta^{B}\right)$.
+We can take $C=\left(Q^{C},q_{0}^{C},q_{f}^{C},\delta^{C}\right)$,
+where:
+\begin{itemize}
+\item $Q^{C}=Q^{A}\cup Q^{B}$;
+\item $q_{0}^{C}=q_{0}^{A}$;
+\item $q_{f}^{C}=q_{f}^{A}$;
+\item $\delta^{C}=\delta^{A}\cup\delta^{B}\cup\left\{ \left(q_{0}^{C},\textrm{Ø},q_{0}^{B}\right),\left(q_{f}^{B},\textrm{Ø},q_{f}^{C}\right)\right\} $.
+\end{itemize}
+In this construction, we allow to go from the initial state of $A$
+to the one of $B$, and to go from the final state of $B$ to the
+one of $A$. Thus, given a string $\mathbf{s}$, $C$ can span exactly
+the $\left(V,\mathbf{s}\right)-$tuples contained in $A\left(\mathbf{s}\right)\cup B\left(\mathbf{s}\right)$.
+Regarding the complexity, the operations that we perform here are
+the union of the states sets and transition functions, with the addition
+of two new transitions.
+Let us consider the size of the input as $n=n_{A}+n_{B}=\left|Q^{A}\right|+\left|\delta^{A}\right|+\left|\textrm{SVars}\left(A\right)\right|+\left|Q^{B}\right|+\left|\delta^{B}\right|+\left|\textrm{SVars}\left(B\right)\right|$.
+If we assume the costs of basic operations as adding/deleting a state/transitions
+to be constant, then obtaining $Q^{C}$ takes $O\left(\left|Q^{B}\right|\right)=O\left(n\right)$
+time, and constructing $\delta^{C}$ takes $O\left(\left|\delta^{B}\right|\right)=O\left(n\right)$
+time.
+\end{proof}
+
+\begin{theorem}
+\noindent Given a well-behaved extended vset-automaton $A$, a set
+$X=\textrm{SVars}\left(A\right)$ and a set $Y\subseteq X$, a vset-automaton
+$A'$ can be produced in polynomial time such that $\left\llbracket A'\right\rrbracket =\pi_{Y}\left\llbracket A\right\rrbracket $,
+in a well-behavedness preserving manner.\end{theorem}
+
+\begin{proof}
+Let us consider $A=\left(Q,q_{0},q_{f},\delta\right)$. We can take
+$A'=\left(Q',q_{0}',q_{f}',\delta'\right)$, where:
+\begin{itemize}
+\item $Q'=Q$;
+\item $q_{0}'=q_{0}$;
+\item $q_{f}'=q_{f}$;
+\item $\delta'=\delta\setminus\delta^{\textrm{unprojected}}\cup\delta^{\textrm{projected}}$,
+where:
+
+\begin{itemize}
+\item $\delta^{\textrm{unprojected}}=\left\{ \left(q,S,q'\right)\in\delta\mid S\cap\textrm{SVOps}\left(X\setminus Y\right)\neq\textrm{Ø}\right\} $;
+\item $\delta^{\textrm{projected}}=\left\{ \left(q,S',q'\right)\mid\exists\left(q,S,q'\right)\in\delta^{\textrm{unprojected}}:S'=S\setminus\textrm{SVOps}\left(X\setminus Y\right)\right\} $.
+\end{itemize}
+\end{itemize}
+This construction removes all the occurencies of variable operations
+that concern variables excluded from the projection in the transitions
+of $A$. Since $A$ is well behaved, we have the guarantee that $A'=A$.
+If this wasn't the case, our construction might result in $A\subseteq A'$.
+$A'$ is still well-behaved, since we maintain the occurrencies of
+operations concerning the variables on which we project, that continue
+to appear exactly once on each path. With the usual assumptions, this
+construction can be carried out in $O\left(n\right)$.
+\end{proof}
+
+\begin{definition}
+An AQL core fragment is an AQL fragment that can be associated with
+a core spanner.
+\end{definition}
+
+\begin{theorem}
+Given a regex formula $\gamma$, $\gamma$ can be converted into an
+equivalent, well-behaved extended vset-automaton $A$ in polynomial
+time.\end{theorem}
 

### 2017-03-02 paper/paper.tex
testing conflicts
Updating the repository.

@@
-%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
+%% LyX 2.2.0 created this file.  For more info, see http://www.lyx.org/.
 %% Do not edit unless you really know what you are doing.
@@
 \usepackage{url}
+\usepackage{pdfpages}
 \usepackage{slashed}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
+%% Because html converters don't know tabularnewline
+\providecommand{\tabularnewline}{\\}
+%% A simple dot to overcome graphicx limitations
+\newcommand{\lyxdot}{.}
+
 \floatstyle{ruled}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
+\usepackage{tikz, pgf, pgfplots}
+\usetikzlibrary{automata,graphs,positioning}
+\usetikzlibrary{arrows,backgrounds,plotmarks}
+\pgfplotsset{width=7cm,compat=newest}
+\pgfplotsset{plot coordinates/math parser=false}
+
+\@ifundefined{showcaptionsetup}{}{%
+ \PassOptionsToPackage{caption=false}{subfig}}
+\usepackage{subfig}
 \makeatother
@@
 \begin{document}
-\selectlanguage{english}%
-
-\title{Engineering a Runtime System for AQL}
-
-\selectlanguage{american}%
-
-\author{Andrea Morciano}
-
-\degreeaward{Docteur en XXX\\
-(Département XXX, Faculté XXX)}
-\maketitle
+\includepdf{PageDeGarde}
 \begin{dedication}
-To my parents.\end{dedication}
+To my parents.
+\end{dedication}
 \begin{acknowledgments}
@@
 \chapter{Information Extraction}
+
+In this chapter, I give a generic introduction to Information Extraction.
+This Chapter is a re-elaboration of \cite{Sarawagi:2008:IE:1498844.1498845}.
+Please refer to the original source for a more comprehensive and detailed
+overview. The figures and examples appearing in this Chapter are mine.
 
@@
 Information Retrieval, Database, Web, Document Analysis. The interest
-in towards it is motivated by the evergrowing amount of data that
-is generated by our society, that is at a point in which manual analysis
-is infeasible. Thus IE promises to bring value to many application
-domains, most notably the enterprise world and the Web. There are
-two main approaches to IE: \emph{rule-based }and \emph{statistical}.
+towards it is motivated by the constantly increasing amount of data
+that is generated by our society. Manual analysis has become infeasible.
+IE promises to bring value to many application domains, most notably
+the enterprise world and the Web.
+
+\medskip{}
+
+There are two prominent approaches to IE: 
+\begin{description}
+\item [{\emph{rule-based}}] which define a set of rules that the output
+has to respect. The rules can be manually coded or learnt from examples;
+\item [{\emph{statistical}}] which seek to identify a decomposition of
+unstructured text and to label its components.
+\end{description}
 
@@
 Extracting structured forms from flat data strings (containing, e.g.,
-addresses). It allows more effective decuplication of information,
+addresses). It allows more effective deduplication of information,
 among the other things.
@@
 
-Articles, conferences sites, individual research sites and similar
+Articles, conference sites, individual research sites and similar
 are explored to obain formatted citations of publications, later stored
 in publicly accessible databases. They are capable to forward references
-and may provide aggregate statistics and scoring information. See
-for instance Google Scholar (\url{https://scholar.google.it/}).
+and may provide aggregate statistics and scoring information. See,
+for instance, Google Scholar (\url{https://scholar.google.com/}).
 
@@
 \section{Issues}
+
+The main issues that have to be dealt with when performing IE can
+be divided in two categories: \emph{accuracy} and \emph{running time}.
 
@@
 \end{itemize}
-While we are able to reach a very good level of accuracy for entity
+While it is possible to reach a very good level of accuracy for entity
 extraction ($90\%$), relationship extraction is still quite unreliable
-($50-70\%$ accuracy), mainly due to its intrinsic complexity, superior
-to that of the task of entity extraction.
+($50-70\%$ accuracy), mainly due to its intrinsic complexity.
 
@@
 
-In its initial phase IE was interesing only to research communities.
-With the enormous amount of data they need to process, companies are
-now concerned with IE too, and \emph{scalability} has become a central
-issue, while many IE systems don't address the problem of efficiently
-carrying out extraction tasks with sufficient attention. Efficiency
-is influenced most notably by the efficiency of the following tasks:
+In its initial phase IE was interesting only to research communities.
+Nowadays companies, because of the enormous amount of data they need
+to process on a regular basis, are concerned with IE too, and \emph{scalability}
+has become a central issue, while many IE systems don't address the
+problem of efficiently carrying out extraction tasks with sufficient
+attention. Efficiency is influenced most notably by the efficiency
+of the following tasks:
 \begin{itemize}
@@
 Recently many solutions have been proposed to target the scalability
-issue; one of these is indeed SystemT.
+issue; one of these is SystemT.
 
@@
 text document (that becomes \emph{annotated}), although there are
-other possibilities.
+other possible output formats.
 
@@
 \end{equation}
-\\
 Contextual patterns are a way to identify entitities by exploiting
@@
 have been identified, and might consist in storing them in a database
-or adding delimiters directly in the text.\\
-Most systems in this category present a cascaded structure: an input
-document goes through a series of processing phases, and the output
-of a phase is the input of the next one. A famous example is that
-of cascading grammars: formal grammars are evaluated in sequence on
-the input.
+or adding delimiters directly in the text.
+
+Most systems in this category present a \emph{cascaded structure}:
+an input document goes through a series of processing phases, and
+the output of a phase is the input of the next one. A famous example
+is that of \emph{cascading grammars}: formal grammars are evaluated
+in sequence on the input.
 
@@
 
-For the examples I use an idealized syntax, as in REF.
+For the examples I use an idealized syntax, as in \cite{Sarawagi:2008:IE:1498844.1498845}.
 \begin{example}
@@
 \end{multline}
-
 \end{example}
 
-a
 \begin{example}
@@
 \end{multline}
-
 \end{example}
@@
 
-\subsubsection{Token-Level Models\label{sub:Token-Level-Models}}
+\subsubsection{Token-Level Models\label{subsec:Token-Level-Models}}
 
@@
 
-Clues are of the form 
+The clues that these models exploit are of the form 
 \begin{equation}
-f:\left(x,y,i\right)\longmapsto\mathbb{R}
+f:\left(\mathbf{x},y,i\right)\longmapsto\mathbb{R}
 \end{equation}
-where $x$ is a sequence of tokens, $i$ is a position in $x$ and
-$y$ is a candidate label for the token at $i$. We distinguish between
-these types of features:
+where $\mathbf{x}$ is a sequence of tokens, $i$ is a position in
+$\mathbf{x}$ and $y$ is a candidate label for the token at $i$.
+We distinguish between these types of features:
 \begin{itemize}
@@
 \end{itemize}
-
-\paragraph{Models for Labeling Token}
+\begin{figure}
+Yesterday I watched a movie called ``The Matrix''
+
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
+\hline 
+i & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\tabularnewline
+\hline 
+\hline 
+x & Yesterday & I & watched & a & movie & called & `` & The & Matrix & ``\tabularnewline
+\hline 
+y & $y_{1}$ & $y_{2}$ & $y_{3}$ & $y_{4}$ & $y_{5}$ & $y_{6}$ & $y_{7}$ & $y_{8}$ & $y_{9}$ & $y_{10}$\tabularnewline
+\hline 
+\end{tabular}
+
+\caption{\label{fig:tokenization}Decomposition of a sentence into a sequence
+of tokens.}
+\end{figure}
+
+\begin{example}
+Consider the sentence shown in Figure \ref{fig:tokenization} and
+its corresponding decomposition. An example of word feature at position
+5 is
+
+\[
+f_{1}\left(y,\mathbf{x},i\right)=\left\llbracket x_{i}\textrm{ equals "Matrix"}\right\rrbracket \cdot\left\llbracket y=\textrm{Movie}\right\rrbracket 
+\]
+where $\left\llbracket P\right\rrbracket =1$ if predicate $P$ is
+true, and 0 otherwise. An orthographic feature might be
+
+\[
+f_{2}\left(y,\mathbf{x},i\right)=\left\llbracket x_{i}x_{i+1}\textrm{ matches INITIAL\_QUOTE CapsWord}\right\rrbracket \cdot\left\llbracket y=\textrm{Movie}\right\rrbracket .
+\]
+
+Finally, an example of dictionary lookup feature is
+
+\[
+f_{3}\left(y,\mathbf{x},i\right)=\left\llbracket x_{i}\textrm{ in Movie\_dictionary}\right\rrbracket \cdot\left\llbracket y=\textrm{Movie}\right\rrbracket .
+\]
+\end{example}
+
+\paragraph{Models for Labeling Tokens}
 
@@
 In these models the label for a segment depends on the properties
-of its token and on the previous segment. We can describe a segment
+of its tokens and on the previous segment. We can describe a segment
 $s_{j}$ as:
@@
 \begin{equation}
-f\left(y_{i},y_{i-1},x,l_{j},u_{j}\right)
+f\left(y_{i},y_{i-1},\mathbf{x},l_{j},u_{j}\right)
 \end{equation}
-where $x$ is the input sequence of tokens, $y_{i-1}$ is the proposed
-label for the previous segment and the other symbols are defined as
-for Equation \ref{eq:segment}. Besides token-level features (see
-subsection \ref{sub:Token-Level-Models}), we can exploit the following
-feature types:
+where $\mathbf{x}$ is the input sequence of tokens, $y_{i-1}$ is
+the proposed label for the previous segment and the other symbols
+are defined as for Equation \ref{eq:segment}. Besides token-level
+features (see Subsection \ref{subsec:Token-Level-Models}), we can
+exploit the following feature types:
 \begin{itemize}
@@
 \end{itemize}
-There are also global segmentation models, that try to find the best
-segmentation of a token sequence by maximizing a target function.
+\begin{figure}
+``The Matrix'' by The Wachowski Brothers
+
+\begin{tabular}{|c|c|c|c|}
+\hline 
+i & %
+\begin{tabular}{cccc}
+1 & 2 & 3 & 4\tabularnewline
+\end{tabular} & 5 & %
+\begin{tabular}{ccc}
+6 & 7 & 8\tabularnewline
+\end{tabular}\tabularnewline
+\hline 
+\hline 
+x & ``The Matrix'' & by & The Wachowski Brothers\tabularnewline
+\hline 
+$\left(l_{j},u_{j},y_{i}\right)$ & $1,4,M$ & $5,5,O$ & $6,8,D$\tabularnewline
+\hline 
+\end{tabular}
+
+\caption{\label{fig:segmentation2}Segmentation of a sentence.}
+\end{figure}
+
+\begin{example}
+Consider the sentence in Figure \ref{fig:segmentation2} and its corresponding
+segmentation, which identifies a movie and a director. A basic feature
+similarity feature for the director could be:
+
+\[
+f\left(y_{i},y_{i-1},\mathbf{x},6,8\right)=\left\llbracket x_{6}x_{7}x_{8}\textrm{ appears in a list of movie directors}\right\rrbracket \cdot\left\llbracket y_{i}=\textrm{Director}\right\rrbracket .
+\]
+
+A more realistic feature would make use of some similarity function,
+rather than requiring an extact match. An example of length feature
+is
+
+\[
+f\left(y_{i},y_{i-1},\mathbf{x},l,u\right)=\left\llbracket u-l=3\right\rrbracket \cdot\left\llbracket y_{i}=\textrm{Director}\right\rrbracket .
+\]
+
+\medskip{}
+\end{example}
+
+There exist also global segmentation models, that try to find the
+best segmentation of a token sequence by maximizing a target function.
 
@@
 
-When exctracting relationship between entities, we might face three
+When exctracting relationships between entities, we might face three
 types of specific tasks:
-\begin{itemize}
+\begin{enumerate}
 \item given a pair of entities, find the relationship between them;
@@
 pairs of entities $e',e''$ such that $\left(e',e''\right)\in r$.
-\end{itemize}
-
-\subsection{Predicting the Relationship between a Entity Pair }
+\end{enumerate}
+
+\subsection{Predicting the Relationship between an a Pair of Entities}
 
@@
 \end{itemize}
-The second task is a special case of the third one so I don't treat
-it directly.
+The second task is a special case of the third, and won't be treated
+here.
 
@@
 way of evaluating AQL queries, that uses a formalism derived from
-\emph{Finite State Automata}. Producing and analysing a concrete runtime
-system for AQL queries that employs this new method is the object
-of my thesis. Since SystemT is a rule-based system, I now focus only
-on this category.
+\emph{Finite State Automata}. This chapter is a re-elaboration from
+\cite{4497502,Krishnamurthy:2009:SSD:1519103.1519105,Fagin:2015:DSF:2772377.2699442}.
+In particular, the mathematical statements from \ref{first-stat}
+to \ref{.last-stat} are taken or adapted from \cite{Fagin:2015:DSF:2772377.2699442},
+except for Remarks \ref{first-rem} to \ref{last-rem}, that come
+from \cite{4497502}. Please refer to the original sources for more
+details on the topic. The figures and examples appearing in this chapter
+are mine, except when I explicitly state otherwise.
 
@@
 
-As I mentioned in the previous chapter, scalability has now become
-a central issue to IE. Companies now rely on IE for many tasks; a
-prominent example is \emph{Business Intelligence}. Unfortunately,
-many systems developed in the past don't address this concern correctly.
-Traditionally, most rule-based system rely on cascading grammars:
-formal grammars are executed in sequence on the input, each grammar
-representing a stage that takes as input the ouput of the previous
-one. Rules in such grammars are matched using regular expressions:
-if a part of the input text satisfies the regular expression associated
-with a rule, that rule is activated (i.e. ``fires'') and the corresponding
-action is executed. Evaluating a grammar on a document tends to be
-costly, because simply evaluating a single rule might require scanning
-the whole document. On large datasets, the running time becomes enormous.
-What's more, these kind of systems are not able to fulfill the expressivity
-requirements of the complex extraction tasks.\\
+As I mentioned in the previous chapter, scalability is now a main
+concern in IE. Companies rely on IE for many tasks; a prominent example
+is \emph{Business Intelligence}. Unfortunately, many systems developed
+in the past don't address this issue correctly. Traditionally, most
+rule-based system rely on cascading grammars: formal grammars are
+executed in sequence on the input, each grammar representing a stage
+that takes as input the ouput of the previous one. Rules in such grammars
+are matched using regular expressions: if a part of the input text
+satisfies the regular expression associated with a rule, that rule
+is activated (i.e. ``fires'') and the corresponding action is executed.
+Evaluating a grammar on a document tends to be costly, because simply
+evaluating a single rule might require scanning the whole document.
+On large datasets, the running time becomes enormous. What's more,
+these kind of systems are not able to fulfill the expressivity requirements
+of the complex extraction tasks.
+
 SystemT was developed at IBM to address these issues. This system
@@
 difficult to resolve in cascading grammars, where they tipically require
-disambiguation policies that only partially solve the problem.\\
+disambiguation policies that only partially solve the problem.
+
 I now give an overview of SystemT's high-level structure. Then, I
-talk about its data model, execution model and algebra.
+talk about its data model, execution model and algebra. For more information
+on SystemT and AQL, please refer to \cite{4497502,Krishnamurthy:2009:SSD:1519103.1519105}.
 
@@
 
-\subsection{Data Model\label{sub:Data-Model}}
+\subsection{Data Model\label{subsec:Data-Model}}
 
 SystemT uses an \emph{object-relational} data model for annotations
-on a text document, that allows applying logical operators over it,
-that can be composed. There is an important assumption to mention
+on a text document, that allows applying logical operators over it.
+Operators can be composed. There is an important assumption to mention
 before continuing: an extraction task over a set of documents is performed
-one document at a time. This means that any relationships between
+on one document at a time. This means that any relationships between
 entities in different documents are disregarded. This assumption is
@@
 (think of the Web), still a large number of relevant tasks can be
-carried out this way. In the following, I define the basic data types
-of SystemT's data model. \\
-I consider a finite alphabet $\varSigma$ of symbols (characters).
-$\sum^{*}$ denotes the set of all strings of finite length over $\sum^{*}$.
-A document is modeled as one such string. 
-\begin{definition}
-A document is a string $s\in\sum^{*}$.
+carried out this way. In the following, I report the definitions of
+the basic data types in the data model of SystemT.
+
+\medskip{}
+
+I consider a finite alphabet $\Sigma$ of symbols (characters). $\Sigma^{*}$
+denotes the set of all strings of finite length over $\Sigma$. A
+document is modeled as one such string. 
+\begin{definition}
+\label{first-stat}A document is a string $\mathbf{s}\in\Sigma^{*}$.
 \end{definition}
@@
 \begin{definition}
-Given a string $s=\sigma_{1}...\sigma_{n}$ where $\forall i$, $\sigma_{i}\in\Sigma$,
-with length$\left|s\right|=n$ and whose characters are indexed in
-the natural way, a span is a substring $\left[i,j\right\rangle $
-, where $i,j$ are indices of $s$ satisfying $1\leq i\leq j\leq n+1$.
-A span of $s$ is also denoted as $s_{\left[i,j\right\rangle }$.
+Given a string $\mathbf{s}=\sigma_{1}...\sigma_{n}$ where $\forall i$,
+$\sigma_{i}\in\Sigma$, with length$\left|\mathbf{s}\right|=n$ and
+whose characters are indexed in the natural way, a span is a substring
+$\left[i,j\right\rangle $, where $i$, $j$ are indices of $s$ satisfying
+$1\leq i\leq j\leq n+1$. A span of $\mathbf{s}$ is also denoted
+as $\mathbf{s}_{\left[i,j\right\rangle }$.
 \end{definition}
@@
 sequences of spans\emph{. }Let us denote the set of all possible spans
-of a string $s$ by $\textrm{Spans}\left(s\right)$ and by $\textrm{SVars}$
-an infinite set of span variables, which can be assigned spans. A
-$\left(V,s\right)-\textrm{tuple}$ is defined as follows:
+of a string $\mathbf{s}$ by $\textrm{Spans}\left(\mathbf{s}\right)$
+and by $\textrm{SVars}$ an infinite set of span variables, which
+can be assigned spans. A $\left(V,\mathbf{s}\right)-\textrm{tuple}$
+is defined as follows:
 \begin{definition}
 Given a set $V\subseteq\textrm{SVars}$, $V$ being finite, and a
-string $s\in\sum^{*}$, a $\left(V,s\right)-\textrm{tuple}$ is a
-mapping $\mu:V\longmapsto\textrm{Spans}\left(s\right)$. When $V$
-is clear from the context, we might call $\mu$ simply a $s-\textrm{tuple}$.
-\end{definition}
-
-\noindent A set of tuples is called a \emph{relation. }Here, I focus
-on \emph{span relations}, which I formally define as $\left(V,s\right)-\textrm{relations}$.
-\begin{definition}
-\label{span-relation}A $\left(V,s\right)-\textrm{relation}$ is a
-set of $\left(V,s\right)-\textrm{tuples}$. As before, we can speak
-of $s-\textrm{relation}$ when $V$ is clear.
-\end{definition}
-
-\noindent Each operator in the algebra takes one or more span relations
-as input and outputs a single span relation.
-
-\subsection{Execution Model\label{sub:Execution-Model}}
+string $\mathbf{s}\in\sum^{*}$, a $\left(V,\mathbf{s}\right)-\textrm{tuple}$
+is a mapping $\mu:V\longmapsto\textrm{Spans}\left(\mathbf{s}\right)$.
+When $V$ is clear from the context, we might call $\mu$ simply a
+$\mathbf{s}-\textrm{tuple}$.
+\end{definition}
+
+\noindent A set of tuples is called a \emph{relation. }Here, we focus
+on \emph{span relations}, which are formally defined as $\left(V,\mathbf{s}\right)-\textrm{relations}$.
+\begin{definition}
+\label{span-relation}A $\left(V,\mathbf{s}\right)-\textrm{relation}$
+is a set of $\left(V,\mathbf{s}\right)-\textrm{tuples}$. As before,
+we can speak of $\mathbf{s}-\textrm{relation}$ when $V$ is clear.
+\end{definition}
+
+\noindent Each operator in the algebra of SystemT takes one or more
+span relations as input and outputs a single span relation.
+
+\subsection{Execution Model\label{subsec:Execution-Model}}
 
@@
 \caption{\label{alg:Annotating-all-local}Annotating all local databases in
-a global database (taken from REF).}
+a global database (taken from \cite{4497502}).}
 
@@
 
-\noindent In this procedure, step 2 in the for loop is the most expensive.
-
-\subsection{Algebra of Operators \label{sub:Algebra-of-Operators}}
+\subsection{Algebra of Operators \label{subsec:Algebra-of-Operators}}
 
@@
 \item \emph{Cartesian product }($\times$);
+\item Join ($\bowtie$);
 \item \emph{Union} ($\cup$);
@@
 Loosely speaking, span extraction operators take a pattern and a document
-as input and output the maximal set of spans that match that pattern.
+as input and output a maximal set of spans that match that pattern.
 There are two main span extraction operators:
@@
 $s_{2}$;
+\item [{$s_{1}\subseteq s_{2}$}] when $s_{1}$ is contained in or equals
+$s_{2}$;
 \item [{$s_{1}=s_{2}$}] when $s_{1}$ equals $s_{2}$.
 \end{lyxlist}
+\begin{figure}
+{\tiny{}``Yesterday I }\textbf{\tiny{}watched 'Annie Hall'}{\tiny{}.
+It is about the relationship between a TV writer and his girlfriend,
+who wants to become an actress. I really }\textbf{\tiny{}loved this
+movie}{\tiny{}. Although the }\textbf{\tiny{}acting wasn't very good
+}{\tiny{}sometimes, the }\textbf{\tiny{}dialogues were smart}{\tiny{}.
+I'd recommend it to anyone.''}{\tiny \par}
+
+\centering
+
+\subfloat[\label{fig:high-level}High-level structure.]{\tikz[level distance=2cm]  	\node[rectangle, draw, align=center] {ReviewInstance \\ Aggregator}  		child { node[rectangle, draw, align=center] {ReviewPart \\ Extractor}  			edge from parent [<-] node[right=1.6cm, pos=1.5, draw=none, font=\tiny, align=left] {"watched 'Annie Hall'" \\ "loved this movie" \\ "acting wasn't very good" \\ "dialogues were smart"} 			edge from parent [<-] node[right=1.6cm, pos=-0.7, draw=none, font=\tiny, align=left] {"\textbf{watched 'Annie Hall'}.
+\\ It is about...
+\textbf{loved this} \\ \textbf{movie}.
+Altough the \textbf{acting} \\ \textbf{wasn't very good}...
+\\ \textbf{dialogues were smart}"}  		};
+
+}
+
+\subfloat[\label{fig:tree}Operator tree.]{\tikz [level distance=1.6cm, level 3/.style={sibling distance=6em}, level 4/.style={sibling distance=2em}] 	\node {$\varOmega_{c}$} 			child { node {$\beta$} 				child { node {$\cup$} 					child { node {$\bowtie_{\preceq_{10}}$} 						child { node {$\varepsilon_{d}$} edge from parent [<-] node[left,draw=none, font=\tiny] {Action} } 						child { node {$\varepsilon_{re}$} edge from parent [<-] node[right,draw=none, font=\tiny] {Title} } 						edge from parent [<-] node[left, draw=none, font=\tiny] {ReviewPart} 					} 					child { node {$\bowtie_{\preceq_{30}}$} 						child { node {$\varepsilon_{d}$} edge from parent [<-] node[left,draw=none, font=\tiny] {Aspect} } 						child { node {$\varepsilon_{d}$} edge from parent [<-] node[right,draw=none, font=\tiny] {Attribute} } 						edge from parent [<-] node[right, draw=none, font=\tiny] {ReviewPart} 					} 					child { node {.
+.
+.
+.} edge from parent[draw=none]} 					child { node {$\bowtie_{\preceq_{30}}$} 						child { node {$\varepsilon_{d}$}  edge from parent [<-] node[left,draw=none, font=\tiny] {Sentiment} } 						child { node {$\varepsilon_{d}$}  edge from parent [<-] node[right,draw=none, font=\tiny] {MovieKeyword} } 						edge from parent [<-] node[right, draw=none, font=\tiny] {ReviewPart} 					} 					edge from parent [<-] node[left, draw=none, font=\tiny] {ReviewPart} 				} 				edge from parent [<-] node[left, draw=none, font=\tiny] {ReviewInstance} 			};
+
+}
+
+\caption{\label{fig:query1} Extraction of informal movie reviews.}
+
+\end{figure}
+
+\medskip{}
+
+\begin{example}
+Consider the task of extracting informal movie reviews from blog posts.
+A possible annotator to perform the task is shown in Figure \ref{fig:query1}.
+Figure \ref{fig:high-level} shows the high-level structure of the
+annotator. There are two main components: the \textbf{ReviewPart}
+module, which extracts the text snippets that are identified as parts
+of a movie review, and the \textbf{ReviewInstance} module that aggregates
+adjacent snippets into single blocks of text. Figure \ref{fig:tree}shows
+the operator tree of the annotator. Review parts are extracted by
+a series of join operators that combine elements of basic relations,
+extracted by running span extractors on the text. Subsequently, the
+block operator aggregates the parts, and review instances that are
+contained into others are discarded.
+\end{example}
 
@@
 
-As I mentioned in the introduction of this chapter, SystemT can make
+As I mentioned in the introduction of this Chapter, SystemT can make
 use of the optimization techniques from traditional database systems,
@@
 \begin{remark}
-The effect of document-at-a-time processing is that the span relations
-produced and consumed for a single document by operators are very
-small in size and often empty.\label{small-relations}
+\label{first-rem}The effect of document-at-a-time processing is that
+the span relations produced and consumed for a single document by
+operators are very small in size and often empty.\label{small-relations}
 \end{remark}
@@
 \begin{remark}
-Evaluating an annotator on a large set of document is a \emph{CPU-intensive
+Evaluating an annotator on a large set of documents is a \emph{CPU-intensive
 process}. This is because the running time is by far dominated by
@@
 \begin{remark}
-Spans are nothing but \emph{intervals}, so we can exploit \emph{interval
-algebra}.\label{interval-algebra}
+\label{last-rem}Spans are nothing but \emph{intervals}, so we can
+exploit \emph{interval algebra}.\label{interval-algebra}
 \end{remark}
@@
 \noindent With Remark \ref{small-relations} and Remark \ref{interval-algebra}
-in mind, running time can be reduced in a number of ways. I present
-them briefly.
+in mind, running time can be reduced in a number of ways. Let us look
+at them briefly.
 \begin{description}
 \item [{Regular~Expression~Strength~Reduction}] Some classes of regular
-expressions, as defined in the POSIX standard, are executed more efficiently
-by using \emph{specialized engines }that are able to improve performance.
+expressions, as defined in the POSIX standard, can be executed by
+using \emph{specialized engines, }that are able to improve performance.
 For example, an expression that looks for a finite number of strings
@@
 expensive operations on selected regions of a document. It is used
-for \emph{join} operators involving dictionary matching operators
-and/or a regular expression matching operators in their argments.
-One input operator to a join is evaluated on the whole document, while
-the other operators are modified to consider only some neighborhood
-of the results from the first one. These neighborhoods are established
+for \emph{join} operators involving dictionary a matching operator
+in their arguments. One input operator is evaluated on the whole document,
+while the others are modified to consider only some neighborhood of
+the results from the first one. These neighborhoods are established
 by some ad hoc heuristic.
@@
 formulated. In the end, the best ones are retained. Results of sample
-experiments on a large dataset are decribed in REF and REF.
+experiments on a large dataset are decribed in \cite{4497502,Krishnamurthy:2009:SSD:1519103.1519105}.
 
@@
 AQL is the concrete language used by SystemT to express annotators,
-designed to support the execution model I described in Subsection
-\ref{sub:Execution-Model}. It is a declarative language with a syntax
-very similar to that of SQL. It supports all the operators that I
-presented in Subsection \ref{sub:Algebra-of-Operators}. When coding
-an extraction task in AQL, we are able to build a series of views
-of a document, that are of progressively higher level, and can be
-based on lower-level views. The content of a view corresponds with
-annotations in the text. An input document is modeled as a view too,
-which is provided by default. The main advantages of AQL are:
-\begin{itemize}
-\item it allows formulating \emph{complex low-level }patterns in a declarative
+designed to support the execution model described in Subsection \ref{subsec:Execution-Model}.
+It is a \emph{declarative language} with a syntax very similar to
+that of SQL. It supports all the operators presented in Subsection
+\ref{subsec:Algebra-of-Operators}. When coding an extraction task
+in AQL, we can build a series of views of a document, of progressively
+higher abstraction level. Higher-level views are tipically based on
+lower-level views. The content of a view corresponds with annotations
+in the text. An input document is modeled as a view too, which is
+provided by default. The main advantages of AQL are:
+\begin{itemize}
+\item it allows formulating complex low-level patterns in a declarative
 fashion;
 \item it enables \emph{modularization }and \emph{reuse} of the queries,
-making development and maintance of \emph{complex high-level structures}
+making development and maintance of complex high-level structures
 easier.
 \end{itemize}
-
-\section{A Formalism for AQL's Core: VStack and VSet Automata}
-
-AUTHORS describe in REF a formal model that captures the core functionality
-of AQL. They show a way to represent annotators expressed in AQL by
-means of modified finite state automata, namely \emph{VStack} and
-\emph{VSet automata}. In my work, I define and implement a runtime
-system based on this new formalism. I now describe the model and discuss
-the relative expressive power of its main elements. First of all,
-we need to be more precise on some basic concepts.
+\begin{figure}
+\centering
+
+\tikz[level distance=2cm]  	\node[rectangle, draw, align=left] {--\emph{Define a dictionary of actions} \\ \textbf{create} dictionary Action as ('saw', 'watched', ...
+); \\ \\ --\emph{Use a regular expression to find movie titles} \\ \textbf{create view} Title \textbf{as} \\ \textbf{extract} regex /["'][A-Z]$\textbackslash$w+(s+$\textbackslash$w+)\{0,3\}["']/ \\ \ \ \ \textbf{on} D.text \\ \ \ \ \textbf{as} title \\ \textbf{from} Document D; \\ \\ --\emph{A single ReviewPart rule.
+It looks for instances} \\ --\emph{of Action followed within 10 characters by a title} \\ \textbf{create view} ReviewPart \textbf{as} \\ \textbf{select} CombineSpans(A.act, T.title) \textbf{as} part \\ \textbf{from} \\  \ \ \ (\textbf{extract} dictionary 'Action' \textbf{on} D.text \textbf{as} act \\ \ \ \ \ \textbf{from} Document D) A, Title T \\ \textbf{where} \\ \ \ \ Follows(A.act, T.title, 0, 10) \\ \textbf{consolidate on} CombineSpans(A.act, T.title);};
+
+\caption{\label{fig:AQL}Parts of the annotator from Figure \ref{fig:query1}
+expressed in AQL.}
+\end{figure}
+
+\begin{example}
+The AQL code in Figure \ref{fig:AQL} realizes the first join from
+the left in Figure \ref{fig:tree}.
+\end{example}
+
+\section{A Formalism for AQL's Core: Variable Stack Automata and Variable
+Set Automata}
+
+In \cite{Fagin:2015:DSF:2772377.2699442}, a formal model capturing
+the core functionality of AQL is described. The authors show a way
+to represent annotators expressed in AQL by means of modified finite
+state automata, namely \emph{variable stack automata} and \emph{variable
+set automata}. I now resume the formal model, and look at the relative
+expressive power of its main elements.
 
@@
 I consider $\varSigma$ and $\varSigma^{\ast}$, defined as in Subsection
-\ref{sub:Data-Model}. Let us start with the definition of a language.
+\ref{subsec:Data-Model}. Let us start with the definition of a language.
 \begin{definition}
@@
 
-\noindent Now we can state formally define regular expressions, by
-giving their language.
-\begin{definition}
-Regular expressions over $\varSigma$ are those that belong to the
-language $\gamma$ defined as:
+\noindent An important concept is that of regular expressions, which
+can be defined by describing their language.
+\begin{definition}
+Regular expressions over $\varSigma$ are the strings that belong
+to the language $\gamma$ defined as:
 \begin{equation}
@@
 \noindent where:
+
 \begin{itemize}
@@
 \item $\cdot$ is the concatenation operator;
-\item $\ast$ is the Kleene Closure.
+\item $\ast$ is the Kleene Star operator.
 \end{itemize}
@@
 $\sigma_{1}\vee...\vee\sigma_{n}$ (with a slight abuse of notation).
-The language $\mathcal{L}\left(\gamma\right)$of a regular expression
+The language $\mathcal{L}\left(\gamma\right)$ of a regular expression
 $\gamma$ is the set of strings $\mathbf{s}$ over $\varSigma$ that
@@
 
-\noindent I have already defined span relations (Definition \ref{span-relation}).
-They allow us to give a precise definition of annotators, by introducing
+\noindent I have already mentioned span relations (Definition \ref{span-relation}).
+They allow giving a precise definition of annotators, by introducing
 the concept of \emph{document spanner}.
 \begin{definition}
-Given a string $s$, a document spanner P is a function that maps
-$s$ to a $\left(V,s\right)-\textrm{relation}$ $r$ , where $V:=\textrm{Svars}\left(P\right)$.
-We have $r=P\left(s\right).$ We say $P$ is n-ary if $\left|V\right|=n$.
-\end{definition}
-
-\noindent I now present two types of spanners that will be useful
-later on. The first one is that of \emph{Hierarchical Spanners}. First,
-let us see when a $s-\textrm{tuple}$ is hierarchical.
-\begin{definition}
-\label{hierarchical-tuple}Given a string $s$, a document spanner
-P and an $s-\textrm{tuple}$ $\mu\in P\left(s\right)$, $\mu$is hiearchical
-if, for every $x,y\in\textrm{SVars}\left(P\right)$, one of the following
-conditions holds:
+Given a string $\mathbf{s}$, a document spanner P is a function that
+maps $\mathbf{s}$ to a $\left(V,s\right)-\textrm{relation}$ $r$
+, where $V:=\textrm{Svars}\left(P\right)$. We have $r=P\left(\mathbf{s}\right).$
+We say $P$ is n-ary if $\left|V\right|=n$.
+
+\begin{figure}
+\centering
+
+\textbf{String s}
+
+\noindent\resizebox{\textwidth}{!}{
+
+\begin{tabular}{ccccccccccccccccccccccccccccccccccccccccccccccc}
+\hline 
+\noalign{\vskip\doublerulesep}
+\textbf{I} & \textbf{\_} & \textbf{w} & \textbf{a} & \textbf{t} & \textbf{c} & \textbf{h} & \textbf{e} & \textbf{d} & \textbf{\_} & \textbf{``} & \textbf{T} & \textbf{h} & \textbf{e} & \textbf{\_} & \textbf{M} & \textbf{a} & \textbf{t} & \textbf{r} & \textbf{i} & \textbf{x} & \textbf{``} & \textbf{\_} & \textbf{t} & \textbf{h} & \textbf{e} & \textbf{n} & \textbf{\_} & \textbf{w} & \textbf{e} & \textbf{\_} & \textbf{s} & \textbf{a} & \textbf{w} & \textbf{\_} & \textbf{``} & \textbf{A} & \textbf{n} & \textbf{n} & \textbf{i} & \textbf{e} & \textbf{\_} & \textbf{H} & \textbf{a} & \textbf{l} & \textbf{l} & \textbf{``}\tabularnewline[\doublerulesep]
+\hline 
+\noalign{\vskip\doublerulesep}
+1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 & 28 & 29 & 30 & 31 & 32 & 33 & 34 & 35 & 36 & 37 & 38 & 39 & 40 & 41 & 42 & 43 & 44 & 45 & 46 & 47\tabularnewline
+\end{tabular}
+
+}
+
+\medskip{}
+
+$P\left(\mathbf{s}\right)$
+
+\begin{tabular}{|c|c|c|c|}
+\cline{2-4} 
+\multicolumn{1}{c|}{} & \multicolumn{1}{c}{{\small{}$x$}} & \multicolumn{1}{c}{{\small{}$y$}} & {\small{}$z$}\tabularnewline
+\hline 
+{\small{}$\mu_{1}$} & {\small{}$\left[3,10\right\rangle $} & {\small{}$\left[11,23\right\rangle $} & {\small{}$\left[3,23\right\rangle $}\tabularnewline
+\hline 
+{\small{}$\mu_{2}$} & {\small{}$\left[32,35\right\rangle $} & {\small{}$\left[36,48\right\rangle $} & {\small{}$\left[33,48\right\rangle $}\tabularnewline
+\hline 
+{\small{}$\mu_{3}$} & {\small{}$\left[3,10\right\rangle $} & {\small{}$\left[36,48\right\rangle $} & {\small{}$\left[3,48\right\rangle $}\tabularnewline
+\hline 
+\end{tabular}
+
+\caption{\label{fig:running-ex}A string $\mathbf{s}$ and the span relation
+obtained by applying spanner $P$ to $\mathbf{s}$.}
+
+\end{figure}
+\end{definition}
+
+\begin{example}
+In Figure \ref{fig:running-ex}, the string $\mathbf{s}$, on the
+alphabet $\Sigma=\left\{ \textrm{A, \ensuremath{\ldots}\ , Z, a, \ensuremath{\ldots}\ , z, \_, "}\right\} $,
+is shown. The character \_ can be thought as a space. The table shown
+below $\mathbf{s}$ represents the span relation $P\left(\mathbf{s}\right)$,
+which is the result of applying the ternary spanner $P$ to $\mathbf{s}$.
+\end{example}
+
+\noindent I now present two special types of spanners that will be
+useful later on. The first one is that of \emph{Hierarchical Spanners}.
+In order to formulate their definition, we need to say when a $\mathbf{s}-\textrm{tuple}$
+is hierarchical.
+\begin{definition}
+\label{hierarchical-tuple}Given a string $\mathbf{s}$, a document
+spanner $P$ and an $\mathbf{s}-\textrm{tuple}$ $\mu\in P\left(\mathbf{s}\right)$,
+$\mu$ is hiearchical if, for every $x,y\in\textrm{SVars}\left(P\right)$,
+one of the following conditions holds:
+
 \begin{itemize}
@@
 \label{hierarchical-spanner}Given a document spanner $P$, $P$ is
-hierarchical if $\forall s\in\varSigma^{\ast}$, $\forall\mu\in P\left(s\right)$,
+hierarchical if $\forall\mathbf{s}\in\varSigma^{\ast}$, $\forall\mu\in P\left(\mathbf{s}\right)$,
 $\mu$ is hierarchical.
@@
 \noindent The class of hierarchical spanners is denoted by $\mathbf{HS}$.
-As it emerges from Definition \ref{hierarchical-tuple} and Definition
-\ref{hierarchical-spanner}, a hierarchical spanner outputs only $s-\textrm{tuples}$
-whose spans don't overlap. The second type of spanners I introduce
-is that of \emph{Universal Spanners}. Some accessory definitions are
-needed.
-\begin{definition}
-Given a string $s$ and a document spanner $P$, $P$ is total on
-$s$ if $P\left(s\right)$ consists of all the possible $s-\textrm{tuples}$
+The second type of spanners I introduce is that of \emph{Universal
+Spanners}. Some accessory definitions are needed.
+\begin{definition}
+Given a string $\mathbf{s}$ and a document spanner $P$, $P$ is
+total on $\mathbf{s}$ if $P\left(\mathbf{s}\right)$ consists of
+all the possible $\mathbf{s}-\textrm{tuples}$ on $\textrm{SVars}\left(P\right)$.
+\end{definition}
+
+\begin{definition}
+Given a string $\mathbf{s}$ and a document spanner $P$, $P$ is
+hierarchically total on $\mathbf{s}$ if $P\left(\mathbf{s}\right)$
+consists of all the possible hierarchical $\mathbf{s}-\textrm{tuples}$
 on $\textrm{SVars}\left(P\right)$.
-\end{definition}
-
-\begin{definition}
-Given a string $s$ and a document spanner $P$, $P$ is hierarchically
-total on $s$ if $P\left(s\right)$ consists of all the possible hierarchical
-$s-\textrm{tuples}$ on $\textrm{SVars}\left(P\right)$.
 \end{definition}
@@
 spanner $P$ such that $\textrm{SVars}\left(P\right)=Y$ and $P$
-is total on every string $s\in\varSigma^{\ast}$. 
+is total on every string $\mathbf{s}\in\varSigma^{\ast}$. 
 \end{definition}
@@
 unique document spanner $P$ such that $\textrm{SVars}\left(P\right)=Y$
-and $P$ is hierarchically total on every string $s\in\varSigma^{\ast}$. 
-\end{definition}
-
-\noindent I now present the formal spanner represantations described
-in REF.
+and $P$ is hierarchically total on every string $\mathbf{s}\in\varSigma^{\ast}$. 
+\end{definition}
+
+\noindent In the following Subsection, I present the formal spanner
+represantations described in \cite{Fagin:2015:DSF:2772377.2699442}.
 
@@
 
-We saw in Subsection \ref{sub:Algebra-of-Operators} that the operations
-used by SystemT and AQL to extract spans relations are regular expression
-matching and dictionary matching. Here I focus on regular expressions.
-AQL's regular expressions can be seen as usual regular expressions
-enriched with \emph{capture variables}, that are nothing but the span
-variables that constitute the span relations I defined. In what follows.
-In REF, three formalisms are described that are able to model this
-kind of modified regular expressions, namely:
+We saw, in Subsection \ref{subsec:Algebra-of-Operators}, that the
+operations used by SystemT and AQL to extract spans relations are
+regular expression matching and dictionary matching. In the remainder
+of this paper I focus on regular expressions. AQL's regular expressions
+can be seen as usual regular expressions enriched with \emph{capture
+variables}, that are precisely the span variables constituting the
+span relations of Definition \ref{span-relation}. In \cite{Fagin:2015:DSF:2772377.2699442},
+three formalisms are described, that are able to model this kind of
+modified regular expressions. They are:
 \begin{itemize}
@@
 
-To define regex formula, I first introduce the concept of \emph{variable
-regex.}
+In order to define a regex formula, let us introduce the concept of
+\emph{variable regex.}
 \begin{definition}
@@
  where:
+
 \begin{itemize}
@@
 A valid parse tree for a variable regex$\gamma$ is called a $\gamma-\textrm{parse}$.
-We accept only those variable regex expressions whose parses have
-only one occurence of each of their variables, otherwise the variable
-assignment would remain unclear. Such expressions are referred to
-as \emph{functional}.
-\begin{definition}
-A variable regex $\gamma$ is functional if $\forall s\in\varSigma^{\ast}$,
-$\forall\gamma-\textrm{parse }t$ for $s$, $\forall x\in\textrm{SVars}\left(\gamma\right)$,
+We accept only a variable regex expression if its parses have exactly
+one occurence of each of the variables appearing in the expression,
+otherwise the variable assignment would remain unclear. Such an expression
+is referred to as \emph{functional}.
+\begin{definition}
+A variable regex $\gamma$ is functional if $\forall\mathbf{s}\in\varSigma^{\ast}$,
+$\forall\gamma-\textrm{parse }t$ for $\mathbf{s}$, $\forall x\in\textrm{SVars}\left(\gamma\right)$,
 $x$ occurs exactly one time in $t$.
@@
 that $\textrm{SVars}\left(\left\llbracket \gamma\right\rrbracket \right)=\textrm{SVars}\left(\gamma\right)$,
-and the span relation $\left\llbracket \gamma\right\rrbracket \left(s\right)$
-is the set $\left\{ \mu^{p}\mid p\textrm{ is a }\gamma-\textrm{parse for \ensuremath{s}}\right\} $,
+and the span relation $\left\llbracket \gamma\right\rrbracket \left(\mathbf{s}\right)$
+is the set $\left\{ \mu^{p}\mid p\textrm{ is a }\gamma-\textrm{parse for \ensuremath{\mathbf{s}}}\right\} $,
 where $\mu^{p}$ is a tuple defined by a parse $p$. The class of
@@
 
+\begin{figure}
+\centering
+
+\noindent\resizebox{\textwidth}{!}{
+
+\tikz [level distance=1.2cm] 	\node {$\cdot$} 	[sibling distance=3cm] 		child[level distance=1cm] { node {$\ast$} 			[sibling distance=1.2cm] 			child { node {$\vee$} 				child {node[align=center] {\textbf{I} \\ 1} } 			} 			child { node {$\vee$} 				child {node[align=center] {\textbf{\_} \\ 2} } 			}			 		} 		child[level distance=2cm] { node {$z$} 			child { node {$\cdot$} 				[sibling distance=3.8cm] 				child{ node {$x$} 				       [level distance = 1cm] 					child { node {$\cdot$} 						[sibling distance=1.2cm, level distance = 2cm] 						child {node {$\vee$} 							[level distance = 1cm] 							child {node[align=center] {\textbf{w} \\ 3} } 						} 						child { node[align=center] {\textbf{a} \\ 4}} 						child { node[align=center] {\textbf{t} \\ 5}} 						child { node[align=center] {\textbf{c} \\ 6}} 						child { node[align=center] {\textbf{h} \\ 7}} 						child { node[align=center] {\textbf{e} \\ 8}} 						child { node[align=center] {\textbf{d} \\ 9}} 					} 				} 				child[level distance=1cm]{ node {$+$} 					[level distance=1cm] 					child { node {$\vee$}  						child { node[align=center] {\textbf{\_} \\ 10}} 						} 				} 				child{ node {$y$} 					child {node {$\cdot$} 						child[sibling distance =1.3cm, level distance=1cm] { node {$\vee$} 							child { node[align=center] {\textbf{"} \\ 11}} 						} 						child[level distance=1.2cm, sibling distance=1.2cm] {node {$\cdot$} 							child { node {$\vee$} 								[level distance=1cm] 								child { node[align=center] {\textbf{T} \\ 12}} 							} 							child { node {$\ast$} 								[sibling distance=1cm, level distance=1cm] 								child { node {$\vee$} 									child { node[align=center] {\textbf{h} \\ 13}} 								} 								child { node {$\vee$} 									child { node[align=center] {\textbf{e} \\ 14}} 								} 							} 						} 						child[sibling distance =3.8cm] { node {$\ast$} 							child {node {$\cdot$}  								[sibling distance=1.2cm, level distance=1.2cm] 								child { node[align=center] {\textbf{\_} \\ 15} } 								child { node {$\cdot$} 									[sibling distance=1.2cm] 									child { node {$\vee$} 										child { node[align=center] {\textbf{M} \\ 16}} 									} 									child[level distance=2.5cm, sibling distance=2.2cm] { node{$\ast$} 										[level distance=1cm, sibling distance=1cm] 										child { node {$\vee$} 											child { node[align=center] {\textbf{a} \\ 17}} 										} 										child { node {$\vee$} 											child { node[align=center] {\textbf{t} \\ 18}} 										} 										child { node {$\vee$} 											child { node[align=center] {\textbf{r} \\ 19}} 										} 										child { node {$\vee$} 											child { node[align=center] {\textbf{i} \\ 20}} 										} 										child { node {$\vee$} 											child { node[align=center] {\textbf{x} \\ 21}} 										} 									} 								}
+							} 						} 						child[sibling distance=2cm] { node {$\vee$} 							[level distance =1cm] 							child { node[align=center] {\textbf{"} \\ 22}} 						} 					} 				} 			} 		} 		child[level distance=1cm] { node {$\ast$} 			[sibling distance=1.2cm] 			child { node {$\vee$} 				child { node[align=center] {\textbf{\_} \\ 23}} 			} 			child { node {$\vee$} 				child { node[align=center] {.
+.
+.}} 			} 			child { node {$\vee$} 				child { node[align=center] {" \\ 47}} 			}		 		};
+
+}
+
+\caption{\label{fig:parse} A $\gamma$-parse $p$ for string $\mathbf{s}$
+(see Figure \ref{fig:running-ex}).}
+
+\end{figure}
+
+\begin{example}
+\label{ex-rgx}Consider the variable regex $\gamma$ defined by:
+
+\begin{equation}
+\Sigma^{\ast}\cdot z\left\{ x\left\{ \gamma_{\textrm{action}}\right\} \cdot\Sigma^{+}\cdot y\left\{ \gamma_{\textrm{title}}\right\} \right\} \cdot\Sigma^{\ast}\label{eq:rgx}
+\end{equation}
+
+where $\gamma_{\textrm{action}}=\left(\textrm{w}\cdot\textrm{a}\cdot\textrm{t}\cdot\textrm{c}\cdot\textrm{h}\cdot\textrm{e}\cdot\textrm{d}\right)\vee\left(\textrm{s}\cdot\textrm{a}\cdot\textrm{w}\right)$
+and $\gamma_{title}=\textrm{"}\cdot\left(\textrm{A}\vee\ldots\lor\textrm{Z}\right)\cdot\left(\textrm{a}\lor\ldots\lor\textrm{z}\right)^{\ast}\cdot\left(\_\cdot\left(\textrm{A}\vee\ldots\lor\textrm{Z}\right)\cdot\left(\textrm{a}\lor\ldots\lor\textrm{z}\right)^{\ast}\right)^{\ast}\cdot\textrm{"}$.
+Notice that $\textrm{SVars}\left(\gamma\right)=\left\{ x,y,z\right\} $.
+Figure \ref{fig:parse} shows a $\gamma$-parse $p$ for string $\mathbf{s}$
+of Figure \ref{fig:running-ex}. In this parse, each variable contained
+in $\textrm{SVars}\left(\gamma\right)$ occurs exactly once. It is
+easy to verify that this holds for every $\gamma$-parse. Hence, $\gamma$
+is functional. We have that $\mu^{p}\left(x\right)=\left[3,10\right\rangle $,
+$\mu^{p}\left(y\right)=\left[11,23\right\rangle $ and $\mu^{p}\left(z\right)=\left[3,23\right\rangle $,
+thus $\mu^{p}$ is the $\mathbf{s}$-tuple $\mu_{1}$ from Figure
+\ref{fig:running-ex}. Finally, considering the span relation $P\left(\mathbf{s}\right)$
+(Figure \ref{fig:running-ex}), we have that $\left\llbracket \gamma\right\rrbracket \left(\mathbf{s}\right)=P\left(\mathbf{s}\right)$.
+\end{example}
+
 \subsubsection{Variable Stack Automata}
 
-Variable stack automata (VStack Automata or vstk-automata for short)
-are representations of document spanners by means of modified finite
-state automata (FAs). Basically, a FA is augmented with a \emph{stack
-of span variables}. A variable is pushed on the stack when its corresponding
-span is opened, and popped when it is closed. The formal definition
-of a vstk-automaton follows.
+Variable stack automata (vstk-automata for short) are representations
+of document spanners by means of modified finite state automata (FSAs).
+Basically, a FSA is augmented with a \emph{stack of span variables}.
+A variable is pushed on the stack when its corresponding span is opened,
+and popped when it is closed. The formal definition of a vstk-automaton
+follows.
 \begin{definition}
@@
 where:
+
 \begin{itemize}
@@
 \noindent Notice that we don't need to specify which variable we want
-to pop, as it is naturally the last that was pushed on the stack.
+to pop, as it is naturally the last that was pushed onto the stack.
 Given a vstk-automaton $A$, the set of variables that appear in its
-transitions is denoted as $\textrm{SVars}\left(A\right)$. Next I
-give the definitions of a \emph{configuration }and of a \emph{run}
-of a vstk-automaton.
-\begin{definition}
-\label{vstk-conf}Given a string $s$ with length $\left|s\right|=n$
+transitions is denoted as $\textrm{SVars}\left(A\right)$.
+
+\begin{figure}
+\centering
+
+\noindent\resizebox{\textwidth}{!}{
+
+\centering\begin{tikzpicture}[shorten >=1pt,node distance=3cm,on grid,auto]  \node[state,initial]  (q_0)                {};  \node[state]          (q_1) [right=of q_0] {};  \node[state]          (q_2) [right=of q_1] {}; \node[state]          (q_3) [above right=of q_2] {}; \node[state]          (q_4) [right=of q_3] {}; \node[state]          (q_5) [right=of q_4] {}; \node[state]          (q_6) [right=of q_5] {}; \node[state]          (q_7) [right=of q_6] {}; \node[state]          (q_8) [right=of q_7] {}; \node[state]          (q_9) [right=of q_2,xshift=3.5cm] {}; \node[state]          (q_10) [right=of q_9,xshift=3.5cm] {}; \node[state]          (q_11) [below right=of q_8] {}; \node[state]          (q_12) [below=of q_11] {}; \node[state]          (q_13) [left=of q_12] {}; \node[state]          (q_14) [left=of q_13] {}; \node[state]          (q_15) [left=of q_14] {}; \node[state]          (q_16) [left=of q_15] {}; \node[state]          (q_17) [above left=of q_16,xshift=4cm] {}; \node[state]          (q_18) [left=of q_17,xshift=-0.8cm] {}; \node[state]          (q_19) [left=of q_16] {}; \node[state,accepting](q_20) [left=of q_19] {};  \path[->] (q_0) edge              node        {$z\vdash$} (q_1)  		edge [loop above] node        {$\Sigma$} ()  	  (q_1) edge 	          node        {$x\vdash$} (q_2)  	  (q_2) edge              node        {w} (q_3)  	  (q_2) edge              node        {s} (q_9)  	  (q_3) edge              node        {a} (q_4)  	  (q_4) edge              node        {t} (q_5)  	  (q_5) edge              node        {c} (q_6)  	  (q_6) edge              node        {h} (q_7)  	  (q_7) edge              node        {e} (q_8)  	  (q_8) edge              node        {d} (q_11)  	  (q_9) edge              node        {a} (q_10)  	  (q_10) edge              node        {w} (q_11)  	  (q_11) edge              node        {$\dashv$} (q_12)  	  (q_12) edge              node        {$\Sigma$} (q_13)  	  (q_13) edge              node        {$y\vdash$} (q_14)  	  	 edge [loop above] node        {$\Sigma$} ()  	  (q_14) edge              node        {"} (q_15)  	  (q_15) edge              node        {A,...,Z} (q_16)  	  (q_16) edge              node        {\_} (q_17)  	  (q_17) edge 		   node        {A,...,Z} (q_18)  	  (q_16) edge              node        {"} (q_19)  	  (q_18) edge [loop below] node        {a,..,z} () 		 edge 		   node        {$\epsilon$} (q_16)  	  (q_19) edge              node        {$\epsilon$} (q_20)  	  	 edge [loop above] node        {$\dashv$} ()  	  (q_20) edge [loop above] node        {$\Sigma$} (); \end{tikzpicture}
+
+}
+
+\caption{\label{fig:vstk}A vstk-automaton $A$ with $\left\llbracket A\right\rrbracket =\left\llbracket \gamma\right\rrbracket $,
+where $\gamma$ is the regex formula of \ref{eq:rgx} (see Example
+\ref{ex-rgx}).}
+\end{figure}
+
+\begin{example}
+Figure \ref{fig:vstk} shows a vstk-Automaton $A$. Each circle represents
+a state. The double circle is the accepting state. A label $a$ on
+an edge from state $q$ to state $q^{'}$ represents the transition
+$\left(q,a,q^{'}\right)$. A sequence $\sigma_{1},\ldots,\sigma_{k}$
+on an edge from $q$ to $q^{'}$ is a shorthand for the $k$ transitions
+$\left(q,a_{1},q^{'}\right),\ldots,\left(q,a_{k},q^{'}\right)$. Assuming
+$\Sigma=\left\{ \sigma_{1},\ldots,\sigma_{l}\right\} $, then the
+label $\Sigma$ is used in place of $\sigma_{1},\ldots,\sigma_{l}$.
+\end{example}
+
+\noindent Next, I report the definitions of a \emph{configuration
+}and of a \emph{run} of a vstk-automaton.
+\begin{definition}
+\label{vstk-conf}Given a string $\mathbf{s}$ with length $\left|\mathbf{s}\right|=n$
 and a vstk-automaton $A$ , a configuration of $A$ is a tuple $c=\left(q,\overrightarrow{v},Y,i\right)$,
 where:
+
 \begin{itemize}
@@
 \item $i\in\left\{ 1,...,n+1\right\} $ is the position of the next character
-to be read in $s$.
+to be read in $\mathbf{s}$.
 \end{itemize}
@@
 For the rest, a run of a vstk-automaton is similar to those of ordinary
-FAs.
+FSAs.
 \begin{definition}
@@
 on $s$ is a sequence of configurations $c_{0},...,c_{m}$ such that:
+
 \begin{itemize}
@@
 \item $\overrightarrow{v_{j}}=\overrightarrow{v_{j+1}}\cdot x$, $x\in Y_{j}$,
-$Y_{j+1}=Y_{j}$, $\left(q_{j},\dashv,q_{j+1}\right)\in\delta$ (a
-variable is popped from the stack).
-\end{itemize}
-\end{itemize}
-\end{itemize}
-\end{definition}
-
-\begin{definition}
-Given a string with length $\left|s\right|=n$ and a vstk-automaton
-$A$, a run $\rho=c_{0},...,c_{m}$ of $A$ on $s$ is accepting if
-$c_{m}=\left(q_{f},\epsilon,\textrm{Ø},n+1\right)$.
+$Y_{j+1}=Y_{j}$, $\left(q_{j},\dashv,q_{j+1}\right)\in\delta$ (the
+variable on top of the stack is popped).
+\end{itemize}
+\end{itemize}
+\end{itemize}
+\end{definition}
+
+\begin{definition}
+Given a string $\mathbf{s}$ with length $\left|\mathbf{s}\right|=n$
+and a vstk-automaton $A$, a run $\rho=c_{0},...,c_{m}$ of $A$ on
+$\mathbf{s}$ is accepting if $c_{m}=\left(q_{f},\epsilon,\textrm{Ø},n+1\right)$.
 \end{definition}
@@
 \noindent The set of all possible accepting runs of a vstk-automaton
-$A$ on a string $s$ is denoted as $\textrm{ARuns}\left(A,s\right)$.
+$A$ on a string $\mathbf{s}$ is denoted as $\textrm{ARuns}\left(A,\mathbf{s}\right)$.
 The spanner represented by $A$ may be referred to as $\left\llbracket A\right\rrbracket $.
 We have that $\textrm{SVars}\left(\left\llbracket A\right\rrbracket \right)=\textrm{SVars}\left(A\right)$,
-and the span relation $\left\llbracket A\right\rrbracket \left(s\right)$
-is the set $\left\{ \mu^{\rho}\mid\rho\in\textrm{ARuns}\left(A,s\right)\right\} $,
+and the span relation $\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)$
+is the set $\left\{ \mu^{\rho}\mid\rho\in\textrm{ARuns}\left(A,\mathbf{s}\right)\right\} $,
 where $\mu^{\rho}$ is a tuple defined by a run $\rho$. In particular,
@@
 \noindent The class of variable stack automata is called $\textrm{VA}_{\textrm{stk}}$.
+\begin{example}
+Consider the regex formula $\gamma$ from Example \ref{ex-rgx} and
+the vstk-automaton $A$ from Figure \ref{fig:vstk}. We have that
+$\textrm{SVars}\left(A\right)=\left\{ x,y,z\right\} $. The reader
+can verify that $\left\llbracket \gamma\right\rrbracket =\left\llbracket A\right\rrbracket $.
+\end{example}
 
@@
 
-Variable Set Automata (VSet Automata or vset-automata for short) are
-another formalism for representing document spanners that is based
-on FAs. Vset-automata are defined in a very similar way to that of
-vstk-automata. The main differences are:
+Variable Set Automata (vset-automata for short) are another formalism
+for representing document spanners that is based on FSAs. Vset-automata
+are defined in a very similar way to that of vstk-automata. The main
+differences are:
 \begin{itemize}
 \item the stack of span variables is replaced by a \emph{set;}
-\item since no order is defined on the variables in the set, when we close
+\item no order is defined on the variables in the set, so when we close
 a variable, we need to specify which one.
 \end{itemize}
-I now give a formal definition of a vset-automaton.
+The following is the formal definition of a vset-automaton.
 \begin{definition}
@@
 \begin{definition}
-Given a string $s$ with length $\left|s\right|=n$ and a vset-automaton
-$A$ , a configuration of $A$ is a tuple $c=\left(q,V,Y,i\right)$,
+Given a string $\mathbf{s}$ with length $\left|\mathbf{s}\right|=n$
+and a vset-automaton $A$ , a configuration of $A$ is a tuple $c=\left(q,V,Y,i\right)$,
 where:
+
 \begin{itemize}
@@
 \item $Y\in Vars$$\left(A\right)$ is the set of available variables (those
-not already placed in the set).
-\end{itemize}
-\end{definition}
-
-\begin{definition}
-Given a string $s$ and a vset-automaton $A$, a run $\rho$ of $A$
-on $s$ is a sequence of configurations $c_{0},...,c_{m}$ such that:
+not already inserted in the set).
+\end{itemize}
+\end{definition}
+
+\begin{definition}
+Given a string $\mathbf{s}$ and a vset-automaton $A$, a run $\rho$
+of $A$ on $\mathbf{s}$ is a sequence of configurations $c_{0},...,c_{m}$
+such that:
+
 \begin{itemize}
@@
 \begin{definition}
-Given a string with length $\left|s\right|=n$ and a vset-automaton
-$A$, a run $\rho=c_{0},...,c_{m}$ of $A$ on $s$ is accepting if
-$c_{m}=\left(q_{f},\textrm{Ø},\textrm{Ø},n+1\right)$.
-\end{definition}
-
-\noindent Given a string $s$ and a vset-automaton $A$, $\textrm{ARuns}\left(A,s\right)$and
-$\left\llbracket A\right\rrbracket $ are defined as for vstk-automata.
-The class of variable set automata is called $\textrm{VA}_{\textrm{stk}}$.
-
-\subsection{Algebras of Spanners\label{sub:Algebras-of-Spanners}}
+Given a string with length $\left|\mathbf{s}\right|=n$ and a vset-automaton
+$A$, a run $\rho=c_{0},...,c_{m}$ of $A$ on $\mathbf{s}$ is accepting
+if $c_{m}=\left(q_{f},\textrm{Ø},\textrm{Ø},n+1\right)$.
+\end{definition}
+
+\noindent Given a string $\mathbf{s}$ and a vset-automaton $A$,
+$\textrm{ARuns}\left(A,\mathbf{s}\right)$and $\left\llbracket A\right\rrbracket $
+are defined as for vstk-automata. The class of variable set automata
+is called $\textrm{VA}_{\textrm{set}}$.$y_{1}\vdash,\ldots y_{m}\vdash,$
+
+\begin{figure}
+\centering
+
+\begin{tikzpicture}[shorten >=1pt,node distance=3cm,on grid,auto]  \node[state,initial,accepting]  (q_0)                      {};  \path[->] (q_0) edge [loop above] node        {$\Sigma$} () 	        edge [loop below] node[align=center]        {$y_{1}\vdash,\ldots y_{m}\vdash,$ \\ $\dashv y_{1},\ldots \dashv y_{m}$} (); \end{tikzpicture}\caption{\label{fig:vset}A vset-automaton $B$ with $\left\llbracket B\right\rrbracket =\Upsilon_{Y}$,
+for $Y=\left\{ y_{1},\ldots,y_{m}\right\} $ (this image is taken
+from \cite{Fagin:2015:DSF:2772377.2699442}).}
+\end{figure}
+
+\begin{example}
+Figure \ref{fig:vset} shows a vset-automaton $B$, with $\textrm{SVars}\left(B\right)=Y$,
+where $Y=\left\{ y_{1},\ldots,y_{m}\right\} $. We have that $\left\llbracket B\right\rrbracket =\Upsilon_{Y}$.
+This shows that vset-automata can express spanners that regex formulas
+and vstk-automata cannot. (This example was adapted from \cite{Fagin:2015:DSF:2772377.2699442}).
+\end{example}
+
+\subsection{Algebras of Spanners\label{subsec:Algebras-of-Spanners}}
 
 Besides mere span extraction, AQL offers the capability to combine,
-transform and filter extracted tuples by using a series of relational
-operators. Here, I introduce the relational operators for spanners,
-described in REF, that are considered to capture the core relational
-operations of AQL. They are:
-\begin{itemize}
-\item \emph{Union }( $\cup$);
+transform and filter extracted tuples by using a series of operators.
+Here, I introduce the relational operators for spanners, described
+in \cite{Fagin:2015:DSF:2772377.2699442}, that are considered to
+capture the core relational operations of AQL. They are:
+\begin{itemize}
+\item \emph{Union }($\cup$);
 \item \emph{Projection }($\pi$);
@@
 A finite set of spanner operators forms a \emph{spanner algebra}.
-In the following, give definitions for each of the listed spanner
+In the following, I look at the definitions of the listed spanner
 operators.
@@
 \begin{definition}
-Given two document spanner $P_{1}$ and $P_{2}$, they are union compatible
-if and only if $\textrm{SVars}\left(P_{1}\right)=\textrm{SVars}\left(P_{2}\right)$.
+Given two document spanners $P_{1}$ and $P_{2}$, they are union
+compatible if and only if $\textrm{SVars}\left(P_{1}\right)=\textrm{SVars}\left(P_{2}\right)$.
 \end{definition}
@@
 we have that:
+
 \begin{itemize}
@@
 a document spanner $P$ over a subset of its span variables $Y$ is
-obtained by reducing the domain of each of the spanner's output $s-\textrm{tuples}$
+obtained by reducing the domain of each of the spanner's output $\mathbf{s}-\textrm{tuples}$
 to $Y$.
@@
 that:
+
 \begin{itemize}
@@
 
+\begin{example}
+Consider again the regex formula $\gamma$ from Example \ref{ex-rgx}.
+The reader can verify that $\gamma$ can be expressed as:
+
+\ldots{}
+\begin{equation}
+\left(\Sigma^{\ast}\cdot x\left\{ \gamma_{action}\right\} \cdot\Sigma^{\ast}\right)\bowtie\left(\Sigma^{\ast}\cdot y\left\{ \gamma_{title}\right\} \cdot\Sigma^{\ast}\right)\bowtie\left(\Sigma^{\ast}\cdot z\left\{ x\left\{ \Sigma^{+}\right\} \cdot\Sigma^{+}\cdot y\left\{ \Sigma^{+}\right\} \right\} \cdot\Sigma^{\ast}\right)
+\end{equation}
+.
+\end{example}
+
 \subsubsection{String Selection}
@@
 $\left\llbracket \textrm{SR}\right\rrbracket $. Formally, we have
-that $\left\llbracket \textrm{SR}\right\rrbracket =\left\{ \left\llbracket r\right\rrbracket \mid\in\textrm{SR}\right\} $.
+that $\left\llbracket \textrm{SR}\right\rrbracket =\left\{ \left\llbracket r\right\rrbracket \mid r\in\textrm{SR}\right\} $.
 Let $O$ be a spanner algebra. I denote by $\textrm{SR}^{O}$ the
@@
 We are ready to start reasoning on the expressive power of the spanner
-representations that we defined.
+representations that we saw. Please refer to \cite{Fagin:2015:DSF:2772377.2699442}
+for the proofs of the mathematical statements in the remainder of
+this Chapter.
 \begin{proposition}
 The following hold:
+
 \begin{enumerate}
@@
 \noindent In the next Subsection, I discuss the class of \emph{regular
-spanners}, which plays a central role in the discussion of the class
-of spanners that constitute the core of AQL.
+spanners}, which plays a central role in the construction of the class
+of spanners constituting the core of AQL.
 
@@
 
-\noindent Let us now see how regular spanners are related to the other
+\noindent Let us see how regular spanners are related to the other
 basic spanner classes. A preliminary result is that regex formulas
@@
 \noindent For what concerns the algebraic operators I presented in
-Subsection \ref{sub:Algebras-of-Spanners}, it can be shown that union,
-projection and join don't increase the expressive power of regular
-spanners.
+Subsection \ref{subsec:Algebras-of-Spanners}, it can be shown that
+union, projection and join don't increase the expressive power of
+regular spanners.
 \begin{theorem}
@@
 
-\noindent On the other hand, the same operators make the spanners
-represented in $\textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}$
-equivalent to regular spanners.
+\noindent On the other hand, applying the same operators to the spanner
+representations in $\textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}$
+results in a class equivalent to regular spanners.
 \begin{theorem}
 \label{.stkext=00003Dset}$\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}^{\left\{ \cup,\pi,\bowtie\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}\right\rrbracket $.
+
+\medskip{}
 \end{theorem}
 
-\noindent Finally, I look at which string relations can be simulated
-by regular spanners. Let us start by defining the concept of \emph{selectable
+\noindent Let us now look at which string relations can be simulated
+by regular spanners, starting with the concept of \emph{selectable
 string relation}.
@@
 it is not selectable by regular spanners, nonetheless it is important
-for selection predicates in AQL, which is why regular spanners are
-not enough to model its core. In the following Subsection, by using
-the notions and results presented so far, I describe the spanners
-that are identified in REF as the core of AQL: \emph{core spanners}.
+for selection predicates in AQL, which is why regular spanners don't
+have enough expressive power to model its core. In the following Subsection,
+I describe the spanners that are identified in \cite{Fagin:2015:DSF:2772377.2699442}
+as the core of AQL: \emph{core spanners}.
 
@@
 \begin{definition}
-A core spanner is a document spanner belonging to $\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket $.
+\label{core-def}A core spanner is a document spanner belonging to
+$\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket $.
 \end{definition}
@@
 \noindent Thanks to Theorems \ref{.rgx=00003Dvstk}, \ref{.setext=00003Dset}
-and \ref{.stkext=00003Dset} we can easily see that $\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket $.
-This shows that core spanners can be reduced to regular spanners extended
-with the algebra $\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} $.
+and \ref{.stkext=00003Dset} we can easily state the next theorem. 
+\begin{theorem}
+\noindent \label{RGX=00003DSET}$\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{stk}}}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{\ensuremath{\textrm{VA}_{\textrm{set}}}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket $.
+\end{theorem}
+
+\noindent This shows that core spanners can be reduced to regular
+spanners extended with the algebra $\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} $.
 But the following lemma tells us an algebra with fewer operators is
@@
 \noindent Another lemma, known as the \emph{core simplification lemma},
-gives a even more simple way of representing core spanners.
+gives an even more simple way of representing core spanners.
 \begin{lemma}
-$\mathbf{(Core\,Simplification\,Lemma)}$ Every core spanner can be
-defined by an expression of the form
+\label{Core-simplif}$\mathbf{(Core\,Simplification\,Lemma)}$ Every
+core spanner can be defined by an expression of the form
 
@@
 
-\noindent Now I look at which string relations can be simulated by
-core spanners. In the next definition, I define three string relations
+\noindent For what concerns which string relations can be simulated
+by core spanners, the next definition presents three string relations
 of relevance.
@@
 Given two strings $\mathbf{s},\mathbf{t}\in\varSigma^{\ast}$:
+
 \begin{itemize}
@@
 we have:
+
 \begin{itemize}
@@
 \begin{theorem}
-$\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket \subsetneq\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=},\setminus\right\} }\right\rrbracket $.
+\label{.last-stat}$\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket \subsetneq\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=},\setminus\right\} }\right\rrbracket $.
 \end{theorem}
+
+\chapter{A Runtime System for AQL's Core}
+
+In this chapter I describe a new runtime system for AQL's core fragments.
+It is based on a modified version of vset-automata, the \emph{extended
+vset-automata (}or \emph{eVset-automata).} In particular, the system
+works with a particular kind of eVset-automata: \emph{well-behaved
+eVset-automata. }I start by defining eVset-automata, then I describe
+well-behaved eVset-automata and their properties, with particular
+attention to their ability to support the operations from the algebra
+presented in Subsection \ref{subsec:Algebras-of-Spanners}. Subsequently,
+I review some interesting ways of running usual nondeterministic automata
+on a string, from which I took inspiration for my implementation.
+Finally I explain how the system works.
+
+\section{Extended Vset-automata}
+
+Before I define extended vset-automata formally, I introduce some
+basic concepts.
+\begin{definition}
+Given a set $X\subseteq\textrm{SVars}$, we define:
+
+\begin{itemize}
+\item $\textrm{SVOps}^{\vdash}\left(X\right)=\left\{ x\vdash\mid x\in X\right\} $;
+\item $\textrm{SVOps}^{\dashv}\left(X\right)=\left\{ \dashv x\mid x\in X\right\} $;
+\item $\textrm{SVOps}\left(X\right)=\textrm{SVOps}^{\vdash}\left(X\right)\cup\textrm{SVOps}^{\dashv}\left(X\right)$.
+\end{itemize}
+\end{definition}
+
+\noindent I now define an extended vset-automaton.
+\begin{definition}
+An extended \emph{variable-set automaton} (or \emph{eVset-automaton})
+is a tuple $\left(Q,q_{0},q_{f},\delta\right)$, where:
+
+\begin{itemize}
+\item $Q$ is a finite set of states;
+\item $q_{0}\in Q$ is the initial state;
+\item $q_{f}\in Q$ is the accepting state;
+\item $\delta=\delta^{\textrm{char}}\cup\delta^{\textrm{op}}$ is a finite
+transition relation consisting of triples, where:
+
+\begin{itemize}
+\item $\delta^{\textrm{char}}=\left\{ \left(q,\sigma,q'\right)\mid q,q'\in Q,\sigma\in\varSigma\right\} $,
+whose elments are called character transitions;
+\item $\delta^{\textrm{op}}=\left\{ \left(q,S,q'\right)\mid q,q'\in Q,S\subseteq\textrm{SVOps\ensuremath{\left(SVars\right)}}\right\} $,
+whose elements are called operation transitions .
+\end{itemize}
+\end{itemize}
+\end{definition}
+
+\noindent In a transition$\left(q,a,q'\right)$ with label $a$ from
+state $q$ to state $q'$, $q$ is called the \emph{source state},
+while $q'$ is called the \emph{destination state}. By a slight abuse
+of notation, I denote by $\textrm{SVOps}\left(A\right)$ the set of
+variable operations appearing in the transitions of an eVset-automaton
+$A$. $\textrm{SVars}\left(A\right)$ is defined as for a usual vset-automaton.
+I now define a configuration and, subsequently, a run of an eVset-automaton.
+\begin{definition}
+Given a string $\mathbf{s}$, $\left|\mathbf{s}\right|=n$, and an
+eVset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$, a configuration
+of A is a tuple $c=\left(q,V,Y,i\right)$, where:
+
+\begin{itemize}
+\item $q\in Q$ is the current state;
+\item $V\subseteq\textrm{SVars}\left(A\right)$ is the active variable set;
+\item $Y\subseteq\textrm{SVars}\left(A\right)$ is the set of available
+variables;
+\item $i$ is an index belonging to $\left\{ 1,\ldots,n+1\right\} $.
+\end{itemize}
+\end{definition}
+
+\begin{definition}
+\label{e-run}Given a string $\mathbf{s}=s_{1},\ldots,s_{n}$ and
+an eVset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$, a run $\rho$
+of $A$ on $\mathbf{s}$ is a sequence $c_{0},\ldots,c_{m}$ of configurations,
+where:
+
+\begin{itemize}
+\item $c_{0}=\left(q_{0},\textrm{Ø},\textrm{SVars}\left(A\right),1\right)$;
+\item for $j=0,\ldots,m-1$ one of the following holds for $c_{j}=\left(q_{j},V_{j},Y_{j},i_{j}\right)$
+and $c_{j+1}=\left(q_{j+1},V_{j+1},Y_{j+1},i_{j+1}\right)$:
+
+\begin{itemize}
+\item $V_{j+1}=V_{j}$, $Y_{j+1}=Y_{j}$, $i_{j+1}=i_{j}+1$ and $\left(q_{j},s_{j}q_{j+1}\right)\in\delta^{\textrm{char}}$;
+\item $i_{j+1}=i_{j}$, and for some $S\subseteq\textrm{SVOps}\left(A\right)$
+we have :
+
+\begin{itemize}
+\item for each $x\in\textrm{SVars}\left(A\right)$:
+
+\begin{itemize}
+\item $\left\{ x\vdash,\dashv x\right\} \cap S=\left\{ x\vdash\right\} $
+or $\left\{ x\vdash,\dashv x\right\} \cap S=\left\{ x\vdash,\dashv x\right\} $
+if and only if $x\in Y_{j}$;
+\item $\left\{ x\vdash,\dashv x\right\} \cap S=\left\{ \dashv x\right\} $
+if and only if $x\in V_{j}$;
+\end{itemize}
+\item $V_{j+1}=V_{j}\cup\left\{ x\mid x\vdash\in S\right\} \setminus\left\{ x\mid\dashv x\in S\right\} $;
+\item $Y_{j+1}=Y_{j}\setminus\left\{ x\mid x\vdash\in S\right\} $;
+\item $\left(q_{j},S,q_{j+1}\right)\in\delta^{\textrm{op}}$.
+\end{itemize}
+\end{itemize}
+\end{itemize}
+$\rho$ is accepting if $c_{m}=\left(q_{f},\textrm{Ø,}\textrm{Ø},n+1\right)$.
+\end{definition}
+
+$\textrm{ARuns}\left(A,\mathbf{s}\right)$ and $\left\llbracket A\right\rrbracket $,
+for an eVset-automaton $A$ and a string $\mathbf{s}$, are defined
+in similar ways to those of usual vset-automata. This new kind vset-automata
+allows to perform an arbitrary number of variable operations in one
+transition. The operations in a transition are to be performed in
+a given order. The exact order isn't very important, except for the
+fact that, in a valid run, the insertion and removal of a variable
+from the active variable set need to happen in the correct order (i.e.
+insertion first).
+
+\begin{figure}
+\centering\begin{tikzpicture}[shorten >=1pt,node distance=3cm,on grid,auto] \node[state,initial]  (q_0)                      {$q_0$}; \node[state]          (q_1) [right=of q_0] {$q_1$}; \node[state]          (q_2) [right=of q_1] {$q_2$}; \node[state,accepting](q_3) [right=of q_2] {$q_3$}; \path[->] (q_0) edge              node        {$\left\{ x\vdash\right\} $} (q_1) 		edge [loop above] node        {$\varSigma$} () 	  (q_1) edge 	          node        {$\left\{ y\vdash\right\} $} (q_2) 		edge [loop above] node        {a} (q_2) 	  (q_2) edge              node        {$\left\{ \dashv x,\dashv y\right\} $} (q_3) 		edge [loop above] node        {b} () 	  (q_3) edge [loop above] node        {$\varSigma$} (); \end{tikzpicture}
+
+\caption{\label{fig:evset}An eVset-automaton $A$.}
+\end{figure}
+
+The reason for using eVset-automata instead of plain vset-automata
+will be clear in the next section. I now introduce some additional
+useful concepts, the most important of which being that of a \emph{path}
+in an eVset-automaton.
+\begin{definition}
+Given a transition $t\in\delta$ in an eVset-automaton $A$, and an
+ordering $\varphi$ on the elements of $\textrm{SVOps}\left(A\right)$,
+we define
+
+\begin{itemize}
+\item $\textrm{Ops}\left(t\right)$as either:
+
+\begin{itemize}
+\item if $t=\left(q,S,q'\right)\in\delta^{\textrm{op}}$, the set $S$;
+\item if $t\in\delta^{\textrm{char}}$, the empty set.
+\end{itemize}
+\item $\textrm{LOps}_{\varphi}\left(t\right)$ as either:
+
+\begin{itemize}
+\item if $t=\left(q,S,q'\right)\in\delta^{\textrm{op}}$, the list $o_{1},\ldots,o_{\left|S\right|}$
+of operations belonging to $S$, ordered according to $\varphi$;
+\item if $t\in\delta^{\textrm{char}}$, the empty list.
+\end{itemize}
+\end{itemize}
+\end{definition}
+
+\begin{definition}
+\noindent Given an extended vset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$
+and a pair of states $q,q'\in Q$, a path $p$ between $q$ and $q'$
+in $A$ is a sequence of transitions $t_{1},$$\ldots,t_{n}\in\delta$,
+such that:
+
+\begin{itemize}
+\item the source state of $t_{1}$ is $q$;
+\item the destination state of $t_{n}$ is $q'$;
+\item for every pair $t_{i},t_{i+1}$, the destination state of $t_{i}$
+equals the source state of $t_{i+1}$.
+\end{itemize}
+We also write $p_{q}^{q'}$. We refer to the set of paths in $A$
+as $\textrm{Paths}\left(A\right)$.
+\end{definition}
+
+A path in an eVset-automaton is closely related to a run, as formalized
+by the next definition.
+\begin{definition}
+Given a string $\mathbf{s}$, an eVset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$,
+a run $\rho=c_{0},$$\ldots,c_{m}$ of $A$ on $\mathbf{s}$ and a
+path $p=t_{0},\ldots,t_{m-1}$ in $A$, we say that $p$ supports
+$\rho$ if, for every pair $c_{i}$, $c_{i+1}$ of configurations,
+$c_{i+1}$is obtained from $c_{i}$ by applying $t_{i}$, using the
+rules given in Definition \ref{e-run}.
+\end{definition}
+
+\begin{definition}
+Given a path $p=t_{1},\ldots,t_{n}$ in an eVset-automaton $A$, and
+considering an ordering $\varphi$ on the elements of $\textrm{SVOps}\left(A\right)$,
+we define
+
+\begin{itemize}
+\item the set $\textrm{Ops}\left(p\right)$ as
+\[
+\bigcup_{i=1}^{n}\textrm{Ops}\left(t_{i}\right)
+\]
+\item the list $\textrm{LOps}_{\varphi}\left(p\right)$ as
+\[
+\bigoplus_{i=1}^{n}\textrm{LOps}_{\varphi}\left(t_{i}\right)
+\]
+where $\bigoplus$ is the usual list concatenation operator.
+\end{itemize}
+If every transition in $p$ belongs to $\delta^{\textrm{op}}$, we
+say that $p$ is an operation-only path.
+\end{definition}
+
+From now on, without loss of generality, I consider a fixed order
+$\varphi$, in which insertions come before deletions.
+
+\medskip{}
+
+As shown in the previous chapter, the class of vset-automata doesn't
+get more expressive power when extended with the algebra$\left\{ \cup,\pi,\bowtie\right\} $.
+Hence, this class is a good candidate to be the base for a class of
+representations for the core spanners, as the Core Simplification
+Lemma suggests. In the following, I show that the same result holds
+for a subclass of extended vset-automata: \emph{well-behaved eVset-automata}.
+Moreover, we will see that an AQL core fragment, as defined in \cite{Fagin:2015:DSF:2772377.2699442},
+can be converted into a well-behaved eVset-automaton. The runtime
+system implicitly assumes that any eVset-automaton it receives as
+input is well-behaved.
+
+\section{Well-Behaved Extended Vset-Automata}
+
+In order to be well-behaved, an eVset-automaton has to respect some
+constraints on its paths that are between its initial and accepting
+state. I call this kind of paths \emph{complete paths.}
+\begin{definition}
+\noindent Given an eVset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$,
+a path $p=t_{1},\ldots,t_{n}$ in $A$ is complete if it is between
+$q_{0}$ and $q_{f}$, that is $p=p_{q_{0}}^{q_{f}}$.
+\end{definition}
+
+I now formally define a well-behaved eVset-automaton.
+\begin{definition}
+\noindent \label{well-behaved}An eVset-automaton $A$ is well-behaved
+if, for every complete path $p$ in $A$ we have:
+
+\begin{enumerate}
+\item for every $o\in\textrm{SVOps}\left(A\right)$, $o$ appears exactly
+once in $\textrm{LOps}\left(p\right)$;\label{enu:well-behaved-1,}
+\item for every pair of operations $x\vdash$,$\dashv x$$\in\textrm{SVOps}\left(A\right)$,
+$x\vdash$ appears before $\dashv x$ in $\textrm{LOps}\left(p\right)$.
+\end{enumerate}
+\end{definition}
+
+According to the definition, well-behaved eVset-automata guarantee
+that any of their complete paths will support a valid accepting run.
+Since states that cannot reach the final state, or that cannot be
+reached from the initial state, are not used in complete paths, we
+might want to consider only well-behaved eVset-automata where these
+states don't exist.
+\begin{definition}
+Given an eVset-automaton $A$, $A$ is pruned if for every state $q$
+in $A$, there is a path from $q_{0}$ to $q$, and a path from $q$
+to $q_{f}$ in $A$.
+\end{definition}
+
+\begin{example}
+The eVset automaton $A$ from Figure \ref{fig:evset} and the eVset-automaton
+$B$ from Figure \ref{evset-2} are well-behaved and pruned.
+\end{example}
+
+The next proposition tells us that the we can always prune a well-behaved
+eVset-automaton, obtaining a new eVset-automaton equivalent to the
+original.
+\begin{proposition}
+Given a well-behaved eVset-automaton $A$, there exists a pruned well-behaved
+eVset automaton $A'$ such that $\left\llbracket A'\right\rrbracket =\left\llbracket A\right\rrbracket $.
+Moreover, $A'$ can be produced in polynomial time.
+\end{proposition}
+
+\begin{proof}
+To obtain $A'$, it is sufficient to remove from $A$ those states
+from which we can't reach the final state, those states that are unreachable
+from the initial state, and all the transitions that have them as
+source or destination states. To test each state for removal we can
+use, e. g., Dijkstra's shortest path algorithm, which ensures $A'$
+can be found in polynomial time. Because of our construction, $A'$
+is well-behaved, as all its complete paths are also in $A$. We now
+show that, for every string $\mathbf{s},$ we have $\left\llbracket A'\right\rrbracket \left(\mathbf{s}\right)=\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)$.
+To see that $\left\llbracket A'\right\rrbracket \left(\mathbf{s}\right)\subseteq\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)$,
+we can notice that each run $\rho'$ of $A'$ on $\mathbf{s}$ that
+is accepting can be supported only by a complete path $p'$, which
+is also in $A$ by construction. $\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)\subseteq\left\llbracket A'\right\rrbracket \left(\mathbf{s}\right)$
+is true as well because we include in $A'$ all the complete paths
+existing in $A$ so every run $p$ of $A$ on $\mathbf{s}$ is also
+a run of $A'$.
+\end{proof}
+
+If a well-behaved eVset-automaton is pruned, it automatically gets
+two other properties.
+\begin{corollary}
+Given a pruned well-behaved eVset-automaton $A$, for every path$p$
+in $A$ $\textrm{LOps}\left(p\right)$ contains no duplicates.
+\end{corollary}
+
+\begin{proof}
+Since $A$ is pruned, $p$ is part of a complete path $p'$. But $A$
+is also well-behaved, so $\textrm{LOps}\left(p'\right)$ contains
+no duplicates, which implies that $\textrm{LOps}\left(p\right)$ has
+no duplicates either.
+\end{proof}
+
+\begin{corollary}
+\label{same-ops}Given a pruned well-behaved eVset-automaton $A$,
+and two states $q,q'$ in $A$, for every pair of paths $p,p'$ between
+$q$ and $q'$ in $A$, $\textrm{Ops}\left(p\right)=\textrm{Ops}\left(p'\right)$.
+\end{corollary}
+
+\begin{proof}
+Since $A$ is pruned, there surely exists a path $p^{0}$ between
+$q_{0}$ and $q$. For the same reason, there exists a path $p^{f}$
+between $q'$ and $q_{f}$. So $q$ and $q'$ appear in two complete
+paths, which we may call $p^{0}pp^{f}$ and $p^{0}p'p^{f}$, that
+differ only in the subpaths between $q$ and $q'$. If $\textrm{Ops}\left(p\right)\neq\textrm{Ops}\left(p'\right)$,
+then one of $p^{0}pp^{f}$ and $p^{0}p'p^{f}$ wouldn't satisfy the
+requirement \ref{enu:well-behaved-1,} of \ref{well-behaved}, making
+$A$ not well-behaved, a contraddiction.
+\end{proof}
+
+\medskip{}
+
+In the following, I describe the constructions that I use in my system
+to support algebraic operations with well-behaved eVset-automata.
+The main advantage of these constructions is that they produce a well-behaved
+eVset-automaton whose size is \emph{polynomial} in the size of the
+input automaton/a. Let us start with projection.
+
+\begin{theorem}
+\noindent \label{ext-proj}Given a well-behaved eVset-automaton $A$,
+the set $X=\textrm{SVars}\left(A\right)$ and a set $Y\subseteq X$,
+a well behaved eVset-automaton $A'$ can be produced in linear time
+such that $\left\llbracket A'\right\rrbracket =\pi_{Y}\left\llbracket A\right\rrbracket $.
+\end{theorem}
+
+\begin{proof}
+Let us consider $A=\left(Q,q_{0},q_{f},\delta\right)$. We can take
+$A'=\left(Q',q_{0}',q_{f}',\delta'\right)$, where:
+
+\begin{itemize}
+\item $Q'=Q$;
+\item $q_{0}'=q_{0}$;
+\item $q_{f}'=q_{f}$;
+\item $\delta'=\left(\delta\setminus\delta^{\textrm{unprojected}}\right)\cup\delta^{\textrm{projected}}$,
+where:
+
+\begin{itemize}
+\item $\delta^{\textrm{unprojected}}=\left\{ \left(q,S,q'\right)\in\delta\mid S\cap\textrm{SVOps}\left(X\setminus Y\right)\neq\textrm{Ø}\right\} $;
+\item $\delta^{\textrm{projected}}=\left\{ \left(q,S',q'\right)\mid\exists\left(q,S,q'\right)\in\delta^{\textrm{unprojected}}:S'=S\setminus\textrm{SVOps}\left(X\setminus Y\right)\right\} $.
+\end{itemize}
+\end{itemize}
+This construction removes all the occurencies of variable operations
+concerning variables excluded from the projection from the transitions
+of $A$. $A'$ is still well-behaved, since we maintain the occurrencies
+of operations concerning the variables on which we project, that continue
+to appear exactly once on each path, and in the right order. Assuming
+the size of $A$, that we may call $n$, as the sum of the sizes of
+$Q$, $\delta$ and $\textrm{SVars}\left(A\right)$, it is easy to
+verify that this construction can be carried out in $O\left(n\right)$.
+Let us show that $\left\llbracket A'\right\rrbracket =\pi_{Y}\left\llbracket A\right\rrbracket $.
+Notice that given a string $\mathbf{s}$, for each accepting run $\rho$
+of $A$ on $\mathbf{s}$, producing an $\mathbf{s}$-tuple $\mu$,
+there is an accepting run $\rho'$ of $A'$ that produces an $\mathbf{s}$-tuple
+$\mu'$, which assigns the same spans as $\mu$ to the variables in
+common with $\mu'$, which by construction belong to $Y$. The path
+$p'$ supporting $\rho'$ is exactly the path obtained by modifying
+the path $p$, that supports $\rho$, by eliminating the operations
+on unprojected variables. This shows that $\pi_{Y}\left\llbracket A\right\rrbracket \subseteq\left\llbracket A'\right\rrbracket $.
+It must also be that $\left\llbracket A'\right\rrbracket \subseteq\pi_{Y}\left\llbracket A\right\rrbracket $,
+because no additional complete paths were added to $A'$.
+\end{proof}
+
+A similar result holds for the union operation.
+
+\begin{theorem}
+\noindent \label{ext-union}Given two well-behaved eVset-automata
+$A$ and $B$ that are union-compatible, a third well-behaved eVset-automaton
+$C$ can be produced in linear time such that $\left\llbracket C\right\rrbracket =\left\llbracket A\right\rrbracket \cup\left\llbracket B\right\rrbracket $.
+\end{theorem}
+
+\begin{proof}
+Let us consider $A$ to be of the form $\left(Q^{A},q_{0}^{A},q_{f}^{A},\delta^{A}\right)$
+and $B$ of the form $\left(Q^{B},q_{0}^{B},q_{f}^{B},\delta^{B}\right)$.
+We can take $C=\left(Q^{C},q_{0}^{C},q_{f}^{C},\delta^{C}\right)$,
+where:
+
+\begin{itemize}
+\item $Q^{C}=Q^{A}\cup Q^{B}\cup\left\{ q_{0}^{C},q_{f}^{C}\right\} $;
+\item $\delta^{C}=\delta^{A}\cup\delta^{B}\cup\left\{ \left(q_{0}^{C},\textrm{Ø},q_{0}^{B}\right),\left(q_{f}^{B},\textrm{Ø},q_{f}^{C}\right),\left(q_{0}^{C},\textrm{Ø},q_{0}^{A}\right),\left(q_{f}^{A},\textrm{Ø},q_{f}^{C}\right)\right\} $.
+\end{itemize}
+In this construction, we allow to go from the initial state of $C$
+to the initial state of either $A$ or $B$, and to go from the accepting
+state of $A$ or that of $B$ to the one of $C$, without any new
+variable operations. Thus, given a string $\mathbf{s}$, $C$ can
+span exactly the $\mathbf{s}-$tuples contained in $\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)\cup\left\llbracket B\right\rrbracket \left(\mathbf{s}\right)$.
+Regarding the complexity of the construction, the operations that
+we perform here are the union of the states sets and transition functions,
+with the addition of a fixed number of new transitions and states.
+Let us consider the size of the input as $n=n_{A}+n_{B}=\left|Q^{A}\right|+\left|\delta^{A}\right|+\left|\textrm{SVars}\left(A\right)\right|+\left|Q^{B}\right|+\left|\delta^{B}\right|+\left|\textrm{SVars}\left(B\right)\right|$.
+If we assume the costs of basic operations, as adding/deleting a state/transitions,
+to be constant, then obtaining $Q^{C}$ takes $O\left(\left|Q^{B}\right|\right)=O\left(n\right)$
+time, and constructing $\delta^{C}$ takes $O\left(\left|\delta^{B}\right|\right)=O\left(n\right)$
+time.
+\end{proof}
+
+In order to show that we can obtain the natural join of two spanners,
+represented by well-behaved eVset-automata, so that the size of the
+correspondent resulting eVset-automaton is polynomial in the size
+of the input automata, we need a few more steps with respect to the
+previous cases. The construction that I present is conceptually very
+similar to that described in \cite{Fagin:2015:DSF:2772377.2699442}
+for the natural join of plain vset-automata. That construction simulates
+running the input automata in parallel, making sure that operations
+on common variables are performed simultaneously. Unfortunately this
+doesn't work in general, as the next example shows.
+
+\begin{figure}
+\centering\begin{tikzpicture}[shorten >=1pt,node distance=3cm,on grid,auto] \node[state,initial]  (q_0)                      {$q_0$}; \node[state]          (q_1) [right=of q_0] {$q_1$}; \node[state]          (q_2) [right=of q_1] {$q_2$}; \node[state,accepting](q_3) [right=of q_2] {$q_3$}; \path[->] (q_0) edge              node        {$\left\{ y\vdash\right\} $} (q_1) 		edge [loop above] node        {$\varSigma$} () 	  (q_1) edge 	          node        {$\left\{ x\vdash\right\} $} (q_2) 		edge [loop above] node        {a} (q_2) 	  (q_2) edge              node        {$\left\{ \dashv x,\dashv y\right\} $} (q_3) 		edge [loop above] node        {b} () 	  (q_3) edge [loop above] node        {$\varSigma$} (); \end{tikzpicture}
+
+\caption{\label{evset-2}An eVset-automaton $B$.}
+
+\end{figure}
+
+\begin{example}
+Consider the eVset-automaton $A$ from Figure \ref{fig:evset}, the
+eVset-automaton $B$ from Figure \ref{evset-2} and the string $\mathbf{s}=$'b'.
+$\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)$ contains
+the $\mathbf{s}$-tuple $\mu$ such that $\mu\left(x\right)=\left[0,0\right\rangle $
+and $\mu\left(y\right)=\left[0,1\right\rangle $. We also have that
+$\mu\in\left\llbracket B\right\rrbracket \left(\mathbf{s}\right)$.
+If we attempt to run $A$ and $B$ in parallel on $\mathbf{s}$, we
+won't be able to span $\mu$ because the two automata disagree on
+the order of the operations $x\vdash$ and $y\vdash$. Thus, we don't
+get $\left\llbracket A\right\rrbracket \bowtie\left\llbracket B\right\rrbracket \left(\mathbf{s}\right)$
+as a result of the execution.
+\end{example}
+
+For the construction to work, the input automata must be modified
+in some way. In my system, I convert them into a particular form of
+eVset-automata, that I call \emph{operation-closed}. The resulting
+automata have generally bigger size of the originals, but still they
+allow to keep the size of the result of the join construction to be
+polynomial in the size of the input. The definition of an operation-closed
+eVset-automaton follows.
+\begin{definition}
+Given a eVset-automaton $A$, $A$ is operation-closed if, for every
+pair of states $q,q'$ in $A$, whenever there exists an operation-only
+path $p$=$t_{1},\ldots,t_{n}$ between $q$ and $q'$, then there
+exists a transition in $A$ of the form $\left(q,\bigcup_{i=1}^{n}\textrm{Ops}\left(t_{i}\right),q'\right)$.
+\end{definition}
+
+As the next proposition states, given a well-behaved eVset-automaton,
+we can always find an equivalent operation-closed well-behaved eVset-automaton,
+whose size is polynomial in the size of the original. This ensures
+the applicability of the join construction.
+\begin{proposition}
+\label{well-b-to-op-clos}Given a well-behaved eVset-automaton $A$,
+there exists an operation-closed well-behaved eVset-automaton $A'$
+such that $\left\llbracket A'\right\rrbracket =\left\llbracket A\right\rrbracket $.
+Moreover, the size of $A'$ is cubic in the size of $A$.
+\end{proposition}
+
+\begin{proof}
+Let us consider $A$ to be of the form $\left(Q,q_{0},q_{f},\delta\right)$.
+Without loss of generality, we can assume that $A$ is pruned. Then
+we can take $A'=\left(Q',q_{0}',q_{f}',\delta'\right)$, where:
+
+\begin{itemize}
+\item $Q'=Q$;
+\item $q_{0}'=q_{0}$;
+\item $q_{f}'=q_{f}$;
+\item $\delta'=\delta\cup\left\{ \left(q,\bigcup_{i=1}^{n}\textrm{Ops}\left(t_{i}\right),q'\right)\mid\exists p=t_{1},\ldots,t_{n}\in\textrm{Paths}\left(A\right):p=p_{q}^{q'}\textrm{, \ensuremath{p} is operation-only}\right\} $.
+\end{itemize}
+This construction does nothing but including in $A'$ the transitions
+that are missing in $A$ to be operation-closed . We now show that,
+for every string $\mathbf{s}\in\varSigma^{\ast}$, $\left\llbracket A\right\rrbracket '\left(\mathbf{s}\right)=\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)$.
+To see that $\left\llbracket A\right\rrbracket '\left(\mathbf{s}\right)\subseteq\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)$,
+consider an accepting run $\rho'$ of $A'$ on $\mathbf{s}$, that
+returns an $\mathbf{s}$-tuple $\mu$. This run is supported by a
+complete path $p'$ in $A'$. We can always find a complete path $p$
+in $A$ that supports a run $\rho$ of $A$ on $\mathbf{s}$, which
+returns $\mu$ as well. To obtain $p$, we substitute every operations
+transition $t$ in $p'$ that doesn't belong to $\delta$ with a operation-only
+path $p''$ in $A$ such that $\textrm{LOps}\left(p''\right)=\textrm{LOps}\left(t\right)$
+and $p''$ is between the source and destination states of $t$. This
+is always possible by construction of $A'$. It is easy to verify
+that the run $\rho$ of $A$ on $\mathbf{s}$, supported by $p$,
+returns indeed $\mu$. $\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)\subseteq\left\llbracket A\right\rrbracket '\left(\mathbf{s}\right)$
+is also true, because we include all the transitions belonging to
+$\delta$ in $\delta'$. The fact that the size of $A'$ is cubic
+in the size of $A$ follows from the fact that we add $O\left(\left|Q^{2}\right|\right)$
+new transitions of size $O\left(\textrm{SVars}\left(A\right)\right)$
+to $\delta$ to form $\delta'$.
+\end{proof}
+
+This construction is the reason why extended vset-automata were used
+for the runtime system instead of plain vset-automata. When we construct
+an operation-closed automaton, starting from an automaton that does
+not have this property, the size of the resulting transition function
+is likely to explode (although it will stay polynomial in the size
+of the original automaton). The ability of performing multiple variable
+operations in a single transition allows for a more compact representation,
+that is also easier to manipulate. I now present a product construction
+that simulates running two operation-closed well-behaved eVset-automata
+in parallel, which is very similar to that presented in \cite{Fagin:2015:DSF:2772377.2699442}
+for vset-automata. Notice that we need the input automata to be operation-closed,
+in order for the construction to give correct results.
+\begin{definition}
+Given two well-behaved operation-closed eVset-automata $A=\left(Q_{A},q_{A}^{0},q_{A}^{f},\delta_{A}\right)$
+and $B=\left(Q_{B},q_{B}^{0},q_{B}^{f},\delta_{B}\right)$, their
+product is a eVset-automaton $C=\left(Q,q^{0},q^{f},\delta\right)$,
+where:
+
+\begin{itemize}
+\item $Q=Q_{A}\times Q_{B}$;
+\item $q^{0}=\left\langle q_{A}^{0},q_{B}^{0}\right\rangle $;
+\item $q^{f}=\left\langle q_{A}^{f},q_{B}^{f}\right\rangle $;
+\item $\delta$ has the following transitions:
+
+\begin{itemize}
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,\sigma,\left\langle q'_{A},q'_{B}\right\rangle \right)$
+whenever $\sigma\in\varSigma$, $\left(q_{A},\sigma,q'_{A}\right)\in\delta_{A}$
+and $\left(q_{B},\sigma,q'_{B}\right)\in\delta_{B}$;
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,S_{A}\cup S_{B},\left\langle q'_{A},q'_{B}\right\rangle \right)$
+whenever $\left(q_{A},S_{A},q'_{A}\right)\in\delta_{A}$, $\left(q_{B},S_{B},q'_{B}\right)\in\delta_{B}$
+and $\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{A}=\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{B}$;
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,S_{A},\left\langle q'_{A},q{}_{B}\right\rangle \right)$
+whenever $\left(q_{A},S_{A},q'_{A}\right)\in\delta_{A}$ and $\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{A}=\textrm{Ø}$;
+\item $\left(\left\langle q_{A},q_{B}\right\rangle ,S_{B},\left\langle q_{A},q'_{B}\right\rangle \right)$
+whenever $\left(q_{B},S_{B},q'_{B}\right)\in\delta_{B}$ and $\left(\textrm{SVOps}\left(A\right)\cap\textrm{SVOps}\left(B\right)\right)\cap S_{B}=\textrm{Ø}$.
+\end{itemize}
+\end{itemize}
+We may write $C=A\times B$.
+\end{definition}
+
+We can now state the following theorem.
+\begin{theorem}
+\label{ext-join}Given two well-behaved eVset-automata $A$ and $B$,
+and given an eVset-automaton $C$ such that $C=A\times B$, then $\left\llbracket C\right\rrbracket =\left\llbracket A\right\rrbracket \bowtie\left\llbracket B\right\rrbracket $.
+Moreover, the size of $C$ is polynomial in the size of $A$ and $B$.
+\end{theorem}
+
+\begin{proof}
+This proof is similar to the proof for the analogue construction for
+plain vset-automata in \cite{Fagin:2015:DSF:2772377.2699442}. Without
+loss of generality, we can assume that $A$ and $B$ are operation
+closed. If this was not the case, we could always obtain equivalent
+automata that are operation closed, in polynomial time, because of
+Proposition \ref{well-b-to-op-clos}. To show that $\left\llbracket C\right\rrbracket \subseteq\left\llbracket A\right\rrbracket \bowtie\left\llbracket B\right\rrbracket $,
+we can decompose a run of $C$ on a string $\mathbf{s}$ into two
+consistent runs of $\rho_{A}$ of $A$ and $\rho_{B}$ of $B$. Two
+runs $\rho$, $\rho'$, with supporting paths $p$ and $p'$ respectively,
+are consistent with each other if, for every pair of operations $o,o'$
+belonging both to $\textrm{Ops}\left(p\right)$ and $\textrm{Ops}\left(p'\right)$,
+$o$ appears before $o'$ in $\textrm{LOps}\left(p\right)$ if and
+only if the same holds in $\textrm{LOps}\left(p'\right)$. Since a
+run of $C$ represents two parallel runs of $A$ and $B$ by construction,
+the decomposition aims to isolate the two individual runs of $A$
+and $B$. The details of this decomposition are not difficult to figure
+out, and are omitted. To show that $\left\llbracket A\right\rrbracket \bowtie\left\llbracket B\right\rrbracket \subseteq\left\llbracket C\right\rrbracket $,
+consider a string $\mathbf{s}$, a $\mathbf{s}$-tuple $\mu_{A}\in\left\llbracket A\right\rrbracket \left(\mathbf{s}\right)$
+and a $\mathbf{s}$-tuple $\mu_{B}\in\left\llbracket B\right\rrbracket \left(\mathbf{s}\right)$
+that assigns the same spans as $\mu_{A}$ to the variables they have
+in common. Given the $\mathbf{s}$-tuple $\mu$ that yelds all the
+variable assignments of $\mu_{A}$ and $\mu_{B}$, we need to find
+a run $\rho$ of $C$ on $\mathbf{s}$ that returns $\mu$. Let us
+call $\rho_{A}\in\textrm{ARuns}\left(A,\mathbf{s}\right)$ and $\rho_{B}\in\textrm{ARuns}\left(B,\mathbf{s}\right)$
+the runs that return $\mu_{A}$ and $\mu_{B}$, respectively. We can
+obtain $\rho$ by combining $\rho_{A}$ and $\rho_{B}$. For this
+construction to work, $\rho_{A}$ and $\rho_{B}$ need to be consistent
+on the order of the variable operations they perform. Since $A$ and
+$B$ are well-behaved and operation-closed, $\rho_{A}$ and $\rho_{B}$
+can always be selected so that they are consistent.
+\end{proof}
+
+Now, let us reason on the relative expressive power of eVset-automata
+with respect to plain vset-automata. Actually, they are equivalent,
+as the following lemmas show.
+\begin{lemma}
+\noindent \label{eVset-to-plain}Given an eVset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$,
+$A$ can be converted into a vset-automaton $A'$, such that $\left\llbracket A'\right\rrbracket =\left\llbracket A\right\rrbracket $,
+in polynomial time, and in a well-behavedness preserving manner\footnote{Note that I didn't define well-behavedness in the case of a plain
+vset-automaton. Nonetheless, the idea underlying the definition for
+extended vset-automata remains unchanged and the actual definition
+for a standard vset-automaton isn't difficult to figure out.}.
+\end{lemma}
+
+\begin{proof}
+Without loss of generality, we consider an ordering of the symbols
+in $\textrm{SVOps}\left(A\right)$ of the following form:
+\[
+x\vdash,\ldots,y\vdash,x\dashv,\ldots,y\dashv
+\]
+ In this ordering, all insertion operations come before the deletion
+operations. Let us define $o\prec o'$, with $o,o'\in\textrm{SVOps}\left(A\right)$,
+if $o$ comes before $o'$ (\emph{not} if they are equal) in the chosen
+ordering. Consider $A'=$$\left(Q',q_{0}',q_{f}',\delta'\right)$,
+with $\textrm{SVars}\left(A'\right)=\textrm{SVars}\left(A\right)$,
+whose components are defined as follows:
+
+\begin{itemize}
+\item $Q'=Q\cup Q^{\textrm{ops}}\cup Q^{\textrm{Ø}}$, where:
+
+\begin{itemize}
+\item $Q^{\textrm{ops}}=\left\{ q_{q',o,q''}\mid\exists\left(q',S,q''\right)\in\delta:o\in S\right\} $;
+\item $Q^{\textrm{Ø}}=\left\{ q_{\textrm{q',Ø,q''}}\mid\exists\left(q',\textrm{Ø},q''\right)\in\delta\right\} $;
+\end{itemize}
+\item $q_{0}'=q_{0};$
+\item $q_{f}'=q_{f}$;
+\item $\delta'=\left(\delta\setminus\delta^{S}\right)\cup\delta^{\textrm{ops}}\cup\delta^{\textrm{Ø}}\cup\delta^{\varepsilon}$,
+where:
+
+\begin{itemize}
+\item $\delta^{S}=\left\{ \left(q,S,q'\right)\in\delta\right\} ;$
+\item $\delta^{\textrm{ops}}=\left\{ \left(q_{q',o,q''},o,q_{q',o',q''}\right)\mid\exists\left(q',S,q''\right)\in\delta:\left(o,o'\epsilon S\land o\prec o'\land\forall o''\in S:o\nprec o''\nprec o'\right)\right\} \cup\left\{ \left(q_{q',o,q''},o,q''\right)\mid\exists\left(q',S,q''\right)\in\delta:\left(o\epsilon S\land\forall o'\in S:o\nprec o'\right)\right\} $;
+\item $\delta^{\textrm{Ø}}=\left\{ \left(q_{\textrm{q',Ø,q''}},\varepsilon,q''\right)\mid\exists\left(q',\textrm{Ø},q''\right)\in\delta\right\} ;$
+\item $\delta^{\varepsilon}=\left\{ \left(q,\varepsilon,q'\right)\mid\left(\exists\left(q',o,q''\right)\in\delta^{\textrm{ops}}:\forall\left(q',o',q''\right)\in\delta^{\textrm{ops}}:o'\nprec o\right)\lor\left(\exists\left(q',\varepsilon,q''\right)\in\delta^{\textrm{Ø}}\right)\right\} $.
+\end{itemize}
+\end{itemize}
+This construction expands the transitions of $A$ that are labeled
+with a set of variable operations into a sequence of transitions performing
+one operation at a time, taking care of putting the insertion operations
+before the deletion ones. This construction clearly preserves well-behavedness,
+because each complete path $p'$ in $A'$ is obtained from a path
+$p$ in $A$, preserving the original operations of $p$ and ensuring
+a correct order of their appearence (if $A$ is well-behaved). Each
+sequence starts with an $\varepsilon$-transition. This is not necessary
+in principle, but it allows to reduce the complexity of the formulation.
+The construction also substitutes transitions labeled with the empty
+set with ordinary $\varepsilon-$transitions. To prove equivalence
+between $A$ and $A'$ it is sufficient to notice that for every string
+$\mathbf{s}$, every run belonging to$\textrm{ARuns}\left(A,\mathbf{s}\right)$
+can be put in correspondence with a run belonging to $\textrm{ARuns}\left(A',\mathbf{s}\right)$,
+and viceversa. Indeed, if we start from $\rho\in\textrm{ARuns}\left(A,\mathbf{s}\right)$
+we can obtain a run $\textrm{\ensuremath{\rho}'\ensuremath{\in}ARuns}\left(A',\mathbf{s}\right)$
+that spans the same $\mathbf{s}$-tuple $\mu$. It is sufficient to
+expand configuration pairs in $\rho$ whose current states are linked
+in $A$ by a transition $t$, which we expand in $A'$, with a series
+of configurations that let us perform the set of variable operations
+in $t$ one at a time. This is always possible by construction of
+$A'$. If we start from $\textrm{\ensuremath{\rho}'\ensuremath{\in}ARuns}\left(A',\mathbf{s}\right)$,
+we can obtain an equivalent run $\rho\in\textrm{ARuns}\left(A,\mathbf{s}\right)$
+in the opposite way, by compressing consecutive configurations. The
+details are omitted.
+
+\noindent It remains to show that the construction of $A'$ can be
+carried in polynomial time. Let us refer to the size of $A$ as $n$.
+We consider $n=\left|Q\right|+\left|\delta\right|+\left|\textrm{SVars}\left(A\right)\right|$.
+With the usual assumptions, adding a new transition, removing a state
+and removing a transition take constant time. Thus, expanding a single
+transition of $A$ takes $O\left(\left|\textrm{SVars}\left(A\right)\right|\right)=O\left(n\right)$
+time. Expanding every transition that is needed will then take $O\left(\left|\textrm{SVars}\left(A\right)\right|\cdot\left|\delta\right|\right)=O\left(n^{2}\right)$
+time.
+\end{proof}
+
+The opposite direction is also true.
+\begin{lemma}
+\noindent Given a vset-automaton $A=\left(Q,q_{0},q_{f},\delta\right)$,
+an eVset-automaton $A'$ can be found such that $\left\llbracket A'\right\rrbracket =\left\llbracket A\right\rrbracket $,
+in polynomial time, and in a well-behavedness preserving manner.
+\end{lemma}
+
+\begin{proof}
+In this case it is sufficient to replace every transition $t$ performing
+a variable operation $o$ in $A$ with an operations transition $t'$
+in $A'$ such that $\textrm{Ops}\left(t'\right)=\left\{ o\right\} $
+and $\textrm{LOps}\left(t'\right)=o$ . $\varepsilon$-transitions
+can be replaced by operations transitions with an empty set. With
+the usual assumptions, this construction can be obtained in linear
+time. It is easy to verify equivalence, and that well-behavedness
+is preserved, thus the details are omitted.
+\end{proof}
+
+These results are interesting. In particular, Lemma \ref{eVset-to-plain}
+ensures that any well-behaved eVset-automaton, equipped with the string
+equality selection operator, can indeed represent an AQL core fragment.
+What about the inverse? In order for the runtime system to have full
+applicability, we need to prove that an AQL core fragment can be represented
+by a well-behaved eVset automaton, extended with string equality selection.
+According to Definition \ref{core-def}, an AQL core fragment is a
+set of regex formulas combined by using the operators described in
+Subsection \ref{subsec:Algebras-of-Spanners}. To achieve our goal,
+the first step is to prove that a regex formula can always be converted
+into a well-behaved eVset-automaton.
+\begin{theorem}
+\label{regex-to-well}Given a regex formula $\gamma$, a well-behaved
+eVset-automaton $A$ can be found, in polynomial time, such that $\left\llbracket A\right\rrbracket =\left\llbracket \gamma\right\rrbracket $.
+\end{theorem}
+
+Let us call the class of well-behaved eVset-automata $\textrm{VA}_{\textrm{WESet}}$.
+By combining Theorems \ref{regex-to-well}, \ref{ext-union} and \ref{ext-join}
+and Lemma \ref{eVset-to-plain}, we can state the following theorem.
+\begin{theorem}
+$\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{VA}_{\textrm{WESet}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{VA}_{\textrm{WESet}}^{\left\{ \pi,\varsigma^{=}\right\} }\right\rrbracket $.
+\end{theorem}
+
+\begin{proof}
+Theorem \ref{regex-to-well} tells us that $\left\llbracket \textrm{RGX}\right\rrbracket \subseteq\left\llbracket \textrm{VA}_{\textrm{WESet}}\right\rrbracket $.
+Thus we can state that
+
+\begin{equation}
+\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket \subseteq\left\llbracket \textrm{VA}_{\textrm{WESet}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket
+\end{equation}
+
+Moreover, Lemma \ref{eVset-to-plain} and Theorem \ref{RGX=00003DSET}
+imply that
+
+\begin{equation}
+\left\llbracket \textrm{VA}_{\textrm{WESet}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket \subseteq\left\llbracket \textrm{VA}_{\textrm{set}}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket =\left\llbracket \textrm{RGX}^{\left\{ \cup,\pi,\bowtie,\varsigma^{=}\right\} }\right\rrbracket
+\end{equation}
+
+Finally, Theorems \ref{ext-union} and \ref{ext-join} justify the
+right-hand equality in the statement of this theorem.
+\end{proof}
+
+We can also formulate a modified version of the Core Simplification
+Lemma (Lemma \ref{Core-simplif}). The proof is the same as that of
+the original lemma, which can be found in \cite{Fagin:2015:DSF:2772377.2699442},
+and is omitted here.
+\begin{lemma}
+Every core spanner can be defined by an expression of the form
+
+\begin{equation}
+\pi_{V}SA
+\end{equation}
+
+\noindent where:
+
+\begin{itemize}
+\item $A$ is a well-behaved eVset-automaton;
+\item $V\subseteq\textrm{SVars}\left(A\right)$;
+\item $S$ is a sequence of string selections $\varsigma_{x,y}^{=}$, for
+$x,y\in\textrm{SVars}\left(A\right)$.
+\end{itemize}
+\end{lemma}
+
+\section{Methods for NFA execution}
+
+\subsection{\label{subsec:The-Thompson-Approach}The Thompson Approach}
+
+Given a regular expression $r$, its corresponding NFA $A$ and a
+string $\mathbf{s}$, the Thompson algorithm will try all the feasible
+runs of $A$ on $\mathbf{s}$ at the same time. More precisely, when
+the string pointer is on a given character, there is a set of current
+states, each representing the advancement in a feasible run. The set
+of outgoing transitions for each state is examined. All $\varepsilon$-transitions
+are fired iteratively and, subsequently, any transition labeled with
+the character corresponding to the one pointed by the string pointer
+is fired. This produces a new set of states and the execution continues
+as descibed. $\mathbf{s}$ is matched by $A$ if, at the end of the
+execution, all the symbols of $\mathbf{s}$ have been consumed and
+at least one state in the current state set is an accepting state.
+It can be shown that the time complexity of this approach is $O\left(mn\right)$,
+where $m$ is the size of $r$.
+
+Usually, we accept \emph{partial matchings,} meaning that we don't
+require to reach the end of a string to produce a matching, instead
+we only require that one state set, among those obtained during the
+execution, contains the final state. This is useful for implementing
+\emph{unanchored matchings. }Partial matchings pose the problem of
+ambiguous matches. This is why concrete algorithms implement a policy
+to discriminate among multiple possible matchings (e.g. greedy leftmost).
+For more information on the Thompson approach, see \cite{cox_regular_2007}.
+
+\begin{figure}
+\centering
+
+\subfloat[An $\varepsilon$-NFA $A$.]{\centering \begin{tikzpicture}[shorten >=1pt,node distance=3cm,on grid,auto] \node[state,initial]  (q_0)                      {$q_0$}; \node[state]          (q_1) [above right=of q_0] {$q_1$}; \node[state]          (q_2) [below right=of q_0] {$q_2$}; \node[state,accepting](q_3) [below right=of q_1] {$q_3$}; \path[->] (q_0) edge              node        {a} (q_1) 		edge              node [swap] {$\varepsilon$} (q_2) 	  (q_1) edge [bend left]  node        {a} (q_3) 		edge [bend right] node        {b} (q_3) 		edge [loop above] node        {a} () 	  (q_2) edge              node [swap] {c} (q_3) 		edge [loop below] node        {a} () 	  (q_3) edge [loop right] node        {a} (); \end{tikzpicture}
+
+\label{thompson-aut}}
+
+\medskip{}
+
+\subfloat[The execution of $A$ on 'aaab' according to the Thompson approach.]{\centering \begin{tikzpicture}[node distance=2.5cm,auto]     \node (one) [align=left] {$\underline{a}aab$\\$\left\{ q_{0}\right\} $};     \node (two) [right of=one,align=left] {$a\underline{a}ab$\\$\left\{ q_{1}, q_{2} \right\} $};     \node (three) [right of=two,align=left] {$aa\underline{a}b$\\$\left\{ q_{1}, q_{2}, q_{3} \right\} $};     \node (four) [right of=three,align=left] {$aaa\underline{b}$\\$\left\{ q_{1}, q_{2}, q_{3} \right\} $};     \node (five) [right of=four,align=left] {aaab\underline{ }\\$\left\{q_{3} \right\} $};     \path[->] (one) edge node {} (two);     \path[->] (two) edge node {} (three);     \path[->] (three) edge node {} (four);     \path[->] (four) edge node {} (five); \end{tikzpicture}
+
+\label{thompson-exec}}\caption{Example of the Thompson algorithm.}
+\end{figure}
+
+\begin{example}
+The automaton $A$ of Figure \ref{thompson-aut} matches the string
+$\mathbf{s}=$'aaab'. Figure \ref{thompson-exec} illustrates the
+execution of $A$ on $\mathbf{s}$. Notice how runs that end up in
+the same state are naturally merged. For instance, if we are in $q_{1}$
+and we read an $a$ we can go to $q_{3}$, but we can also read an
+$a$ while in $q_{3}$, remaining in that state. Nonetheless, in the
+cases where both $q_{1}$ and $q_{3}$ are in the current state set,
+$q_{3}$ appears only once in the next. This is because the two originally
+distinct runs are now equivalent. This lets the algorithm keep a reduced
+list of current states, and justifies the runtime complexity presented.
+\end{example}
+
+One interesting implementation of the Thompson algorithm views an
+NFA as a program that can be executed by a virtual machine on a string.
+This method is described in \cite{cox_regular_2009}.
+
+\subsubsection{The Virtual Machine Implementation}
+
+In this implementation, a NFA (or a regex) is converted into a program,
+written in a simple assembly language with few instructions. The basic
+instructions are character match (CHAR), string match (MATCH), and
+control flow instructions (SPLIT and JUMP). A virtual machine is provided,
+which treats each concurrent run on a string $\mathbf{s}$ as a conceptual
+thread. The virtual machine advances all the threads in lockstep,
+in the spirit of the Thompson approach. Actual implementations can
+enforce a policy for ambiguous matches. Different threads that reach
+the same instruction in a program are merged.
+
+The main advantage of this method is that it is easy to enrich the
+assembly language with new instructions, in order to support new features.
+To execute new instructions, modyifing the virtual machine is required.
+For instance, capturing groups can be implemented by equipping each
+thread with an array of saved pointers that are grouped by two: the
+first element would point to the beginning of a span of text and the
+second one would point to its end. Then, a SAVE instruction could
+be added, that would make the virtual machine record the current position
+in the input into the pointer specified by the instruction.
+
+\begin{figure}
+\centering
+
+\subfloat[A program for the regex $a^{+}b^{\ast}$.]{\begin{tikzpicture} \matrix [ampersand replacement=\&]{ 	\node {0}; \& \node {CHAR a};    \\ 	\node {1}; \& \node {SPLIT 0, 2};\\ 	\node {2}; \& \node {SPLIT 3, 5};\\ 	\node {3}; \& \node {CHAR b};    \\ 	\node {4}; \& \node {JUMP 2};    \\ 	\node {5}; \& \node {MATCH};     \\ 	\node {}; \& \node {};     \\ 	\node {}; \& \node {};     \\ 	\node {}; \& \node {};     \\
+}; \end{tikzpicture}
+
+\label{prog}}\subfloat[A program for the regex $\left(a^{+}\right)\left(b^{\ast}\right)$.]{\begin{tikzpicture} \matrix [ampersand replacement=\&]{ 
+	\node{0}; \& \node {SAVE 0};    \\ 	\node{1}; \& \node {CHAR a};    \\ 	\node{2}; \& \node {SPLIT 1, 3};\\ 	\node{3}; \& \node {SAVE 1};    \\ 	\node{4}; \& \node {SAVE 2};    \\ 	\node{5}; \& \node {SPLIT 6, 8};\\ 	\node{6}; \& \node {CHAR b};    \\ 	\node{7}; \& \node {JUMP 5};    \\ 	\node{8}; \& \node {SAVE 3};    \\ 	\node{9}; \& \node {MATCH};     \\
+}; \end{tikzpicture}
+
+\label{prog-capt}}\caption{Example programs.}
+
+\end{figure}
+
+\begin{example}
+The program shown in Figure \ref{prog} can be used to match the regular
+expression $a^{+}b^{*}$. SPLIT instructions explictly divide a thread
+into two, telling each of the generated threads which position in
+the program to reach. The program in Figure \ref{prog-capt} matches
+the same expression as the one in Figure \ref{prog}, but it retains
+the substrings matching the subexpressions between parentheses too.
+
+\begin{figure}
+\centering\begin{tikzpicture} \matrix [ampersand replacement=\&]{ 
+	\node{0}; \& \node {SPLIT 1, 8};    \\ 	\node{1}; \& \node {CHAR a};    \\ 	\node{2}; \& \node {SPLIT 1, 3};\\ 	\node{3}; \& \node {SPLIT 4, 6};    \\ 	\node{4}; \& \node {CHAR a};    \\ 	\node{5}; \& \node {JUMP 7};\\ 	\node{6}; \& \node {CHAR b};    \\ 	\node{7}; \& \node {JUMP 12};    \\ 	\node{8}; \& \node {SPLIT 9, 11};    \\ 	\node{9}; \& \node {CHAR a};     \\ 	\node{10}; \& \node {JUMP 8};     \\ 	\node{11}; \& \node {CHAR c};     \\ 	\node{12}; \& \node {SPLIT 13, 15};     \\ 	\node{13}; \& \node {CHAR a};     \\ 	\node{14}; \& \node {JUMP 12};     \\ 	\node{15}; \& \node {MATCHS};     \\
+}; \end{tikzpicture}
+
+\label{prog-aut}\caption{A program for the automaton $A$ of Figure \ref{thompson-aut}.}
+\end{figure}
+\end{example}
+
+\begin{example}
+The program shown in Figure \ref{prog-aut} corresponds to the automaton
+$A$ of Figure \ref{thompson-aut}. Notice how the $\varepsilon$-transition
+between $q_{0}$ and $q_{2}$ is automatically omitted from the program.
+Blocks of SPLIT instructions correspond to multiple outgoing transitions
+from a state. JUMP instructions are used to merge execution branches. 
+\end{example}
+
+\subsection{Ordered Binary Decision Diagrams}
+
+Ordered Binary Decision Diagrams (OBDDs) are tree-like structures,
+that can be used to efficiently represent, manipulate and evaluate
+boolean formulas. One of their advantages is that first order logic
+quantifiers can be easily supported by performing a series of operations
+that change the structure of an OBDD opportunely. In \cite{Yang:2012:FSE:2396556.2396594},
+it is shown how to encode NFAs as OBDDs, with support for submatch
+extraction. Then, a procedure for evaluating a string by using an
+OBDD representation of a NFA is described.
+
+Given a regular expression $r$ of size $m$ and a string of size
+$n$, the runtime cost of this approach is shown to be between $O\left(m\right)$
+and $O\left(mn\right)$. The authors of \cite{Yang:2012:FSE:2396556.2396594}
+suggest that this approach performs best when matching patterns are
+combined togheter.
+
+\subsection{Kleenex}
+
+A \emph{transducer} is a kind of finite state machine whose main difference
+with an NFA is that it has an \emph{output tape}. Thus, the transition
+function of a transducer is enriched with \emph{output actions.} Output
+actions can be used to write to the output tape, thus providing a
+way to support submatch extraction.
+
+In \cite{Grathwohl:2016:KCN:2914770.2837647}, the Kleenex language
+is presented. It can be used to specify context free grammars, representing
+regular expressions with capturing groups. A Kleenex program is converted
+to a transducer, which is in turn decomposed into an \emph{oracle
+machine} and an \emph{action machine}. The oracle machine traverses
+an input string trying to match it (according to a disambiguation
+policy), while the action machine performs the ouput actions.
+
+\section{Implementation}
+
+In the system, core spanners are represented by well-behaved eVset-automata,
+for which a set of string equality constraints can be specified on
+pairs of their span variables. Moreover, the implementation fixes
+$\textrm{SVars}=\mathbb{N}$. Hence, span variables are identified
+by nonnegative integers. The system can be used in two modes: 
+\begin{description}
+\item [{compilation}] reads an AQL core fragment specification and produces
+an equivalent core spanner representation;
+\item [{evaluation}] evaluates a core spanner representation on a series
+of text documents, returning, for each of them, a $\left(V,\mathbf{s}\right)$-relation
+as output.
+\end{description}
+
+\subsection{Compilation Mode}
+
+The compilation mode allows to easily compose spanner representations
+to obtain more complex ones. The composition process is performed
+according to an input AQL core fragment, written in an ad-hoc syntax
+that linearizes its operator tree. \emph{// I will put an example. }
+
+The compilation mode supports all the operators described in Section
+\ref{subsec:Algebras-of-Spanners}. For projection, union and natural
+join, the constructions described in Theorems \ref{ext-proj}, \ref{ext-union}
+and \ref{ext-join} are respectively used (keep in mind that core
+spanners are represented by well-behaved eVset-automata). String equality
+selection operations are simply reported in the resulting representation,
+for use of the evaluation engine. \emph{// Need to describe the 'isFollowedBy'
+join here.}
+
+\subsection{Evaluation Mode}
+
+\begin{algorithm}
+$E\longleftarrow\left\{ \textrm{core spanner representation}\right\} $
+
+for document in inputDirectory do
+
+begin
+\begin{enumerate}
+\item $\left\{ \textrm{Read document into main memory}\right\} $
+\item $R\longleftarrow E\left(\textrm{document}\right)$
+\item $\left\{ \textrm{Write \ensuremath{R} to disk}\right\} $
+\end{enumerate}
+\caption{\label{alg:The-execution-algorithm}The execution algorithm of a core
+spanner representation on a set of input documents.}
+
+\end{algorithm}
+
+The evaluation mode requires as input a core spanner representation
+and the path of the directory containing the text documents to span.
+The execution algorithm, represented by Algorithm \ref{alg:The-execution-algorithm},
+is similar to that of SystemT, whose high-level structure is described
+by Algorithm \ref{alg:Annotating-all-local}. In particular, Algorithm
+\ref{alg:The-execution-algorithm} conforms to document-at-a-time
+processing as well. On the other hand, the results are written to
+a separate file, instead of annotating the input documents, as in
+Algorithm \ref{alg:Annotating-all-local}.
+
+\medskip{}
+
+The engine that evaluates a spanner representation on a document draws
+inspiration from the Thompson approach, described in Subsection \ref{subsec:The-Thompson-Approach}.
+The procedure defined by this approach can't be used as is. This is
+because when evaluating an eVset-automaton $A$ on a string $\textrm{\ensuremath{\mathbf{s}}}$,
+we are interested in all the $\mathbf{s}$-tuples defined by the runs
+belonging to $\textrm{ARuns}\left(A,\mathbf{s}\right)$, instead of
+a single match. Subsequently, the implementation of the engine doesn't
+discard any of the feasible runs in $\textrm{ARuns}\left(A,\mathbf{s}\right)$.
+
+Initially, the implementation followed the Virtual Machine method:
+a core spanner representation was converted into an NFA program and
+executed by a modified implementation of the virtual machine. However
+this implementation revealed itself to be slow. Thus, the engine was
+changed so that it directly uses input eVset-automata: in this implementation,
+a set of current runs is maintained, and each run is advanced by following
+the outgoing transitions of its current state. All the runs are advanced
+in lockstep. There's another difference with the original Thompson
+approach: operation transitions, which may be considered as a special
+kind of $\epsilon$-transitions, aren't followed iteratively, thus
+different runs may point to different positions of the input string
+at the same moment.
+
+\chapter{Experiments}
+
+After the development of the system, an experimental validation phase
+followed. The experiments were designed to test whether the evaluation
+approach described in this paper (converting a query into a well-behaved
+eVset-automaton and using a modified regular expression engine to
+find matches) brings performance benefits over the approach of SystemT.
+In order to make a fair comparison, a subsystem that imitates SystemT
+was developed. This systems doesn't support the optimization techniques
+described in \cite{4497502,Krishnamurthy:2009:SSD:1519103.1519105}.
+
+\section{A Subsystem for the Algebraic Approach}
+
+The subsystem implements the operators of the algebra as operations
+on $\left(V,\mathbf{s}\right)$-relations, that return a new $\left(V,\mathbf{s}\right)$-relation.
+\emph{// To be continued...}
+
+\section{Queries}
+
+\emph{// Description of the benchmark queries will go here...}
+
+\section{Experimental Results}
+
+\emph{// Just the diagrams for now. Discussion will go here.}
+
+\begin{figure}
+\pgfplotstableread[row sep=\\,col sep=&]{     interval      & baseT \\     Action        & 13.03 \\     Aspect        & 16.75 \\     Attribute     & 18.67 \\     Genre         & 20.23 \\     MovieKeyword  & 13.30 \\     Name          & 12.38 \\     PlotClue      & 16.83 \\     Role          & 14.32 \\     RoleClue 	  & 12.25 \\     Sentiment     & 15.45 \\     Title         & 12.03 \\     }\mydata
+\begin{tikzpicture}     \begin{axis}[             ybar,             bar width=.5cm,             width=\textwidth,             height=.5\textwidth,             symbolic x coords={Action,Aspect,Attribute,Genre,MovieKeyword,Name,PlotClue,Role,RoleClue,Sentiment,Title},             xtick=data, 	    xticklabel style={ 		inner sep=0pt, 		anchor=north east, 		rotate=45 	    },             nodes near coords,             nodes near coords align={vertical},             ymin=0,ymax=25,             ylabel={Running Time (min.)},         ]         \addplot table[x=interval,y=baseT]{\mydata};     \end{axis} \end{tikzpicture}
+
+\caption{}
+\end{figure}
+\begin{figure}
+\centering
+
+\pgfplotstableread[row sep=\\,col sep=&]{     interval & cls & vst \\     Query 1     & 22.93  & 13.23 \\     Query 2     & 29.33 & 18.25 \\     Query 3   & 26.95 & 13.77 \\     }\mydata
+\begin{tikzpicture}     \begin{axis}[             ybar,             bar width=.5cm,             width=.6\textwidth,             height=.5\textwidth,             symbolic x coords={Query 1,Query 2,Query 3},             xtick=data, 	    xticklabel style={ 		inner sep=0pt, 		anchor=north east, 		rotate=45 	    },             nodes near coords,             nodes near coords align={vertical},             ymin=0,ymax=35,             ylabel={Running Time (min.)},         ]         \addplot table[x=interval,y=cls]{\mydata}; 	\addplot table[x=interval,y=vst]{\mydata};     \end{axis} \end{tikzpicture}
+
+\caption{}
+
+\end{figure}
+\begin{figure}
+\pgfplotstableread[row sep=\\,col sep=&]{     interval  & cls   & vst   & vstF  \\     Query 4   & 108.02 & 88.63 & 0 \\     Query 5   & 134.33 & 97.87 & 86.17 \\     Query 6   & 136.38 & 135.17 & 100.98 \\     Query 7   & 117.67 & 381.02 & 188.15 \\     }\mydata
+\begin{tikzpicture}     \begin{axis}[             ybar,             bar width=.5cm,             width=\textwidth,             height=.5\textwidth,             symbolic x coords={Query 4,Query 5,Query 6,Query 7},             xtick=data, 	    xticklabel style={ 		inner sep=0pt, 		anchor=north east, 		rotate=45 	    },             nodes near coords,             nodes near coords align={vertical},             ymin=0,ymax=400,             ylabel={Running Time (min.)},         ]         \addplot table[x=interval,y=cls]{\mydata}; 	\addplot table[x=interval,y=vst]{\mydata}; 	\addplot table[x=interval,y=vstF]{\mydata};     \end{axis} \end{tikzpicture}
+
+\caption{}
+
+\end{figure}
+\begin{figure}
+\pgfplotstableread[row sep=\\,col sep=&]{     interval  & cls   & vst  \\     Query 8   & 38.52 & 14.08 \\     Query 9   & 34.55 & 14.10 \\     Query 10  & 50.08 & 16.6 \\     }\mydata
+\begin{tikzpicture}     \begin{axis}[             ybar,             bar width=.5cm,             width=\textwidth,             height=.5\textwidth,             symbolic x coords={Query 8,Query 9,Query 10},             xtick=data, 	    xticklabel style={ 		inner sep=0pt, 		anchor=north east, 		rotate=45 	    },             nodes near coords,             nodes near coords align={vertical},             ymin=0,ymax=60,             ylabel={Running Time (min.)},         ]         \addplot table[x=interval,y=cls]{\mydata}; 	\addplot table[x=interval,y=vst]{\mydata};     \end{axis} \end{tikzpicture}
+
+\caption{}
+
+\end{figure}
+\begin{figure}
+\pgfplotstableread[row sep=\\,col sep=&]{     interval  & cls   & vst  \\     Query 11  & 82.63 & 30.98 \\     Query 12  & 117.57 & 47.72 \\     Query 13  & 132.18 & 61.7 \\     Query 14  & 133.63 & 79.05 \\     }\mydata
+\begin{tikzpicture}     \begin{axis}[             ybar,             bar width=.5cm,             width=\textwidth,             height=.5\textwidth,             symbolic x coords={Query 11,Query 12,Query 13,Query 14},             xtick=data, 	    xticklabel style={ 		inner sep=0pt, 		anchor=north east, 		rotate=45 	    },             nodes near coords,             nodes near coords align={vertical},             ymin=0,ymax=200,             ylabel={Running Time (min.)},         ]         \addplot table[x=interval,y=cls]{\mydata}; 	\addplot table[x=interval,y=vst]{\mydata};     \end{axis} \end{tikzpicture}
+
+\caption{}
+
+\end{figure}
 
@@
 \bibliographystyle{apalike}
-\bibliography{xampl}
+\nocite{*}
+\bibliography{survey,systemT1,systemT2,spanners,datalog,regex,vm,regex2,obdd,kleenex}
 
 \selectlanguage{american}%
+
 \appendix

### 2017-03-02 project_sbt/arc/src/main/scala/Main.scala

@@ -1,0 +1,307 @@
+import java.io._
+
+import be.ac.ulb.arc.core.{AQLCoreFragmentSpecificationReader, CoreSpannerFileReader, CoreSpannerGenerator}
+import be.ac.ulb.arc.runtime.{ClassicalImplementation, OutputWriter}
+
+import scala.collection.mutable.{HashSet => VSRelation, ArrayBuffer}
+import scala.io._
+import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
+
+/**
+  * Created by andrea on 21/06/16.
+  */
+package object Main {
+
+  val defaultOutputTuplesFileName = "tuples"
+  val defaultOutputSpannerFileName = "spanner.csp"
+
+  def main(args: Array[String]):Unit = {
+
+    val coreSpannerFilePattern = ".+\\.csp"
+    val AQLSpecificationFilePattern = ".+\\.aqls"
+    val illegalCharPattern = ".*[^a-zA-Z0-9_./-].*"
+
+    try {
+
+      if(args.size == 0) throw new MissingArgumentException("Not enough arguments.")
+      // If the input file is a core spanner file
+      if(args.last.matches(coreSpannerFilePattern)) {
+
+        // Get the input spanner
+        if(!(new File(args.last)).exists) throw new IllegalArgumentException("The file" + args.last + " does not exist.")
+        val spannerOpt = CoreSpannerFileReader.getCoreSpanner(args.last)
+        if(spannerOpt == None) throw new IllegalArgumentException("The input spanner is not valid.")
+        val spanner = spannerOpt.get
+
+        // look for the input documents folder
+        val docsI = args.indexOf("-d")
+        if(docsI == -1) throw new MissingArgumentException("Missing argument: -d")
+        val docsFolder = args(docsI + 1)
+        val docsFolderFile = new File(docsFolder)
+        if(!docsFolderFile.exists) throw new IllegalArgumentException("The folder" + docsFolder + " does not exist.")
+        if(!docsFolderFile.isDirectory) throw new IllegalArgumentException(docsFolder + " is not a directory.")
+
+        var outputFileName = defaultOutputTuplesFileName
+
+        // Look for possible output filename
+        val outI = args.indexOf("-o")
+        if(outI != -1) {
+
+          val fileName = args(outI + 1)
+          if(fileName.matches(illegalCharPattern)) throw new IllegalArgumentException("The name " + fileName + " is not a valid filename.")
+          outputFileName = fileName
+        }
[…]

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/core/AQLCoreFragmentSpecification.scala

@@ -1,6 +1,6 @@
 package be.ac.ulb.arc.core
 
-import scala.collection.immutable.{HashSet => VSRelation}
+import scala.collection.mutable.{HashSet => VSRelation}
 import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
 import be.ac.ulb.arc.runtime.ClassicalInterpreter
 
@@ -15,6 +15,10 @@ import scala.util.control.Breaks._
   * @param operations
   */
 class AQLCoreFragmentSpecification(val spanners:CoreSpannersCollection[String, CoreSpanner], val operations:Operations[Operation]) {
+
+  // Obtain the NFA program representations of the base spanners
+  for((k, s) <- spanners)
+    s.toNFAProgram
 
   /**
     * Executes this AQL specification on a given input document.
@@ -38,7 +42,7 @@ object AQLCoreFragmentSpecificationReader {
     * @param file
     * @return
     */
-  def getSpecification(file: String): Option[AQLCoreFragmentSpecification] = {
+  def getSpecification(file: String, specialize:Boolean = false): Option[AQLCoreFragmentSpecification] = {
 
     import scala.io.Source
 
@@ -67,6 +71,10 @@ object AQLCoreFragmentSpecificationReader {
 
       line = lineIterator.next()
     }
+
+    // Set the specialization flags to true
+    if(specialize)
+      OperationExtractors.specializeJoin = true
 
     // Acquire the operations
     for(line <- lineIterator) {

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/core/Basics.scala

@@ -3,6 +3,12 @@ package be.ac.ulb.arc.core
 import scala.collection.immutable.{HashSet => SVars}
 import scala.collection.immutable.{HashSet => SVOps}
 import scala.{Int => SVar}
+import be.ac.ulb.arc.runtime.StringPointerCollection
+import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
+import scala.collection.mutable.{HashSet => VSRelation}
+
+import scala.collection.mutable.{HashSet => RunSet}
+import scala.collection.mutable.ArrayBuffer
 
 trait ⊢/⊣
 
@@ -83,6 +89,7 @@ abstract class Transition[State](val q:State, val V:SVars[SVar], val q1:State) e
 
 /**
   * Represents an ordinary transition, with a character label.
+  *
   * @param q
   * @param σ
   * @param q1
@@ -105,7 +112,7 @@ case class OrdinaryTransition[State](override val q: State, σ: Char, override v
 
       case RangeTransition(q:State, σ:Range, v:SVars[SVar], q1:State) => {
 
-        if(this.σ >= σ.min || this.σ <= σ.max)
+        if(this.σ >= σ.min && this.σ <= σ.max)
           Some(OrdinaryTransition[State2]((this.q, q), this.σ, this.V.union(v), (this.q1, q1)))
         else None
 
@@ -238,15 +245,15 @@ case class RangeTransition[State](override val q: State, σ: Range, override val
   */
 case class OperationsTransition[State](override val q: State, S: SVOps[SVOp], override val V:SVars[SVar], override val q1: State) extends Transition[State](q, V, q1) {
 
-  override def &(other:Transition[State]):Option[Transition[State2]]  = {
+  override def &(other: Transition[State]): Option[Transition[State2]] = {
 
     other match {
 
-      case OrdinaryTransition(q:State, σ:Char, v:SVars[SVar], q1:State) => {
+      case OrdinaryTransition(q: State, σ: Char, v: SVars[SVar], q1: State) => {
 
-        for(o <- this.S) {
+        for (o <- this.S) {
 
-          if(v.contains(o.x))
+          if (v.contains(o.x))
             return None
         }
 
@@ -254,11 +261,11 @@ case class OperationsTransition[State](override val q: State, S: SVOps[SVOp], ov
 
       }
 
-      case RangeTransition(q:State, σ:Range, v:SVars[SVar], q1:State) => {
[…]

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/core/CoreSpanner.scala

@@ -3,7 +3,7 @@ package be.ac.ulb.arc.core
 import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
 import be.ac.ulb.arc.runtime._
 
-import scala.collection.immutable.{HashSet => VSRelation}
+import scala.collection.mutable.{HashSet => VSRelation}
 import scala.collection.immutable.{HashSet => SVars}
 import scala.collection.mutable.{Map, ArrayBuffer => Program}
 import scala.{Int => SVar}
@@ -11,15 +11,29 @@ import scala.collection.mutable.{Map => CoreSpannersCollection}
 import scala.collection.immutable.{HashSet => SVOps}
 import scala.collection.immutable.{HashSet => TransitionFunction}
 import scala.collection.immutable.{HashSet => StateSet}
+import scala.collection.mutable.{HashSet => RunSet}
 
 /**
   * Represents a core spanner.
   */
 class CoreSpanner(val automaton:VSetAutomaton, val equalities:Option[Set[(SVar, SVar)]]) {
 
-  // Convert the vset automaton into an NFA program
+  // Convert the vset-automaton into an NFA program
   val prog = new Program[Instruction]()
-  automaton.toNFAProgram(prog, Map[State, Int](), automaton.q0, 0)
+  var map = scala.collection.immutable.Map[State, Array[Transition[State]]]()
+
+  // Pair each state with its outgoing transitions for performance
+  for(q <- automaton.Q)
+    map = map + ((q, automaton.δ.filter((t:Transition[State]) => t.q == q && t.q1 == q).toArray ++
+      automaton.δ.filter((t:Transition[State]) => t.q == q && t.q1 != q)))
+
+  /**
+    * Produces a NFA program representation of the spanner.
+    */
+  def toNFAProgram(): Unit = {
+    automaton.toNFAProgram(prog, Map[State, Int](), automaton.q0, 0)
+  }
+
   /**
     * Evaluates the spanner on the given document.
     *
@@ -28,7 +42,7 @@ class CoreSpanner(val automaton:VSetAutomaton, val equalities:Option[Set[(SVar, 
     */
   def evaluate(doc: String):Option[VSRelation[VSTuple]]= {
 
-    val eqs = if(equalities != None) equalities.get else Set[(SVar, SVar)]()
+    /*val eqs = if(equalities != None) equalities.get else Set[(SVar, SVar)]()
 
[…]

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/core/CoreSpannerFileReader.scala

@@ -153,10 +153,16 @@ object CoreSpannerFileReader {
         δ = δ + new OrdinaryTransition[State](q, '.', V, q1)
         δ = δ + new OrdinaryTransition[State](q, ';', V, q1)
       }
+      else if(label.matches("NEWLINE")) {
+
+        δ = δ + new OrdinaryTransition[State](q, '\n', V, q1)
+      }
     }
 
     source.close
 
-    (Some(new CoreSpanner(new VSetAutomaton(Q, q0, qf, V, δ), eqsSet)))
+    val result = new CoreSpanner(new VSetAutomaton(Q, q0, qf, V, δ), eqsSet)
+    result.toNFAProgram
+    Some(result)
   }
 }

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/core/Operation.scala

@@ -2,7 +2,7 @@ package be.ac.ulb.arc.core
 
 import scala.collection.mutable.{Map => CoreSpannersCollection}
 import scala.collection.mutable.{Map => VSRelationsCollection}
-import scala.collection.immutable.{HashSet => VSRelation}
+import scala.collection.mutable.{HashSet => VSRelation}
 import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
 import be.ac.ulb.arc.runtime.ClassicalImplementation
 import scala.collection.immutable.{HashSet => SVars}
@@ -141,24 +141,12 @@ case class ⋈(val a1:String, val a2:String, val a3:String) extends Operation(a3
 
       var a3Eqs:Option[Set[(SVar, SVar)]] = None
 
-      // Get the union of the equalities
-      if(_a1.equalities != None || _a2.equalities != None) {
+      val _a3 = aut3Opt.get
 
-        var eqSet = Set[(SVar, SVar)]()
-        if(_a1.equalities != None)
-          eqSet = eqSet ++ _a1.equalities.get
-        if(_a2.equalities != None)
-          eqSet = eqSet ++ _a2.equalities.get
+      // Add the result to the spanner collection
+      spanners += ((a3, _a3))
 
-        a3Eqs = Some(eqSet)
-
-        val _a3 = aut3Opt.get
-
-        // Add the result to the spanner collection
-        spanners += ((a3, _a3))
-
-        return Some(_a3)
-      }
+      return Some(_a3)
 
     }
     None
@@ -196,24 +184,12 @@ case class ∪(val a1:String, val a2:String, val a3:String) extends Operation(a3
 
       var a3Eqs:Option[Set[(SVar, SVar)]] = None
 
-      // Get the union of the equalities
-      if(_a1.equalities != None || _a2.equalities != None) {
+      val _a3 = aut3Opt.get
 
-        var eqSet = Set[(SVar, SVar)]()
-        if(_a1.equalities != None)
-          eqSet = eqSet ++ _a1.equalities.get
-        if(_a2.equalities != None)
-          eqSet = eqSet ++ _a2.equalities.get
+      // Add the result to the spanner collection
+      spanners += ((a3, _a3))
 
[…]

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/core/VSetAutomaton.scala

@@ -222,12 +222,21 @@ class VSetAutomaton(val Q:StateSet[State], val q0:State, val qf:State, val V:SVa
     // Create the new final and initial states
     var newQ0 = new State
     var newQf = new State
+    newQ = newQ + newQ0 + newQf
 
     // Connect the original and initial final states to the new ones
     newδ = newδ + new OperationsTransition[State](newQ0, new SVOps, this.V, this.q0)
     newδ = newδ + new OperationsTransition[State](newQ0, new SVOps, this.V, other.q0)
     newδ = newδ + new OperationsTransition[State](this.qf, new SVOps, this.V, newQf)
     newδ = newδ + new OperationsTransition[State](other.qf, new SVOps, this.V, newQf)
+
+    // Eliminate loops on old initial and final states
+    newδ = newδ.diff(newδ.filter((t:Transition[State]) => (t.q == this.qf && t.q1 == this.qf) || (t.q == other.qf && t.q1 == other.qf)))
+    newδ = newδ.diff(newδ.filter((t:Transition[State]) => (t.q == this.q0 && t.q1 == this.q0) || (t.q == other.q0 && t.q1 == other.q0)))
+
+    //Unanchored matching
+    newδ = newδ + new RangeTransition[State](newQ0, new Range(Char.MinValue, Char.MaxValue), this.V, newQ0)
+    newδ = newδ + new RangeTransition[State](newQf, new Range(Char.MinValue, Char.MaxValue), this.V, newQf)
 
     Some(new VSetAutomaton(newQ, newQ0, newQf, this.V, newδ))
   }
@@ -319,7 +328,11 @@ class VSetAutomaton(val Q:StateSet[State], val q0:State, val qf:State, val V:SVa
 
     val result = new VSetAutomaton(newQ, q0, qf, t.V.union(o.V), newδ)
 
-    if(closed) Some(result.removeStates(result.hasOnlyOperationsTransitions))
+    if(closed) {
+      val reduced = result.removeStates(result.hasOnlyOperationsTransitions)
+      val  (prδ, prQ) = reduced.prune[State](reduced.δ, reduced.Q, reduced.q0, reduced.qf)
+      Some(new VSetAutomaton(prQ, reduced.q0, reduced.qf, reduced.V, prδ))
+    }
     else Some(result)
   }
 
@@ -355,9 +368,7 @@ class VSetAutomaton(val Q:StateSet[State], val q0:State, val qf:State, val V:SVa
[…]

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/package.scala

@@ -1,189 +1,0 @@
-package be.ac.ulb.arc
-
-import core.CoreSpannerFileReader
-import runtime.OutputWriter
-import core.AQLCoreFragmentSpecificationReader
-import core.CoreSpannerGenerator
-
-import java.io._
-import scala.io._
-
-/**
-  * Created by andrea on 21/06/16.
-  */
-package object arc {
-
-  val defaultOutputTuplesFileName = "tuples"
-  val defaultOutputSpannerFileName = "spanner.csp"
-
-  def main(args: Array[String]):Unit = {
-
-    val coreSpannerFilePattern = ".+\\.csp"
-    val AQLSpecificationFilePattern = ".+\\.aqls"
-    val illegalCharPattern = ".*[^a-zA-Z0-9_./-].*"
-
-    try {
-
-      if(args.size == 0) throw new MissingArgumentException("Not enough arguments.")
-      // If the input file is a core spanner file
-      if(args.last.matches(coreSpannerFilePattern)) {
-
-        // Get the input spanner
-        if(!(new File(args.last)).exists) throw new IllegalArgumentException("The file" + args.last + " does not exist.")
-        val spannerOpt = CoreSpannerFileReader.getCoreSpanner(args.last)
-        if(spannerOpt == None) throw new IllegalArgumentException("The input spanner is not valid.")
-        val spanner = spannerOpt.get
-
-        // look for the input documents folder
-        val docsI = args.indexOf("-d")
-        if(docsI == -1) throw new MissingArgumentException("Missing argument: -d")
-        val docsFolder = args(docsI + 1)
-        val docsFolderFile = new File(docsFolder)
-        if(!docsFolderFile.exists) throw new IllegalArgumentException("The folder" + docsFolder + " does not exist.")
-        if(!docsFolderFile.isDirectory) throw new IllegalArgumentException(docsFolder + " is not a directory.")
-
-        var outputFileName = defaultOutputTuplesFileName
-
-        // Look for possible output filename
-        val outI = args.indexOf("-o")
-        if(outI != -1) {
-
-          val fileName = args(outI + 1)
-          if(fileName.matches(illegalCharPattern)) throw new IllegalArgumentException("The name " + fileName + " is not a valid filename.")
-          outputFileName = fileName
-        }
-
-        // Open output file
-        val outputFile = new File(outputFileName)
-        // Check if it already exists
[…]

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/ClassicalImplementation.scala

@@ -2,7 +2,7 @@ package be.ac.ulb.arc.runtime
 
 import scala.collection.immutable.{HashSet => SVars}
 import scala.{Int => SVar}
-import scala.collection.immutable.{HashSet => VSRelation}
+import scala.collection.mutable.{HashSet => VSRelation}
 import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
 import scala.collection.mutable.{Map => VSRelationsCollection}
 import be.ac.ulb.arc.core.AQLCoreFragmentSpecification
@@ -13,7 +13,7 @@ object ClassicalInterpreter {
 
   /**
     * Executes an AQL fragment on a document.
- *
+    *
     * @param fragment
     * @param doc
     * @param lazyEv
@@ -41,6 +41,8 @@ object ClassicalInterpreter {
       for(op <- fragment.operations) {
 
         result = op.execute(fragment.spanners, relations, doc)
+        if(result != None)
+          relations += ((op.res, result.get))
       }
     }
     else {
@@ -61,9 +63,114 @@ object ClassicalInterpreter {
   */
 object ClassicalImplementation {
 
+  def Ωo(table:VSRelation[VSTuple], v:SVar):VSRelation[VSTuple] = {
+
+    if(table.size <= 1) return table
+
+    val par = new Array[Int](1)
+    par(0) = 0
+    var result = new VSRelation[VSTuple]
+    val V = new SVars + v
+
+    // Sort on beginning index
+    var list = sort(table.toList, v, follows, par)
+
+    var list2 = list.tail
+
+    var current = list.head
+
+    while(!list2.isEmpty) {
+
+      while(!list2.isEmpty && current(v*2 + 1) >= list2.head(v*2)) {
+
+        current = mergeTuples(current, V, v, list2.head, V, v, V, v)
+        list2 = list2.tail
+      }
+
+      result += current
+
+      if(!list2.isEmpty) {
+
+        list = list2
+        current = list2.head
+        list2 = list2.tail
+      }
+    }
+
+    result += current
+
+    result
+
+  }
+
+  def β(table:VSRelation[VSTuple], v:SVar, dist:Int, count:Int):List[VSTuple] = {
+
+    if(table.size <= 1) return List[VSTuple]() ++ table
+
+    val par = new Array[Int](1)
+    par(0) = 0
+    var result = List[VSTuple]()
+    val V = new SVars + v
+
[…]

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/OutputWriter.scala

@@ -1,11 +1,10 @@
 package be.ac.ulb.arc.runtime
 
 import scala.{Int => Position}
-import scala.collection.immutable.{HashSet => VSRelation}
 import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
 import scala.collection.mutable.ArrayBuffer
 
-import java.io.FileWriter
+import java.io.BufferedWriter
 
 /**
   * A writer for outputs of NFA programs executed on input strings.
@@ -17,7 +16,7 @@ object OutputWriter {
     * @param input
     * @param output
     */
-  def printOutput(input:String, output:VSRelation[VSTuple]):Unit = {
+  def printOutput(input:String, output:Iterable[VSTuple], lengthConstraint:Int = -1):Unit = {
 
     for(t <- output) {
 
@@ -29,6 +28,7 @@ object OutputWriter {
         spans = newSpans
 
         print("(" + span(0) + "," + span(1) + ") ")
+
       }
 
       spans = new ArrayBuffer[Position]() ++ t.toArray
@@ -52,11 +52,13 @@ object OutputWriter {
     * @param output
     * @param outputFileWriter
     */
-  def writeOutput(input:String, output:VSRelation[VSTuple], outputFileWriter:FileWriter):Unit = {
+  def writeOutput(input:String, output:Iterable[VSTuple], outputFileWriter:BufferedWriter, lengthConstraint:Int = -1):Unit = {
 
     for(t <- output) {
 
       var spans = new ArrayBuffer[Position]() ++ t.toArray
+
+      var printed = 0
 
       while(!spans.isEmpty) {
 
@@ -64,6 +66,7 @@ object OutputWriter {
         spans = newSpans
 
         outputFileWriter.write("(" + span(0) + "," + span(1) + ") ")
+
       }
 
       spans = new ArrayBuffer[Position]() ++ t.toArray
@@ -76,9 +79,11 @@ object OutputWriter {
         if(span(0) < span(1))
           outputFileWriter.write(" " + input.substring(span(0), span(1)) + " ")
         else outputFileWriter.write(" ε ")
+
       }
       outputFileWriter.write('\n')
 
     }
   }
+
 }

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/Thread.scala

@@ -82,11 +82,7 @@ class StringPointerArray(override val vars:SVars[SVar]) extends StringPointerCol
   }
 }
 
-/**
-  * Represents a tree of string pointers.
-  *
-  * @param size
-  */
+
 /*
 class StringPointerTree(override val vars:SVars[SVar]) extends StringPointerCollection(vars) { self =>
 

### 2017-03-02 project_sbt/arc/src/main/scala/be.ac.ulb.arc/runtime/VirtualMachine.scala

@@ -4,7 +4,7 @@ import scala.collection.immutable.{HashSet => SVars}
 import scala.{Int => SVar}
 import scala.{Int => Position}
 import scala.{Array => Program}
-import scala.collection.immutable.{HashSet => VSRelation}
+import scala.collection.mutable.{HashSet => VSRelation}
 import be.ac.ulb.arc.runtime.{StringPointerCollection => VSTuple}
 import scala.collection.mutable.{MutableList => ThreadList}
 

### 2017-03-02 project_sbt/arc/src/test/scala/be.ac.ulb.arc/benchmark/benchmarkTestSuite.scala

@@ -1,8 +1,10 @@
 package be.ac.ulb.arc.benchmark
 
-import be.ac.ulb.arc.runtime.OutputWriter
-import be.ac.ulb.arc.core.{CoreSpannerFileReader, OrdinaryTransition, Transition, State}
+import be.ac.ulb.arc.runtime.{ClassicalImplementation, OutputWriter}
+import be.ac.ulb.arc.core._
 import org.scalatest.FunSuite
+
+import scala.io.{BufferedSource, Source}
 /**
   * Created by andrea on 20/06/16.
   */
@@ -17,6 +19,18 @@ class benchmarkTestSuite extends FunSuite{
     val sentimentFile = "src/test/scala/be.ac.ulb.arc/benchmark/sentiment.csp"
     val genreFile = "src/test/scala/be.ac.ulb.arc/benchmark/genre.csp"
     val titleFile = "src/test/scala/be.ac.ulb.arc/benchmark/title.csp"
+    val nameFile = "src/test/scala/be.ac.ulb.arc/benchmark/name.csp"
+
+    val documentFile = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt/189575.female.23.indUnk.Libra.xml"
+    val documentFile2 = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt/476271.female.25.Publishing.Pisces.xml"
+    val documentFile3 = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt/317450.female.25.indUnk.Taurus.xml"
+    val fragmentFile = "src/test/scala/be.ac.ulb.arc/benchmark/movieFollowsTitle.aqls"
+    val fragmentFile2 = "src/test/scala/be.ac.ulb.arc/benchmark/test2.aqls"
+    val fragmentFile3 = "src/test/scala/be.ac.ulb.arc/benchmark/test3.aqls"
+    val fragmentFile5 = "src/test/scala/be.ac.ulb.arc/benchmark/test5.aqls"
+    val fragmentFile6 = "src/test/scala/be.ac.ulb.arc/benchmark/sentiment_genre_movie.aqls"
+    val fragmentFile7 = "src/test/scala/be.ac.ulb.arc/benchmark/keywords'.aqls"
+    val fragmentFile8 = "src/test/scala/be.ac.ulb.arc/benchmark/title''.aqls"
 
     val action = CoreSpannerFileReader.getCoreSpanner(actionFile).get
[…]

### 2017-03-02 project_sbt/arc/src/test/scala/be.ac.ulb.arc/core/coreSpannerTestSuite.scala

@@ -24,7 +24,6 @@ class coreSpannerTestSuite extends FunSuite{
     val keywordsFile = "src/test/scala/be.ac.ulb.arc/benchmark/keywords.csp"
     val titleFile = "src/test/scala/be.ac.ulb.arc/benchmark/title.csp"
 
-
     val spanner = CoreSpannerFileReader.getCoreSpanner(spannerFile1).get
     val spanner3 = CoreSpannerFileReader.getCoreSpanner(spannerFile3).get
     val spanner4 = CoreSpannerFileReader.getCoreSpanner(spannerFile4).get
@@ -35,7 +34,6 @@ class coreSpannerTestSuite extends FunSuite{
     val spanner9 = CoreSpannerFileReader.getCoreSpanner(spannerFile9).get
     val spanner10 = CoreSpannerFileReader.getCoreSpanner(spannerFile10).get
     val keywords = CoreSpannerFileReader.getCoreSpanner(keywordsFile).get
-    val title = CoreSpannerFileReader.getCoreSpanner(titleFile).get
   }
 
   test("A core spanner should be correctly converted into an NFA program") {
@@ -213,8 +211,4 @@ class coreSpannerTestSuite extends FunSuite{
     }
   }
 
-  test("toString method should give a string representation of a spanner") {
-
-    println(data.title.toString())
-  }
 }

### 2017-03-02 project_sbt/arc/src/test/scala/be.ac.ulb.arc/mainTestSuite.scala

@@ -1,6 +1,7 @@
 package be.ac.ulb.arc
 
 import be.ac.ulb.arc.core.{AQLCoreFragmentSpecificationReader, CoreSpannerFileReader, CoreSpannerGenerator}
+import be.ac.ulb.arc.runtime.OutputWriter
 import org.scalatest.FunSuite
 
 import scala.io.Source
@@ -13,38 +14,123 @@ class mainTestSuite extends FunSuite{
 
     val args = new Array[String](5)
     args(0) = "-d"
-    args(1) = "/home/andrea/Documenti/ulb/thesis/amorcian/project_sbt/arc/src/test/scala/be.ac.ulb.arc/benchmark/excerpt"
+    args(1) = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt"
     args(2) = "-o"
-    args(3) = "/home/andrea/Documenti/ulb/thesis/output/output.txt"
+    args(3) = "/home/output/output.txt"
     args(4) = "src/test/scala/be.ac.ulb.arc/benchmark/title.csp"
 
     val args1 = new Array[String](5)
     args1(0) = "-d"
-    args1(1) = "/home/andrea/Documenti/ulb/thesis/amorcian/project_sbt/arc/src/test/scala/be.ac.ulb.arc/benchmark/excerpt"
+    args1(1) = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt"
     args1(2) = "-o"
-    args1(3) = "/home/andrea/Documenti/ulb/thesis/output/output.txt"
+    args1(3) = "/home/output/output.txt"
     args1(4) = "src/test/scala/be.ac.ulb.arc/benchmark/movieFollowsTitle.aqls"
 
     val args2 = new Array[String](4)
     args2(0) = "-c"
     args2(1) = "-o"
-    args2(2) = "/home/andrea/Documenti/ulb/thesis/output/spanner"
+    args2(2) = "/tmp/movieFollowedByTitle"
     args2(3) = "src/test/scala/be.ac.ulb.arc/benchmark/movieFollowsTitle.aqls"
+
+    val args3 = new Array[String](6)
+    args3(0) = "-d"
+    args3(1) = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt"
+    args3(2) = "-s"
+    args3(3) = "-o"
+    args3(4) = "/home/output/output.txt"
[…]

### 2017-03-02 project_sbt/arc/src/test/scala/be.ac.ulb.arc/runtime/classicalImplementationTestSuite.scala

@@ -1,14 +1,13 @@
 package be.ac.ulb.arc.runtime
 
-import be.ac.ulb.arc.core.CoreSpannerFileReader
+import be.ac.ulb.arc.core.{AQLCoreFragmentSpecificationReader, CoreSpannerFileReader}
+
 import scala.collection.immutable.{HashSet => SVars}
 import org.scalatest.FunSuite
 /**
   * Created by andrea on 19/06/16.
   */
 class classicalImplementationTestSuite extends FunSuite{
-
-
 
   val data = new {
 
@@ -25,6 +24,11 @@ class classicalImplementationTestSuite extends FunSuite{
     val spannerFile11 = "src/test/scala/be.ac.ulb.arc/vset/spanner11.txt"
     val spannerFile12 = "src/test/scala/be.ac.ulb.arc/vset/spanner12.txt"
     val spannerFile13 = "src/test/scala/be.ac.ulb.arc/vset/spanner13.txt"
+
+    val documentFile = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt/189575.female.23.indUnk.Libra.xml"
+    val documentFile2 = "src/test/scala/be.ac.ulb.arc/benchmark/excerpt/476271.female.25.Publishing.Pisces.xml"
+    val fragmentFile = "src/test/scala/be.ac.ulb.arc/benchmark/movieFollowsTitle.aqls"
+    val fragmentFile2 = "src/test/scala/be.ac.ulb.arc/benchmark/test2.aqls"
 
     val spanner = CoreSpannerFileReader.getCoreSpanner(spannerFile1).get
     val spanner3 = CoreSpannerFileReader.getCoreSpanner(spannerFile3).get
@@ -126,4 +130,6 @@ class classicalImplementationTestSuite extends FunSuite{
 
     for(t <- tNEq) assert(s.substring(t(0), t(1)) != s.substring(t(2), t(3)))
   }
+
+
 }
